---
title: "Traffic Collisions Data Cleaning"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Data for File 5. Demographic Trends
This script is intended to clean some of the raw data collected from the US Census Bureau. As these transformations can take a long time to run, this script is separated from the rest of the data transformations (located in file 2. Data Transformations), and the outputs here are saved to their own CSVs. This script should not have to be run more than once.

# Census data
```{r}
parse_census_data <- function(year) {
  
  census_overall <- read.csv(paste('Data/Census Data/demographics_',
                                 year, '.csv', sep='')) %>%
      mutate(Year = local(year),
             Demographic = str_trim(`Label..Grouping.`),
             Demographic = case_when(
               Demographic %in% c('White Alone') ~ 'White',
               Demographic %in% c('Black or African American alone') ~ 'Black',
               Demographic %in% c('Asian alone') ~ 'Asian',
               Demographic %in% c('Some Other Race alone') ~ 'Other',
               Demographic %in% c('Two or More Races') ~ 'Two or More Races',
               Demographic %in% c('Hispanic or Latino (of any race)') ~ 'Hispanic',
               TRUE ~ Demographic),
             DemographicType = case_when(
               Demographic %in% c('Male', 'Female') ~ 'Sex',
               Demographic %in% c('Under 5 years', '5 to 9 years', '10 to 14 years', 
                                  '15 to 19 years', '20 to 24 years', '25 to 34 years', 
                                  '35 to 44 years', '45 to 54 years', '55 to 59 years', 
                                  '60 to 64 years', '65 to 74 years', '75 to 84 years', 
                                  '85 years and over') ~ 'Age',
               Demographic %in% c('White', 'Black', 'Asian', 'Other', 
                                  'Two or More Races', 'Hispanic') ~ 'RaceEthnicity',
               TRUE ~ NA),
             Population = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ',')),
             TotalPopulation = ifelse(Demographic %in% c('Total population'), Population, NA)) %>%
      group_by(Year) %>% mutate(TotalPopulation = max(TotalPopulation, na.rm=TRUE)) %>% ungroup() %>%
      filter(!is.na(DemographicType)) %>%
      select(Year, DemographicType, Demographic, Population, TotalPopulation)

  return(census_overall)
}

census_data <- data.frame()

for (year in c(2013:2024)) {census_data <- rbind(census_data, parse_census_data(year))}

census_data %>% write.csv('Cleaned Data/LACensusDataOverall.csv', row.names=FALSE)
```

# Commutes over Time
## Loading and cleaning data
```{r}
pivot_commuting_overall <- function(year) {
  commuting_overall <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                               year, '.csv', sep='')) %>%
    mutate(Year = local(year),
           Labels = str_trim(`Label..Grouping.`),
           Labels = case_when(
             Labels == 'Taxi or ride-hailing services' ~ 'Taxicab', 
             Labels == 'Public transportation (excluding taxicab):' ~ 'Public transportation:',
             Labels == 'Worked at home' ~ 'Worked from home',
             TRUE ~ Labels),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, Labels, Values) %>%
    filter(Labels %in% c('Total:', 'Drove alone', 'Carpooled:', 'Public transportation:', 
                         'Walked', 'Taxicab', 'Motorcycle', 'Bicycle', 'Other means', 'Worked from home')) %>%
    pivot_wider(names_from=Labels, values_from=Values) %>%
    mutate(TaxiBikeAndOther = Taxicab + Motorcycle + Bicycle + `Other means`) %>% 
    select(Year, `Total:`, `Drove alone`, `Carpooled:`, `Public transportation:`, 
           Walked, TaxiBikeAndOther, `Worked from home`)
  
  colnames(commuting_overall) <- c('Year', 'TotalWorkers', 'Driving', 'Carpooling', 'PublicTransit', 
                                   'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(commuting_overall)
}

pivot_commuting_data_by_ethn <- function(year, file_subname, race_ethnicity_label) {
  
  ethn_df <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                            file_subname, '_', year, '.csv', sep='')) %>% 
    mutate(RaceEthnicity = local(race_ethnicity_label), Year = local(year),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, RaceEthnicity, `Label..Grouping.`, Values) %>%
    pivot_wider(names_from=`Label..Grouping.`, values_from=Values)
  
  colnames(ethn_df) <- c('Year', 'RaceEthnicity', 'TotalWorkers', 'Driving', 'Carpooling',
                         'PublicTransit', 'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(ethn_df)
}

parse_commute_data <- function(years) {
  overall <- data.frame()
  by_ethn <- data.frame()
  by_lang <- data.frame()
  
  for (year in years) {
    overall <- rbind(overall, pivot_commuting_overall(year))
    by_ethn <- rbind(by_ethn, pivot_commuting_data_by_ethn(year, 'asian', 'Asian') %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'black', 'Black')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'hispanic', 'Hispanic')) %>%
                       # rbind(pivot_commuting_data_by_ethn(year, 'white', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'white_nonhispanic', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'multiple_races', 'Two or More Races')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'other', 'Other')))
    # by_lang <- rbind(by_lang, read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_by_language_', year, '.csv', sep='')))
  }
  
  return(list(overall, by_ethn, by_lang))
}

commuting_data <- parse_commute_data(c(2013:2024))
commuting_data_overall <- commuting_data[1] %>% as.data.frame()
commuting_data_by_ethnicity <- commuting_data[2] %>% as.data.frame()

commuting_data_overall %>% write.csv('Cleaned Data/LACommutingDataOverall.csv', row.names=FALSE)
commuting_data_by_ethnicity %>% write.csv('Cleaned Data/LACommutingDataByEthnicity.csv', row.names=FALSE)
```

# Parsing ZIP Codes
```{r}
zip_code_boundaries <- read_json('Data/ZIP Codes/Zipcodes.geojson')

zip_codes_df <- data.frame()

for (i in c(1:length(zip_code_boundaries$features))) {
  row <- zip_code_boundaries$features[[i]]

  object_id <- row$properties$OBJECTID
  zip_code <- row$properties$ZIPCODE
  shape_type <- row$geometry$type

  if (shape_type == 'Polygon') {
    coordinates <- row$geometry$coordinates[[1]]
  } else {
    coordinates <- row$geometry$coordinates[[1]][[1]]
  }
  for (coord_set in coordinates) {
    zip_codes_df <- rbind(zip_codes_df,
                          list('ObjectID'=object_id, 'ZIPCode'=zip_code,
                               'Longitude'=coord_set[[1]],
                               'Latitude'=coord_set[[2]]))
  }
}

zip_codes_df <- zip_codes_df %>% distinct()
```

## Interpolating the ZIP code boundary coordinates
```{r}
zip_codes_indexed <- zip_codes_df %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  group_by(ZIPCode) %>% 
  mutate(Index =  row_number()) %>% 
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2))
```

```{r}
interpolate_coordinates <- function(distance_to_next, boundary_latitude, boundary_longitude, 
                                    next_boundary_latitude, next_boundary_longitude, interpol_distance) {
    num_steps <- floor(distance_to_next/interpol_distance)
    latitude_diff <- next_boundary_latitude - boundary_latitude
    longitude_diff <- next_boundary_longitude - boundary_longitude
    
    new_latitudes <- ''
    new_longitudes <- ''
    
    for (i in c(1:num_steps)) {
      new_latitudes <- paste(new_latitudes, ', ', boundary_latitude + latitude_diff*i/(1+num_steps))
      new_longitudes <- paste(new_longitudes, ', ', boundary_longitude + longitude_diff*i/(1+num_steps))
    }
    
    return(paste(new_latitudes, new_longitudes, sep=';'))
  }

interpolation_distance <- 0.00005

zip_codes_interpolated <- zip_codes_indexed %>%
  rowwise() %>%
  mutate(
    InterpolatedCoordinates = ifelse(
      DistanceToNext > local(interpolation_distance) & (DistanceToNext < 0.0095 | 
                                                          !(ZIPCode %in% c('90803', '90732'))), # 90803 has some non-contiguous areas 
      interpolate_coordinates(DistanceToNext, Latitude, Longitude,
                              NextLatitude, NextLongitude, local(interpolation_distance)), NA)
  ) %>%
  separate(InterpolatedCoordinates, sep=';', 
           into=c('InterpolatedLatitude', 'InterpolatedLongitude')) %>%
  separate_longer_delim(c(InterpolatedLatitude, InterpolatedLongitude), delim=', ') %>%
  mutate(
    InterpolatedLongitude = as.numeric(InterpolatedLongitude),
    InterpolatedLatitude = as.numeric(InterpolatedLatitude),
    Longitude = ifelse(is.na(InterpolatedLongitude), Longitude, InterpolatedLongitude),
    Latitude = ifelse(is.na(InterpolatedLatitude), Latitude, InterpolatedLatitude),
  ) %>% select(-InterpolatedLatitude, -InterpolatedLongitude)

zip_codes_reindexed <- zip_codes_interpolated %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  mutate(Latitude = round(Latitude, 4), Longitude = round(Longitude, 4)) %>%
  distinct() %>%
  group_by(ZIPCode) %>%
  mutate(Index = row_number()) %>%
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2 ))

zip_codes_reindexed %>% filter(DistanceToNext > 0.0001) # 0.0001414214
zip_codes_reindexed %>% mutate(LatitudeDiff = NextLatitude - Latitude) %>% filter(LatitudeDiff > 0.0001)
zip_codes_reindexed %>% mutate(LongitudeDiff = NextLongitude - Longitude) %>% filter(LongitudeDiff > 0.0001)

# zip_codes_reindexed %>% write.csv('Cleaned Data/LAZipCodesInterpolated.csv', row.names=FALSE)
```


# Breaking Down Demographics by ZIP Code
I used (all_coordinates %>% filter(CoordinateType == 'Collision') %>% select(ZIPCode) %>% distinct() %>% arrange(ZIPCode))$ZIPCode to list all the relevant ZIP codes and collected the overall census and commuter data for 2013-2024 for these ZIP codes. Each CSV contains the data for 20-40 of the ZIP codes due to filesize limits
```{r}
census_filenames <-list.files('Data/Census Data/By ZIP Code')
commuter_filenames <-list.files('Data/Commuting Data/By ZIP Code')
```

## Parsing census data
First, we need to standardize the Label Groupings, as they vary year-to-year. I've used the following script to narrow them down to the most important characteristics that are observed each year
```{r}
label_groupings <- data.frame()

parse_label_groupings <- function(label_grouping) {
  stripped_label_grouping <- str_trim(label_grouping)
  num_indents <- floor((str_length(label_grouping) - str_length(stripped_label_grouping))/4)
  
  if (num_indents == 0) {
    return(paste(stripped_label_grouping, 'None', 'None', 'None', 'None', sep=';'))
  } else if (num_indents == 1) {
    return(paste('', stripped_label_grouping, 'None', 'None', 'None', sep=';'))
  } else if (num_indents == 2) {
    return(paste('', '',  stripped_label_grouping, 'None', 'None', sep=';'))
  } else if (num_indents == 3) {
    return(paste('', '', '', stripped_label_grouping, 'None', sep=';'))
  } else if (num_indents == 4) {
    return(paste('', '', '', '', stripped_label_grouping, sep=';'))
  }
}

for (filename in census_filenames) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()
  
  label_groupings <- rbind(label_groupings, 
    read.csv(paste('Data/Census Data/By ZIP Code/', filename, sep='')) %>%
      rowwise() %>%
      mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
             Year = local(year_of_data)) %>%
      separate(LabelGroupings, sep=';', into=c('LabelGroup1',
                                               'LabelGroup2',
                                               'LabelGroup3',
                                               'LabelGroup4',
                                               'LabelGroup5')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~na_if(.,'')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5), ~toupper(.)) %>%
      fill(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
      mutate(
        LabelGroup1 = ifelse(LabelGroup1 == 'TOTAL POPULATION',
                             LabelGroup2, LabelGroup1),
        LabelGroup2 = case_when(
          LabelGroup2 == 'TOTAL POPULATION' |
            LabelGroup2 == LabelGroup1 ~ LabelGroup3,
          TRUE ~ LabelGroup2),
        LabelGroup3 = case_when(
          LabelGroup3 == 'TOTAL POPULATION' | 
            LabelGroup3 == LabelGroup2 ~ LabelGroup4,
          TRUE ~ LabelGroup3),
        LabelGroup4 = case_when(
          LabelGroup4 == 'TOTAL POPULATION' | 
            LabelGroup4 == LabelGroup3 ~ LabelGroup5,
          TRUE ~ LabelGroup4),
        LabelGroup5 = case_when(
          LabelGroup5 == 'TOTAL POPULATION' | 
            LabelGroup5 == LabelGroup4 ~ 'NONE',
          TRUE ~ LabelGroup5),
      ) %>%
      select(Year, LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5))
}

label_groupings_summarised <- label_groupings %>% 
  group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5) %>%
  summarise(Count= n_distinct(Year)) %>% arrange(desc(Count)) %>% ungroup() %>%
  filter(LabelGroup4 == 'NONE', LabelGroup5 == 'NONE', LabelGroup2 != 'NONE',
         LabelGroup3 == 'NONE' | (LabelGroup1 == 'HISPANIC OR LATINO AND RACE' & 
                                    LabelGroup2 == 'NOT HISPANIC OR LATINO') |
           (LabelGroup1 == 'RACE' & LabelGroup2 == 'ONE RACE'),
         LabelGroup3 != 'NONE' | (LabelGroup1 != 'HISPANIC OR LATINO AND RACE' & LabelGroup1 != 'RACE') |
           (LabelGroup2 != 'NOT HISPANIC OR LATINO' & LabelGroup2 != 'ONE RACE'),
         Count == 12, LabelGroup2 != 'MEDIAN AGE (YEARS)',
         !(grepl('YEARS AND OVER', LabelGroup2)) | LabelGroup2 == '85 YEARS AND OVER',
         LabelGroup2 != 'TOTAL POPULATION',
         ) %>% select(LabelGroup1, LabelGroup2, LabelGroup3)
```

Now, we can use label_groupings_summarised to only pull the most important data from each CSV
```{r}
parse_census_data_by_zip_code <- function(filename) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()

  census_data_by_zip <- read.csv(paste('Data/Census Data/By ZIP Code/', filename, sep=''))  %>%
    rowwise() %>%
    mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
           Year = local(year_of_data)) %>%
    separate(LabelGroupings, sep=';', into=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3',
                                             'LabelGroup4', 'LabelGroup5')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~na_if(.,'')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5), ~toupper(.)) %>%
    fill(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
    mutate(
      LabelGroup1 = ifelse(LabelGroup1 == 'TOTAL POPULATION', 
                           LabelGroup2, LabelGroup1),
      LabelGroup2 = case_when(
        LabelGroup2 == 'TOTAL POPULATION' | 
          LabelGroup2 == LabelGroup1 ~ LabelGroup3,
        TRUE ~ LabelGroup2),
      LabelGroup3 = case_when(
        LabelGroup3 == 'TOTAL POPULATION' | 
          LabelGroup3 == LabelGroup2 ~ LabelGroup4,
        TRUE ~ LabelGroup3),
      LabelGroup4 = case_when(
        LabelGroup4 == 'TOTAL POPULATION' | 
          LabelGroup4 == LabelGroup3 ~ LabelGroup5,
        TRUE ~ LabelGroup4),
      LabelGroup5 = case_when(
        LabelGroup5 == 'TOTAL POPULATION' | 
          LabelGroup5 == LabelGroup4 ~ 'NONE',
        TRUE ~ LabelGroup5),
    )  %>% 
    filter(LabelGroup4 == 'NONE', LabelGroup5 == 'NONE') %>%
    select(-`Label..Grouping.`, -LabelGroup4, -LabelGroup5) %>%
    select(Year, LabelGroup1, LabelGroup2, LabelGroup3, everything()) %>% 
    pivot_longer(cols=starts_with('ZCTA'), names_to = 'ZIP', values_to = 'Population', values_transform = as.character) %>%
    filter(grepl('..Estimate', ZIP))
  
  return(census_data_by_zip)
}
```

```{r}
census_by_zip_raw <- data.frame()

for (filename in census_filenames) {
  census_by_zip_raw <- rbind(census_by_zip_raw, parse_census_data_by_zip_code(filename))
}

census_by_zip <- census_by_zip_raw %>%
    mutate(ZIP = substring(ZIP, 7, 11),
           Population = str_remove(Population, ',') %>% as.numeric()) %>%
    inner_join(label_groupings_summarised, by=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3'))

# checking the completeness of the data; ZIP code 90090 is missing data for 2021-2024, but these years are simply not available in the census data
for (year in c(2013:2024)) {
  census_by_zip %>% 
    # group_by(ZIP) %>% summarise(Count = n()) %>% arrange(desc(Count))
    group_by(ZIP) %>% mutate(Count = n()) %>% filter(Count < 372, Year == local(year)) %>% ungroup() %>%
    arrange(ZIP, LabelGroup1, LabelGroup2, LabelGroup3) %>%
    print()
}

census_by_zip %>% write.csv('Cleaned Data/LACensusDataByZIP.csv', row.names=FALSE)
```

## Parsing commuting data
```{r}
label_groupings <- data.frame()

for (filename in commuter_filenames) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()
  
  label_groupings <- rbind(label_groupings, 
    read.csv(paste('Data/Commuting Data/By ZIP Code/', filename, sep='')) %>%
      rowwise() %>%
      mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
             LabelGroupings = str_replace(LabelGroupings, ':', ''),
             Year = local(year_of_data)) %>%
      separate(LabelGroupings, sep=';', into=c('LabelGroup1',
                                               'LabelGroup2',
                                               'LabelGroup3',
                                               'LabelGroup4')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3), ~na_if(.,'')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~toupper(.)) %>%
      fill(LabelGroup1, LabelGroup2, LabelGroup3) %>%
      select(Year, LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4))
}

label_groupings %>% group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>% 
  summarise(Count = n()) %>% arrange(desc(Count))

label_groupings_summarised <- label_groupings %>% 
  group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
  summarise(Count= n_distinct(Year)) %>% arrange(desc(Count)) %>% ungroup() %>%
  filter(LabelGroup4 == 'NONE', LabelGroup2 != 'NONE',
         LabelGroup3 == 'NONE' | (LabelGroup3 %in% c('DROVE ALONE', 'CARPOOLED'))
         ) %>% select(LabelGroup1, LabelGroup2, LabelGroup3)
```

```{r}
parse_commuting_data_by_zip_code <- function(filename) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()

  commuting_data_by_zip <- read.csv(paste('Data/Commuting Data/By ZIP Code/', filename, sep=''))  %>%
    rowwise() %>%
    mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
           LabelGroupings = str_replace_all(LabelGroupings, ':', ''),
           Year = local(year_of_data)) %>%
    separate(LabelGroupings, sep=';', into=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3', 'LabelGroup4')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3), ~na_if(.,'')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~toupper(.)) %>%
    fill(LabelGroup1, LabelGroup2, LabelGroup3) %>%
    filter(LabelGroup4 == 'NONE', LabelGroup2 != 'NONE') %>%
    select(-`Label..Grouping.`, -LabelGroup4) %>%
    select(Year, LabelGroup1, LabelGroup2, LabelGroup3, everything()) %>% 
    pivot_longer(cols=starts_with('ZCTA'), names_to = 'ZIP', values_to = 'Population', values_transform = as.character) %>%
    filter(grepl('..Estimate', ZIP)) %>%
    mutate(ZIP = substring(ZIP, 7, 11),
           Population = str_remove(Population, ',') %>% as.numeric()) %>%
    inner_join(label_groupings_summarised, by=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3'))
  
  return(commuting_data_by_zip)
}
```

```{r}
commuting_by_zip <- data.frame()

for (filename in commuter_filenames) {
  commuting_by_zip <- rbind(commuting_by_zip, parse_commuting_data_by_zip_code(filename))
}

# checking the completeness of the data; ZIP code 90090 is missing data for 2021-2024, but these years are simply not available in the census data
for (year in c(2013:2024)) {
  commuting_by_zip %>% 
    # group_by(ZIP) %>% summarise(Count = n()) %>% arrange(desc(Count))
    group_by(ZIP) %>% mutate(Count = n()) %>% filter(Count < 264, Year == local(year)) %>% ungroup() %>%
    arrange(ZIP, LabelGroup1, LabelGroup2, LabelGroup3) %>%
    print()
}

commuting_by_zip %>% write.csv('Cleaned Data/LACommutingDataByZIP.csv', row.names=FALSE)
```