---
title: "Traffic Collisions Data Transformatiosn"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Data for File 3. Data Exploration
```{r}
# for runtime logging
file_1_start <- Sys.time()
runtime_checkpoints <- c(file_1_start)

print_checkpoint_runtime <- function(current_time) {
  last_checkpoint <- max(runtime_checkpoints)
  runtime_checkpoints <<- c(runtime_checkpoints, current_time)
  
  print(paste('Runtime to checkpoint ', length(runtime_checkpoints) - 1, ': ', 
              format(current_time - last_checkpoint), sep=''))
}
```

# Load the Cleaned Collisions Data
```{r}
collisions_cleaned <- read.csv('Cleaned Data/LACollisionsDataCleaned.csv') %>%
  mutate(
    ReportedDate = as.Date(ReportedDate, format='%Y-%m-%d'),
    OccurredDate = as.Date(OccurredDate, format='%Y-%m-%d'),
  )
```

# Breakdown of Fields
## Time series graphs
```{r}
collisions_by_reported <- collisions_cleaned %>% 
  group_by(ReportedDate) %>% summarise(Collisions = n())
collisions_by_occurred <- collisions_cleaned %>% 
  group_by(OccurredDate) %>% summarise(Collisions = n())

collisions_by_reported_by_week <- collisions_cleaned %>% 
  group_by(ReportedDate = floor_date(ReportedDate, unit='weeks')) %>% 
  summarise(Collisions = n())
collisions_by_occurred_by_week <- collisions_cleaned %>% 
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())
```

```{r}
collisions_by_week <- collisions_cleaned %>% 
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())

collisions_by_area_and_week <- collisions_cleaned %>% 
  group_by(AreaName, OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print_checkpoint_runtime(Sys.time())
```

### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- collisions_cleaned %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
```

### Occurred time distribution
```{r}
occurred_time_dist_over_time <- collisions_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(OccurredDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())
```

## Scatterplot of Latitude, Longitude, 
```{r, fig.width=6, fig.height=3.5}
rounding_factor <- 100

collisions_by_coordinates_and_year <- collisions_cleaned %>% 
  group_by(Address, CrossStreet, Latitude, Longitude, OccurredYear = year(OccurredDate)) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Latitude = floor(Latitude*local(rounding_factor))/local(rounding_factor), 
           Longitude = floor(Longitude*local(rounding_factor))/local(rounding_factor),
           OccurredYear) %>% 
  summarise(TotalCollisions = sum(Collisions),
            AvgCollisionsPerLocation = mean(Collisions, na.rm=TRUE),
            NumCollisionRecords = n()) %>%
  filter(Latitude != 0 | Longitude != 0)

print_checkpoint_runtime(Sys.time())
```

# Forecasting Overall Collision Counts
## Pre-COVID
```{r}
collisions_over_time_pre_covid <- collisions_cleaned %>% 
  filter(OccurredDate < as.Date('2020-03-15')) %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>%
  summarise(Collisions = n())
```

```{r}
collisions_over_time_post_covid <- collisions_cleaned %>% 
  filter(OccurredDate >= as.Date('2021-01-03')) %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>%
  summarise(Collisions = n())

print_checkpoint_runtime(Sys.time())
```

# Identifying Collision Hot Spots
```{r}
rounding_factor <- 100

collisions_by_location_2024 <- collisions_cleaned %>%
  mutate(OccurredYear = year(OccurredDate)) %>%
  filter(OccurredYear == 2024) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Latitude = round(Latitude*local(rounding_factor))/local(rounding_factor), 
           Longitude = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>% 
  summarise(TotalCollisions = sum(Collisions),
            AvgCollisionsPerLocation = mean(Collisions, na.rm=TRUE),
            NumCollisionRecords = n(),
            RangeOfCollisions = max(Collisions) - min(Collisions),
            StdCollisions = sd(Collisions)
            ) %>%
  filter(Latitude != 0 | Longitude != 0)
```

## Classifying the MO Codes, based on the PDF I found
```{r}
collision_mo_codes <- collisions_cleaned %>% 
  mutate(MOCode = str_replace(MOCodes, ' ', ', ')) %>%
  separate_rows(MOCode, convert=TRUE)
```

# Data for File 4. MO Code Analysis
```{r}
# trying again to round out the intersection areas by name and coordinates
collisions_at_intersection_areas <- collisions_cleaned %>%
  filter(Latitude != 0, Longitude != 0) %>%
  mutate(
    HasCrossStreet = as.numeric(!is.na(CrossStreet) & !(CrossStreet == '')),
    NearestIntersection = ifelse(
      HasCrossStreet,  paste0(Address, ' and ', CrossStreet), Address)) %>%
  group_by(
    HasCrossStreet, NearestIntersection,
    RoundedLat = round(Latitude, 4),
    RoundedLon = round(Longitude, 4)) %>%
  summarise(Count = n(), 
            ProportionListedAsAtIntersection = mean(AtIntersection),
            ProportionListedAsNotAtIntersection = mean(NotAtIntersection),
            NumCollisionsAtIntersection = sum(AtIntersection),
            NumCollisionsNotAtIntersection = sum(NotAtIntersection),
            ) %>%
  arrange(desc(Count)) %>%
  group_by(HasCrossStreet, RoundedLat, RoundedLon) %>%
  summarise(
    Collisions = sum(Count),
    NumIntersections = n(),
    NumCollisionsAtIntersection = sum(NumCollisionsAtIntersection),
    NumCollisionsNotAtIntersection = sum(NumCollisionsNotAtIntersection),
    ProportionListedAsAtIntersection = NumCollisionsAtIntersection/Collisions,
    ProportionListedAsNotAtIntersection = NumCollisionsNotAtIntersection/Collisions,
    NearestIntersections = paste(NearestIntersection, collapse=', ')
  ) %>% 
  mutate(
    ListedAsIntersection = as.numeric(ProportionListedAsAtIntersection > 0),
    ListedAsNotIntersection = as.numeric(ProportionListedAsNotAtIntersection > 0),
    ListedMoreAsAtIntersection = (ProportionListedAsAtIntersection > 
                                    ProportionListedAsNotAtIntersection)
  )

print_checkpoint_runtime(Sys.time())
```

# Collision Object
```{r}
collisions_by_object <- collisions_cleaned %>%
    filter(CollisionWithVehicle) %>%
    mutate(CollisionObject = 'Vehicle') %>% 
  rbind(
    collisions_cleaned %>%
      filter(CollisionWithParkedVehicle) %>%
      mutate(CollisionObject = 'Parked Vehicle')
  ) %>% 
  rbind(
    collisions_cleaned %>%
      filter(CollisionWithFixedObject) %>%
      mutate(CollisionObject = 'Fixed Object')
  ) %>% 
  rbind(
    collisions_cleaned %>%
      filter(CollisionWithPedestrian) %>%
      mutate(CollisionObject = 'Pedestrian')
  ) %>% 
  rbind(
    collisions_cleaned %>%
      filter(CollisionWithBike) %>%
      mutate(CollisionObject = 'Bike') 
  ) %>% 
  rbind(
    collisions_cleaned %>%
      filter(CollisionWithOtherObject) %>%
      mutate(CollisionObject = 'Other') 
  ) %>% 
  select(CollisionObject, OccurredDate, OccurredYear, Latitude, Longitude, 
         CollisionAtIntersection, InjuryInvolved, MostSevereInjuryReported,
         NonInjury, ComplaintOfInjury, VisibleInjury, SevereInjury, FatalInjury)

print_checkpoint_runtime(Sys.time())
```

# Defining Functions for Future Analysis
```{r}
intro_graphs <- function(graph_data, field_name, label=field_name, year=2024, text_size=5) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))
  
  # histogram
  if (local(year) == 'all') {
    print(graph_data %>%
            filter(OccurredDate >= '2013-01-01',
                   OccurredDate < '2025-01-01') %>%
            group_by(TargetFeature) %>%
            summarise(Collisions = n()) %>%
            ggplot(aes(x=TargetFeature, y=Collisions, fill=TargetFeature)) +
            geom_bar(stat='identity') +
            theme_set(theme_gray(base_size = text_size)) +
            theme(axis.text.x = element_blank()) +
            labs(x=local(label), y='Collisions', fill=local(label)) +
            ggtitle(paste('Number of Collisions by ', label, ', 2013-2024', sep='')))
  } else {
    print(graph_data %>%
            filter(OccurredYear == local(year)) %>%
            group_by(TargetFeature) %>%
            summarise(Collisions = n()) %>%
            ggplot(aes(x=TargetFeature, y=Collisions, fill=TargetFeature)) +
            geom_bar(stat='identity') +
            theme_set(theme_gray(base_size = text_size)) +
            theme(axis.text.x = element_blank()) +
            labs(x=local(label), y='Collisions', fill=local(label)) +
            ggtitle(paste('Number of Collisions by ', label, ', ', year, sep='')))
  }
  
  # number over time
  print(graph_data %>%
          filter(OccurredDate >= '2013-01-01',
                 OccurredDate < '2025-03-01') %>%
          group_by(OccurredDate = floor_date(OccurredDate, unit='months'), TargetFeature) %>%
          summarise(Collisions = n()) %>%
          ggplot(aes(x=OccurredDate, y=Collisions, color=TargetFeature)) +
          geom_line() +
          labs(x='Date of Collisions', y='Collisions', color=local(label)) +
          theme_set(theme_gray(base_size = text_size)) +
          ggtitle(paste('Number of Collisions over Time by', label)))
  
  # rate over time
  print(graph_data %>%
          filter(OccurredDate >= '2013-01-01',
                 OccurredDate < '2025-03-01') %>%
          group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
          mutate(TotalCollisions = n()) %>% ungroup() %>%
          group_by(OccurredDate, TargetFeature) %>%
          summarise(Proportion = n()/first(TotalCollisions)) %>%
          ggplot(aes(x=OccurredDate, y=Proportion, color=TargetFeature)) +
          geom_line() +
          labs(x='Date of Collisions', y='Proportion of Collisions', color=local(label)) + 
          scale_y_continuous(labels = scales::percent) +
          theme_set(theme_gray(base_size = text_size)) +
          ggtitle(paste('Proportion of Collisions over Time by', label)))
}

mapping_hotspots <- function(graph_data, field_name, label=field_name,
                             year='all', high_percentile=95, text_size=5, dot_size=1) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))
  
  if (year != 'all') {
    graph_data <- graph_data %>% filter(OccurredYear == local(year))
    collision_threshold <- 10
    year_label <- year
  } else {
    graph_data <- graph_data %>% filter(OccurredDate >= '2013-01-01', 
                                        OccurredDate < '2025-01-01')
    collision_threshold <- 50
    year_label <- '2013-2024'
  }
  
  grouped_data <- graph_data %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      TargetFeatureRate = round(mean(TargetFeature), 2),
      TargetFeatureCount = sum(TargetFeature)) %>%
    filter(Collisions > local(collision_threshold))

  feature_percentiles <- grouped_data %>%
    group_by(1) %>%
    mutate(CountNTile = ntile(TargetFeatureCount, 100),
           RateNTile = ntile(TargetFeatureRate, 100)) %>%
    ungroup()
  
  count_percentile <- (feature_percentiles %>% 
                        filter(CountNTile == local(high_percentile)) %>% 
                        summarise(Threshold = min(TargetFeatureCount)))$Threshold[1]
  rate_percentile <- (feature_percentiles %>% 
                        filter(RateNTile == local(high_percentile)) %>% 
                        summarise(Threshold = min(TargetFeatureRate)))$Threshold[1]

  # histogram of collisions by intersection
  print(
    grouped_data %>%
      arrange(desc(TargetFeatureCount)) %>%
      ggplot(aes(x=TargetFeatureCount)) +
      geom_histogram(bins=25) +
      geom_vline(xintercept=local(count_percentile), linetype='dashed') +
      labs(x=paste('Collisions Involving ', label, sep=''), y='Number of Intersections') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Number of Collisions Involving ', label, ' by Intersection, ', year_label, sep=''),
              subtitle=paste(high_percentile, 'th Percentile: ', count_percentile, ' Collisions', sep=''))
  )
  
  # histogram of rates by intersection
  print(
    grouped_data %>%
      arrange(desc(TargetFeatureRate)) %>%
      ggplot(aes(x=TargetFeatureRate)) +
      geom_histogram(bins=25) +
      geom_vline(xintercept=local(rate_percentile), linetype='dashed') +
      scale_x_continuous(labels=scales::percent) +
      labs(x=paste('Rate of Collisions Involving ', label, sep=''), y='Number of Intersections') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Rate of Collisions Involving ', label, ' by Intersection, ', year_label, sep=''),
              subtitle=paste(high_percentile, 'th Percentile: ', rate_percentile*100, '%', sep=''))
  )

  # mapping hotspots
  print(
    grouped_data %>%
      mutate(
        Hotspot = factor(case_when(
          TargetFeatureCount >= local(count_percentile) & 
            TargetFeatureRate >= local(rate_percentile) ~ 'High Rate and Volume',
          TargetFeatureCount >= local(count_percentile) ~ 'High Volume',
          TargetFeatureRate >= local(rate_percentile) ~ 'High Rate',
          TRUE ~ 'Not a Hotspot'
        ), levels=c('High Rate and Volume', 'High Volume', 'High Rate', 'Not a Hotspot'))
      ) %>%
      arrange(desc(Hotspot)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Hotspot)) +
      geom_point(size=dot_size) +
      labs(x='', y='', color='') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Map of Hotspots for Collisions Involving ', label, ', ', year_label, sep=''))
  )
}

graphing_injury_rates <- function(graph_data, field_name, label1=field_name,
                                  label2=label1, year=2024, text_size=5) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))

  if (year != 'all') {
    graph_data <- graph_data %>% filter(OccurredYear == local(year))
    year_label <- year
  } else {
    graph_data <- graph_data %>% filter(OccurredDate >= '2013-01-01', 
                                        OccurredDate < '2025-01-01')
    year_label <- '2013-2024'
  }
  
  ## proportion of hit and run severity by injury outcome
  print(
    graph_data %>%
      group_by(TargetFeature) %>%
      mutate(TotalCollisions = n()) %>% ungroup() %>%
      group_by(TargetFeature, MostSevereInjuryReported) %>%
      summarise(Proportion = n()/first(TotalCollisions)) %>%
      ggplot(aes(x=TargetFeature, y=Proportion, fill=TargetFeature)) +
      geom_bar(stat='identity', position='dodge') +
      labs(x='', y='Proportion of Collisions', fill=label1) +
      scale_y_continuous(labels = scales::percent) +
      facet_wrap(~MostSevereInjuryReported) +
      theme_set(theme_gray(base_size = text_size)) +
      theme(axis.text.x = element_blank()) +
      ggtitle(paste('Distribution of Injury Outcomes by ', label1, ', ', year_label, sep=''))
  )
  
  ## proportion of injury outcome by hit and run severity
  print(
    graph_data %>%
      group_by(MostSevereInjuryReported) %>%
      mutate(TotalCollisions = n()) %>% ungroup() %>%
      group_by(TargetFeature, MostSevereInjuryReported) %>%
      summarise(Proportion = n()/first(TotalCollisions)) %>%
      ggplot(aes(x=TargetFeature, y=Proportion, fill=TargetFeature)) +
      geom_bar(stat='identity', position = position_dodge2(width = 0.9, preserve = "single")) +
      labs(x='', y='Proportion of Collisions', fill=label1) +
      scale_y_continuous(labels = scales::percent) +
      facet_wrap(~MostSevereInjuryReported) +
      theme_set(theme_gray(base_size = text_size)) +
      theme(axis.text.x = element_blank()) +
      ggtitle(paste('Distribution of ', label2, ' Causes by Injury Severity, ', year_label, sep=''))
  )
}

print_checkpoint_runtime(Sys.time())
```

# Tracking Number of MOCodes over Time
```{r}
mo_code_descriptions <- read.csv('Data/MOCodes.csv') %>% 
  mutate(Description = substring(MOCode, 6),
         MOCode = as.numeric(substring(MOCode, 1, 4)))

num_mo_codes <- collisions_cleaned %>% 
  mutate(MOCode = str_replace(MOCodes, ' ', ', ')) %>%
  separate_rows(MOCode, convert=TRUE) 
  group_by(DRKey) %>% summarise(NumMOCodes = n_distinct(MOCode))
```

```{r}
collision_mo_codes_labeled <- collision_mo_codes %>% 
  filter(!is.na(MOCode)) %>%
  left_join(
    collision_mo_codes %>% group_by(MOCode) %>%
      summarise(TotalCollisions = n()) %>%
      arrange(desc(TotalCollisions)) %>%
      mutate(Rank=row_number()) %>% select(-TotalCollisions), by='MOCode') %>%
  left_join(mo_code_descriptions, by='MOCode')

collision_mo_codes_pre_post_covid <- collision_mo_codes_labeled %>%
  filter(OccurredDate < '2020-01-01' | OccurredDate >= '2021-01-01') %>%
  mutate(TimePeriod = ifelse(OccurredDate < '2020-01-01', 'Pre-COVID', 'Post-COVID'),
         PreCOVIDCollision = ifelse(TimePeriod == 'Pre-COVID', DRKey, NA),
         PostCOVIDCollision = ifelse(TimePeriod == 'Post-COVID', DRKey, NA)) %>%
  group_by(1) %>%
  mutate(
    TotalPreCOVIDCollisions = n_distinct(PreCOVIDCollision),
    TotalPostCOVIDCollisions = n_distinct(PostCOVIDCollision)
  ) %>% ungroup() %>%
  mutate(Description = ifelse(Rank > 20, 'Other', Description)) %>%
  group_by(Description) %>%
  summarise(
    PreCOVIDCollisions = n_distinct(PreCOVIDCollision),
    PostCOVIDCollisions = n_distinct(PostCOVIDCollision),
    ProportionPreCOVID = PreCOVIDCollisions/first(TotalPreCOVIDCollisions),
    ProportionPostCOVID = PostCOVIDCollisions/first(TotalPostCOVIDCollisions),
    ChangeInProportion = ProportionPostCOVID - ProportionPreCOVID
  )

print_checkpoint_runtime(Sys.time())
```

## Proportion of collisions w/ conflicting MO codes over time
```{r}
check_multi_substring <- function(patterns, text) {
  count <- 0
  
  for (pattern in unique(patterns)) {
    count <- count + as.numeric(grepl(as.character(pattern), text))
  }
  
  return(as.numeric(count > 1))
}

collisions_mo_code_conflicts <- collisions_cleaned %>% 
  mutate(
    ConflictingIntersection = check_multi_substring(c(3037, 3036), MOCodes),
    ConflictingCityProperty = check_multi_substring(c(3035, 3034), MOCodes),
    ConflictingPrivateProperty = check_multi_substring(c(3033, 3032), MOCodes),
    ConflictingDUI = check_multi_substring(c(3039, 3038), MOCodes),
    ConflictingHitAndRun = check_multi_substring(c(3030, 3029), MOCodes),
    ConflictingInjury = check_multi_substring(c(3028, 3027, 3026, 3025, 3024), MOCodes),
    ConflictingRegion = check_multi_substring(c(4027, 4026, 4025, 4024), MOCodes),
    ConflictingPCF = check_multi_substring(c(3104, 3103, 3102), MOCodes),
    ConflictingLocation = check_multi_substring(c(4001:4021), MOCodes),
    ConflictingMOCodes = ConflictingIntersection == 1 |
      ConflictingCityProperty == 1 | ConflictingPrivateProperty == 1 |
      ConflictingDUI == 1 | ConflictingHitAndRun == 1 | 
      ConflictingInjury == 1 | ConflictingRegion == 1 |
      ConflictingPCF == 1 | ConflictingLocation == 1
  )

print_checkpoint_runtime(Sys.time())
```

# Data for File 5. Demographic Trends
# Census data
```{r}
census_data <- read.csv('Cleaned Data/LACensusDataOverall.csv')
```

# Commutes over Time
## Commuting shifts after COVID
```{r}
commuting_data_overall <- read.csv('Cleaned Data/LACommutingDataOverall.csv')
commuting_data_by_ethnicity <- read.csv('Cleaned Data/LACommutingDataByEthnicity.csv')

commuting_by_covid_period <- commuting_data_by_ethnicity %>%
  filter(Year >= 2016, Year != 2020) %>%
  mutate(Period = ifelse(Year < 2020, 'Pre-COVID', 'Post-COVID') %>%
           factor(levels = c('Pre-COVID', 'Post-COVID'))) %>%
  group_by(Period, RaceEthnicity) %>%
  summarise(
    AvgWorkersPerYear = mean(TotalWorkers) %>% round(),
    AvgCommutingByCar = mean(Driving + Carpooling) %>% round(),
    AvgCommutingByPublicTransit = mean(PublicTransit) %>% round(),
    AvgCommutingByWalking = mean(Walking) %>% round(),
    AvgCommutingByTaxiBikeAndOther = mean(TaxiBikeAndOther) %>% round(),
    AvgWorkFromHome = mean(WorkingFromHome) %>% round(),
    
    CommutingByCarRate = round(AvgCommutingByCar/AvgWorkersPerYear, 2),
    CommutingByPublicTransitRate = round(AvgCommutingByPublicTransit/AvgWorkersPerYear, 2),
    CommutingByWalkingRate = round(AvgCommutingByWalking/AvgWorkersPerYear, 2),
    CommutingByTaxiBikeAndOtherRate = round(AvgCommutingByTaxiBikeAndOther/AvgWorkersPerYear, 2),
    WFHRate = round(AvgWorkFromHome/AvgWorkersPerYear, 2),
  )

print_checkpoint_runtime(Sys.time())
```

## Comparing number of collision victims by ethnicity, compared to WFH rate
```{r}
collisions_by_ethnicity <- collisions_cleaned %>%
  mutate(
    VictimEthnicity = case_when(
      VictimDescent == 'H' ~ 'Hispanic',
      VictimDescent %in% c('A', 'C', 'D', 'F', 'G', 'J', 'K', 'L', 'V', 'Z') ~ 'Asian',
      VictimDescent %in% c('P', 'S', 'U') ~ 'Pacific Islander',
      VictimDescent == 'B' ~ 'Black',
      VictimDescent == 'W' ~ 'White',
      VictimDescent %in% c('O', 'X', '', '-') ~ 'Other',
      VictimDescent == 'I' ~ 'Native American/Alaskan',
      TRUE ~ '?'    )
  ) %>%
  group_by(OccurredYear, VictimEthnicity) %>%
  summarise(Collisions = n()) %>%
  right_join(commuting_data_by_ethnicity, 
             by=c('VictimEthnicity'='RaceEthnicity', 'OccurredYear'='Year')) %>%
  right_join(census_data %>% filter(DemographicType == 'RaceEthnicity'),
             by=c('VictimEthnicity'='Demographic', 'OccurredYear'='Year')) %>%
  mutate(
    Car = Driving + Carpooling,
    DrivingProportion = Car/TotalWorkers,
    WalkingProportion = Walking/TotalWorkers,
    PublicTransitProportion = PublicTransit/TotalWorkers,
    WFHProportion = WorkingFromHome/TotalWorkers,
    CollisionRate = Collisions/Population)

print_checkpoint_runtime(Sys.time())
```

# Parsing ZIP Codes
```{r}
all_coordinates <- read.csv('Cleaned Data/LACoordinatesByZIPCode.csv')

relevant_zip_codes <- (all_coordinates %>% group_by(ZIPCode) %>% 
  summarise(NumCollisionPoints = sum(IsCollisionPoint)) %>%
  filter(NumCollisionPoints > 0) %>% ungroup() %>%
  select(ZIPCode))$ZIPCode

conflicted_coordinates <- all_coordinates %>%
  filter(CoordinateType == 'Collision',
         Backfilled != Frontfilled) %>%
  group_by(Backfilled, Frontfilled) %>% 
  summarise(1)

print_checkpoint_runtime(Sys.time())
```

# Breaking Down Demographics by ZIP Code
```{r}
census_by_zip <- read.csv('Cleaned Data/LACensusDataByZIP.csv')
commuting_by_zip <- read.csv('Cleaned Data/LACommutingDataByZIP.csv')
```

```{r}
census_by_zip_pivoted <- census_by_zip %>%
  mutate(LabelGroup = case_when(
    LabelGroup1 == 'RACE' & LabelGroup2 == 'ONE RACE' ~ case_when(
      LabelGroup3 == 'BLACK OR AFRICAN AMERICAN' ~ 'BLACK',
      LabelGroup3 == 'AMERICAN INDIAN AND ALASKA NATIVE' ~ 'INDIGENOUS',
      LabelGroup3 == 'NATIVE HAWAIIAN AND OTHER PACIFIC ISLANDER' ~ 'PACIFIC ISLANDER',
      LabelGroup3 == 'SOME OTHER RACE' ~ 'OTHER',
      TRUE ~ LabelGroup3),
    LabelGroup1 == 'HISPANIC OR LATINO AND RACE' & 
      LabelGroup2 == 'NOT HISPANIC OR LATINO' ~ case_when(
        LabelGroup3 == 'BLACK OR AFRICAN AMERICAN ALONE' ~ 'BLACK ALONE',
        LabelGroup3 == 'AMERICAN INDIAN AND ALASKA NATIVE ALONE' ~ 'INDIGENOUS ALONE',
        LabelGroup3 == 'NATIVE HAWAIIAN AND OTHER PACIFIC ISLANDER ALONE' ~ 'PACIFIC ISLANDER ALONE',
        LabelGroup3 == 'SOME OTHER RACE ALONE' ~ 'OTHER ALONE',
        LabelGroup3 == 'TWO OR MORE RACES' ~ 'TWO OR MORE RACES ALONE',
        TRUE ~ LabelGroup3),
    LabelGroup1 == 'HISPANIC OR LATINO AND RACE' ~ 'HISPANIC',
    TRUE ~ LabelGroup2) %>% str_to_title() %>% 
        str_replace_all(' To ', ' to ') %>% str_replace_all(' ', '')
  ) %>% select(Year, ZIP, LabelGroup, Population) %>% 
  # filter(LabelGroup != 'TwoOrMoreRaces', LabelGroup != 'TwoOrMoreRacesAlone') %>%
  group_by(Year, ZIP, LabelGroup) %>% summarise(Population = first(Population)) %>% ungroup() %>%
  pivot_wider(names_from=LabelGroup, values_from=Population) %>%
  mutate(TotalPopulation = Male + Female) %>%
  select(Year, ZIP, TotalPopulation, Female, Male, Hispanic, White, Asian, Black, 
         Indigenous, PacificIslander, TwoOrMoreRaces, Other, WhiteAlone, 
         AsianAlone, BlackAlone, IndigenousAlone, PacificIslanderAlone,
         TwoOrMoreRacesAlone, OtherAlone, Under5Years, `5to9Years`, everything())

print_checkpoint_runtime(Sys.time())
```

```{r}
commuting_by_zip_pivoted <- commuting_by_zip %>%
  mutate(LabelGroup = case_when(
    LabelGroup3 == 'DROVE ALONE' ~ 'Driving',
    LabelGroup3 == 'CARPOOLED' ~ 'Carpooling',
    LabelGroup2 %in% c('PUBLIC TRANSPORTATION (EXCLUDING TAXICAB)',
                       'PUBLIC TRANSPORTATION') ~ 'PublicTransit',
    LabelGroup2 == 'WALKED' ~ 'Walking',
    LabelGroup2 %in% c('WORKED AT HOME', 'WORKED FROM HOME') ~ 'WorkingFromHome',
    LabelGroup2 == 'BICYCLE' ~ 'Biking',
    LabelGroup2 %in% c('MOTORCYCLE', 'TAXICAB', 
                       'TAXI OR RIDE-HAILING SERVICES', 'OTHER MEANS') ~ 'TaxiAndOther'
  )) %>% 
  filter(!is.na(LabelGroup)) %>%
  select(Year, ZIP, LabelGroup, Population) %>%
  group_by(Year, ZIP, LabelGroup) %>% summarise(Population = sum(Population)) %>% ungroup() %>%
  pivot_wider(names_from=LabelGroup, values_from=Population)

print_checkpoint_runtime(Sys.time())
```

```{r}
set.seed(55)

zip_codes_all_metrics <- collisions_cleaned %>%
  filter(OccurredYear >= 2013, OccurredYear <= 2024) %>%
  left_join(
    all_coordinates %>%
      filter(CoordinateType == 'Collision') %>%
      rowwise() %>%
      mutate(ZIPCode = ifelse(Backfilled != Frontfilled, 
                              sample(c(Backfilled, Frontfilled), 1), Backfilled)) %>%
      select(Latitude, Longitude, ZIPCode), by=c('Latitude', 'Longitude')) %>%
  left_join(census_by_zip_pivoted %>% mutate(ZIPCode = as.numeric(ZIP)) %>%
              rename(OccurredYear = Year) %>% select(-ZIP), 
            by=c('ZIPCode', 'OccurredYear')) %>%
  left_join(commuting_by_zip_pivoted %>% mutate(ZIPCode = as.numeric(ZIP)) %>%
              rename(OccurredYear = Year) %>% select(-ZIP), 
            by=c('ZIPCode', 'OccurredYear')) %>%
  ungroup()

print_checkpoint_runtime(Sys.time())
```

## Comparing the census and commuter data to collisions data
```{r}
population_rates_graph <- function(pop_df, population='', collision_type='', 
                                   textlabel1 = 'Population Proportion', 
                                   textlabel2 = 'Collision Proportion', 
                                   graph_title = paste(population, textlabel1, 'vs', 
                                                       collision_type, textlabel2, sep=' '),
                                   victim_threshold = 10, legend_position = 'none') {
 
  print(
    pop_df %>% 
      filter(OccurredYear >= 2016, OccurredYear <= 2024, OccurredYear != 2020) %>% 
      rename(
        TargetFeature1 = as.symbol(local(population)),
        TargetFeature2 = as.symbol(local(collision_type))
      ) %>%
      group_by(ZIPCode, OccurredYear) %>%
      summarise(
        TargetFeature1 = first(TargetFeature1),
        TotalPopulation = first(TotalPopulation),
        
        TargetFeature2 = sum(TargetFeature2, na.rm=TRUE),
        TotalCollisions = n()
      ) %>%
      filter(TargetFeature2 >= local(victim_threshold)) %>%
      mutate(TimePeriod = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID')) %>%
      group_by(ZIPCode, TimePeriod) %>%
      summarise(
        TargetFeature1 = mean(TargetFeature1, na.rm=TRUE),
        TotalPopulation = mean(TotalPopulation),
        TargetFeature2 = mean(TargetFeature2),
        TotalCollisions = mean(TotalCollisions, na.rm=TRUE),
      ) %>%
      mutate(
        PopulationProportion = TargetFeature1/TotalPopulation,
        VictimRate = TargetFeature2/TotalCollisions
      ) %>%
      ggplot(aes(x=PopulationProportion, y=VictimRate, 
                 color=as.factor(TimePeriod))) +
      geom_point() +
      labs(x=local(textlabel1), y=local(textlabel2), 
           color='Time Period') +
      theme(legend.position=local(legend_position)) +
      scale_y_continuous(labels=scales::percent, limits = c(0, NA)) +
      scale_x_continuous(labels=scales::percent, limits = c(0, NA)) +
      ggtitle(local(graph_title), subtitle='By ZIP Code in LA, Pre-COVID vs Post-COVID')
  )  
}

print_checkpoint_runtime(Sys.time())
```

```{r}
collision_volume_graphs <- function(pop_df, population = '', collision_type = '', 
                                    textlabel1 = 'Population', textlabel2 = 'Collision Volume',
                                    graph_title = paste(population, textlabel1, 'vs', 
                                                        collision_type, textlabel2, sep=' '),
                                    victim_threshold = 10, legend_position = 'none',
                                    bounded_timeframe = TRUE, include_2020 = FALSE) {
  print(
    pop_df %>% 
      filter( (OccurredYear >= 2016 & OccurredYear <= 2024) |
               local(!bounded_timeframe),
             OccurredYear != 2020 | local(include_2020)) %>% 
      rename(
        TargetFeature1 = as.symbol(local(population)),
        TargetFeature2 = as.symbol(local(collision_type))
      ) %>%
      group_by(ZIPCode, OccurredYear) %>%
      summarise(
        TargetFeature1 = first(TargetFeature1),
        TargetFeature2 = sum(TargetFeature2, na.rm=TRUE),
      ) %>%
      filter(TargetFeature2 >= local(victim_threshold)) %>%
      mutate(TimePeriod = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID')) %>%
      group_by(ZIPCode, TimePeriod) %>%
      summarise(
        PopulationVolume = mean(TargetFeature1, na.rm=TRUE),
        CollisionVolume = sum(TargetFeature2),
      ) %>%
      ggplot(aes(x=PopulationVolume, y=CollisionVolume, 
                 color=as.factor(TimePeriod))) +
      geom_point() +
      labs(x=local(textlabel1), y=local(textlabel2), 
           color='Time Period') +
      theme(legend.position=local(legend_position)) +
      ggtitle(local(graph_title), subtitle='By ZIP Code in LA')
  )  
}
```

```{r}
# for runtime logging
file_1_end <- Sys.time()

print(paste('Total runtime: ', format(file_1_end - file_1_start)))
```
