---
title: "COVID Cases Forecasting"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Load and Clean Data
```{r}
collisions_df <- read.csv('Data/Traffic_Collision_Data_from_2010_to_Present.csv')

collisions_cleaned <- collisions_df %>%
  rename(DRKey = `DR.Number`,
         ReportedDate = `Date.Reported`,
         OccurredDate = `Date.Occurred`,
         OccurredTime = `Time.Occurred`,
         AreaKey = `Area.ID`,
         AreaName = `Area.Name`,
         ReportingDistrict = `Reporting.District`,
         CrimeCode = `Crime.Code`,
         CrimeCodeDescription = `Crime.Code.Description`,
         MOCodes = `MO.Codes`,
         VictimAge = `Victim.Age`,
         VictimSex = `Victim.Sex`,
         VictimDescent = `Victim.Descent`,
         PremiseCode = `Premise.Code`,
         PremiseDescription = `Premise.Description`,
         CrossStreet = `Cross.Street`
         ) %>%
  mutate(
    ReportedDate = as.Date(ReportedDate, format='%m/%d/%Y'),
    OccurredDate = as.Date(OccurredDate, format='%m/%d/%Y'),
    Address = str_squish(Address),
    CrossStreet = str_squish(CrossStreet),
    AreaName = str_squish(AreaName),
    CrimeCodeDescription = str_squish(CrimeCodeDescription),
    MOCodes = str_squish(MOCodes)
  ) %>% rowwise() %>% mutate(
    Latitude = as.numeric(str_split(gsub('\\)', '', gsub('\\(', '', Location)), ', ')[[1]][[1]]),
    Longitude = as.numeric(str_split(gsub('\\)', '', gsub('\\(', '', Location)), ', ')[[1]][[2]]),
  ) %>% select(-Location)
```

```{r}
collisions_cleaned
```

# Breakdown of Fields
## Categorical Fields
Histograms of OccurredTime, AreaKey/AreaName, ReportingDistrict, CrimeCode/CrimeCodeDescription, VictimAge, VictimSex, VictimDescent, PremiseCode/PremiseDescription, Address, CrossStreet; need to make sure I'm correctly tracking what is or isn't a collision
```{r, fig.height=5, fig.width=7}
for (col in c('OccurredTime', 'AreaKey', 'VictimAge')) {
  print(
    collisions_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram() +
      ggtitle(paste0(col, ' Distribution'))
  )
}

for (col in c('AreaName', 'VictimSex', 'VictimDescent')) {
  print(
    collisions_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram(stat='count') +
      theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust=1)) +
      ggtitle(paste0(col, ' Distribution'))
  )
}

# CrimeCode and CrimeCodeDescription are all the same - Traffic Collision

# No clear concentration of collisions
print(collisions_cleaned %>% group_by(ReportingDistrict) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

# Most common by far is STREET (101)
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

# Each PremiseCode only has one PremiseDescription
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% ungroup() %>% group_by(PremiseCode) %>% 
       summarise(Count=n()) %>% arrange(desc(Count)))

# PremiseDecription for 418 is empty, which comes up once; should replace with NA or vice versa
# Otherwise, each PremiseDescription only has one PremiseCode
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% ungroup() %>% group_by(PremiseDescription) %>% 
       summarise(Count=n()) %>% arrange(desc(Count)))
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% filter(PremiseDescription == ''))

# No clear concentration of collisions
print(collisions_cleaned %>% group_by(Address, CrossStreet) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

```

## Time series graphs
```{r}
collisions_by_reported <- collisions_cleaned %>% 
  group_by(ReportedDate) %>% summarise(Collisions = n())
collisions_by_occurred <- collisions_cleaned %>% 
  group_by(OccurredDate) %>% summarise(Collisions = n())

collisions_by_reported_by_week <- collisions_cleaned %>% 
  group_by(ReportedDate = floor_date(ReportedDate, unit='weeks')) %>% 
  summarise(Collisions = n())
collisions_by_occurred_by_week <- collisions_cleaned %>% 
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(collisions_by_reported %>%ggplot(aes(x=ReportedDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Reported Date'))
print(collisions_by_occurred %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Occurred Date'))

print(collisions_by_reported_by_week %>%ggplot(aes(x=ReportedDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Reported Date'))
print(collisions_by_occurred_by_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

### Looking at it by week and by area
```{r}
collisions_by_week <- collisions_cleaned %>% 
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())

collisions_by_area_and_week <- collisions_cleaned %>% 
  group_by(AreaName, OccurredDate = floor_date(OccurredDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(collisions_by_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(collisions_by_area_and_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```
### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- collisions_cleaned %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
print(occurred_time_stats_over_time %>% ggplot(aes(x=OccurredDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=OccurredDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```
### Occurred time distribution
```{r}
occurred_time_dist_over_time <- collisions_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(OccurredDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())

print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

```{r}
collisions_by_coordinates_and_year %>% 
  filter(OccurredYear == local(year),
         Latitude <= local(34.4),
         Latitude >= local(33.6),
         Longitude <= local(-118.125),
         Longitude >= local(-118.75),
         )
```

## Scatterplot of Latitude, Longitude, 
```{r, fig.width=6, fig.height=3.5}
rounding_factor <- 100

collisions_by_coordinates_and_year <- collisions_cleaned %>% 
  group_by(Address, CrossStreet, Latitude, Longitude, OccurredYear = year(OccurredDate)) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Latitude = floor(Latitude*local(rounding_factor))/local(rounding_factor), 
           Longitude = floor(Longitude*local(rounding_factor))/local(rounding_factor),
           OccurredYear) %>% 
  summarise(TotalCollisions = sum(Collisions),
            AvgCollisionsPerLocation = mean(Collisions, na.rm=TRUE),
            NumCollisionRecords = n()) %>%
  filter(Latitude != 0 | Longitude != 0)

for (year in c(2011:2025)) {
  print(
    collisions_by_coordinates_and_year %>%
      arrange(TotalCollisions) %>% 
      filter(OccurredYear == local(year),
             Latitude <= local(34.4),
             Latitude >= local(33.6),
             Longitude <= local(-118.125),
             Longitude >= local(-118.75),
             ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=TotalCollisions)) +
      geom_point(size=2) +
      ggtitle(paste0('Total Collisions by Coordinates in ', year, sep=''))
  )
}
```

```{r, fig.width=6, fig.height=3.5}
for (year in c(2011:2025)) {
  print(
    collisions_by_coordinates_and_year %>%
      arrange(AvgCollisionsPerLocation) %>%
      filter(OccurredYear == local(year),
             Latitude <= local(34.4),
             Latitude >= local(33.6),
             Longitude <= local(-118.125),
             Longitude >= local(-118.75),
             ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=AvgCollisionsPerLocation)) +
      geom_point(size=2) +
      ggtitle(paste0('Avg Collisions Per Location by Coordinates in ', year, sep=''))
  )
}
```

# Forecasting Overall Collision Counts
## Pre-COVID
```{r}
collisions_over_time_pre_covid <- collisions_cleaned %>% 
  filter(OccurredDate < as.Date('2020-03-15')) %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(collisions_over_time_pre_covid$OccurredDate)), 
                month(min(collisions_over_time_pre_covid$OccurredDate)))

collisions_pre_covid_ts <- ts(collisions_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(collisions_pre_covid_ts, round(length(collisions_pre_covid_ts) * 0.8))
h <- length(collisions_pre_covid_ts) - length(train)
test <- tail(collisions_pre_covid_ts, h)
```


```{r}
collisions_pre_covid_auto_model <- auto.arima(train)
collisions_pre_covid_auto_model
```

### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

### AR
```{r}
pacf(train_dff1) # 3-4
```
### MA
```{r}
acf(train_dff1) #2
```
### ARIMA
```{r}
collisions_pre_covid_model <- Arima(train, order=c(3, 1, 2))
collisions_pre_covid_model
```

### Comparing AIC
```{r}
exp((collisions_pre_covid_auto_model$aic - collisions_pre_covid_model$aic)/2)
```
### Forecast
```{r}
forecast_1 <- forecast(collisions_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(collisions_pre_covid_model, h=length(test))

autoplot(collisions_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

## Post-COVID
```{r}
collisions_over_time_post_covid <- collisions_cleaned %>% 
  filter(OccurredDate >= as.Date('2021-01-03')) %>%
  group_by(OccurredDate = floor_date(OccurredDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(collisions_over_time_post_covid$OccurredDate)), 
                month(min(collisions_over_time_post_covid$OccurredDate)))

collisions_post_covid_ts <- ts(collisions_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(collisions_post_covid_ts, round(length(collisions_post_covid_ts) * 0.8))
h <- length(collisions_post_covid_ts) - length(train)
test <- tail(collisions_post_covid_ts, h)
```


```{r}
collisions_post_covid_auto_model <- auto.arima(train)
collisions_post_covid_auto_model
```

### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

### AR
```{r}
pacf(train) # 1-2
```
### MA
```{r}
acf(train) #3
```
### ARIMA
```{r}
collisions_post_covid_model <- Arima(train, order=c(1, 0, 3))
collisions_post_covid_model
```

### Comparing AIC
```{r}
exp((collisions_post_covid_auto_model$aic - collisions_post_covid_model$aic)/2)
```
### Forecast
```{r}
forecast_1 <- forecast(collisions_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(collisions_post_covid_model, h=length(test))

autoplot(collisions_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```






# Identifying Collision Hot Spots
```{r}
rounding_factor <- 100

collisions_by_location_2024 <- collisions_cleaned %>%
  mutate(OccurredYear = year(OccurredDate)) %>%
  filter(OccurredYear == 2024) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Latitude = round(Latitude*local(rounding_factor))/local(rounding_factor), 
           Longitude = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>% 
  summarise(TotalCollisions = sum(Collisions),
            AvgCollisionsPerLocation = mean(Collisions, na.rm=TRUE),
            NumCollisionRecords = n(),
            RangeOfCollisions = max(Collisions) - min(Collisions),
            StdCollisions = sd(Collisions)
            ) %>%
  filter(Latitude != 0 | Longitude != 0)
```
## Check the spread of collision statistics
```{r}
print(
  collisions_by_location_2024 %>%
    ggplot(aes(x=NumCollisionRecords)) +
    geom_histogram()
)
```

```{r}
print(
  collisions_by_location_2024 %>% 
    ggplot(aes(x=RangeOfCollisions, y=StdCollisions, size=NumCollisionRecords)) +
    geom_point()
)
```


```{r}
collisions_by_location_2024 %>% arrange(desc(RangeOfCollisions))
```

```{r}
collisions_by_location_2024 %>% arrange(desc(StdCollisions))
```

```{r}
collisions_cleaned %>%
  mutate(OccurredYear = year(OccurredDate),
         RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
         RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor),
         ) %>%
  filter(
    OccurredYear == 2024,
    (RoundedLat == 33.99 & RoundedLon == -118.28) |
    (RoundedLat == 34.22 & RoundedLon == -118.45) |
    (RoundedLat == 34.03 & RoundedLon == -118.52) |
    (RoundedLat == 34.04 & RoundedLon == -118.56) 
  ) %>%
  group_by(Address, CrossStreet, Latitude, Longitude, RoundedLat, RoundedLon) %>% 
  summarise(Collisions = n()) %>% arrange(desc(RoundedLat), desc(RoundedLon))
```
### Checking for intersections with multiple sets of coordinates, which I should've done before
```{r}
collisions_cleaned %>% 
  filter(Latitude != 0 | Longitude != 0) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Address, CrossStreet) %>% 
  summarise(
    NumCoordinates = n(),
    MaxLat = max(Latitude),
    MinLat = min(Latitude),
    MaxLon = max(Longitude),
    MinLon = min(Longitude),
    LatRange = MaxLat - MinLat,
    LonRange = MaxLon - MinLon,
    LatStd = sd(Latitude),
    LonStd = sd(Longitude),
    AvgLat = mean(Latitude),
    AvgLon = mean(Longitude),
    ModeLat = mode(Latitude),
    ModeLon = mode(Longitude)
  ) %>%
  arrange(desc(NumCoordinates))
```
```{r}
print(collisions_cleaned %>% filter(Address == 'PACIFIC COAST HY', CrossStreet == 'FIGUEROA ST') %>%
        select(Address, CrossStreet, Latitude, Longitude, PremiseDescription) %>%
        group_by(Latitude, Longitude) %>% summarise(Count = n()) %>%
        ggplot(aes(x=Longitude, y=Latitude, size=Count)) +
        geom_point())
```

```{r}
collisions_cleaned %>% filter(Latitude == 33.9892, Longitude ==	-118.2827)
```

```{r}
collision_mo_codes <- collisions_cleaned %>% 
  mutate(MOCode = str_replace(MOCodes, ' ', ', ')) %>%
  separate_rows(MOCode, convert=TRUE)
```

## Classifying the MO Codes, based on the PDF I found
```{r}
mo_code_descriptions <- read.csv('Data/MOCodes.csv') %>% 
  mutate(Description = substring(MOCode, 6),
         MOCode = as.numeric(substring(MOCode, 1, 4)))

mo_code_descriptions %>% write.csv('Cleaned Data/MOCodeDescriptions.csv', row.names=FALSE)

collision_mo_codes %>% 
  filter(!is.na(MOCode)) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

# Select the MOCodes of interest
collision_mo_codes %>% 
  filter(MOCode %in% c(3004, 3006, 3011, 3003, 3008, 3012,
                       3037, 3036, 3030, 3029, 3028, 3026, 3025, 3035, 3034, 3602, 3039, 3027, 3038, 1218
                       )) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

# 3004, 3006, 3011, 3003, 3008, 3012
# 3037 vs 3036, 3030/3029, 3028 vs 3026/3025, 3035 vs 3034, 3602, 3039, 3027, 3038, 1218
```

```{r}
collision_mo_codes %>% 
  filter(MOCode %in% c(0947, 3025,			
3101, 3401, 3701, 3006, 3030, 3032,			
0605, 3101, 3401, 3701, 3011, 3030, 3036,		
0605, 3101, 3401, 3701, 3006, 3030,			
3101, 3401, 3701, 3006, 3034, 3032,			
3101, 3401, 3701, 3006, 3030, 0605,			
3028, 4011, 3030, 3034, 0605, 3101, 3401, 3701, 3011, 3104)) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

collisions_with_mo_labels %>% 
  filter(CrossStreet == '') %>%
  group_by(Latitude, Longitude) %>%
  summarise(AtIntersection = max(AtIntersection),
            PrivatePropertyInvolved = max(PrivatePropertyInvolved)) %>%
  filter(AtIntersection == 1, PrivatePropertyInvolved == 1) %>%
  select(Latitude, Longitude)

collisions_with_mo_labels %>%
  filter(Latitude == 34.1367	, Longitude == -118.2713) %>%
  select(AtIntersection, PrivatePropertyInvolved, CrossStreet)
```

```{r}
# checking for MO Code overlap
## at intersection y/n - rare overlap seems to be more at intersections
collisions_cleaned %>% filter(grepl('3037', MOCodes), grepl('3036', MOCodes))

## collision w/; rare overlap, should probably classify as multiple and also make individual columns
collisions_cleaned %>% filter(grepl('3004', MOCodes) + grepl('3006', MOCodes) + grepl('3011', MOCodes) + 
                              grepl('3003', MOCodes) + grepl('3008', MOCodes) + grepl('3012', MOCodes) > 1)

## hit and run misd vs felony; rare overlap, should probably classify as felony
collisions_cleaned %>% filter(grepl('3030', MOCodes), grepl('3029', MOCodes))

## injury y/n; rare overlap; should probably classify as complaint of injury
collisions_cleaned %>% filter(grepl('3028', MOCodes),  grepl('3026', MOCodes))
# also 3024, 3027

## city property involved y/n; rare overlap; should probably classify as y
collisions_cleaned %>% filter(grepl('3035', MOCodes), grepl('3034', MOCodes))
# misc - 3025, 3602, 3027, 3038, 1218

collisions_with_mo_labels <- collisions_cleaned %>%
  mutate(
    AtIntersection = grepl('3036', MOCodes),
    NotAtIntersection = grepl('3037', MOCodes),
    MovementPrecedingCollisions = grepl('3701', MOCodes),
    TypeOfCollision = grepl('3401', MOCodes),
    CollisionWithVehicle = grepl('3004', MOCodes),
    CollisionWithParkedVehicle = grepl('3006', MOCodes),
    CollisionWithFixedObject = grepl('3011', MOCodes),
    CollisionWithPedestrian = grepl('3003', MOCodes),
    CollisionWithBike = grepl('3008', MOCodes),
    CollisionWithOtherObject = grepl('3012', MOCodes),
    HitAndRun = grepl('3030', MOCodes) | grepl('3029', MOCodes),
    HitAndRunMisdemeanor = grepl('3030', MOCodes),
    HitAndRunFelony = grepl('3029', MOCodes),
    InjuryInvolved = grepl('3026', MOCodes) | grepl('3025', MOCodes) |
      grepl('3024', MOCodes) | grepl('3027', MOCodes),
    NonInjury = grepl('3028', MOCodes),
    ComplaintOfInjury = grepl('3026', MOCodes),
    VisibleInjury = grepl('3025', MOCodes),
    SevereInjury = grepl('3024', MOCodes),
    FatalInjury = grepl('3027', MOCodes),
    CityPropertyInvolved = grepl('3034', MOCodes),
    CityPropertyNotInvolved = grepl('3035', MOCodes),
    PrivatePropertyInvolved = grepl('3032', MOCodes),
    PrivatePropertyNotInvolved = grepl('3033', MOCodes),
    UnlicensedMotorist = grepl('3602', MOCodes),
    DUIMisdemeanor = grepl('3039', MOCodes),
    DUIFelony = grepl('3038', MOCodes),
    Sobriety = grepl('3801', MOCodes),
    HomelessVictim = grepl('1218', MOCodes)
  )

collisions_with_mo_labels %>% 
  write.csv('Cleaned Data/CollisionsWithMOLabels.csv',
            row.names=FALSE)
```


# Cleaning Up Some More
```{r}
collisions_with_mo_labels
```

```{r}
# Check that DRKey values are unique
collisions_with_mo_labels %>% nrow()
collisions_with_mo_labels %>% distinct(DRKey) %>% nrow()

# Range of ReportedDate and OccurredDate
collisions_with_mo_labels %>% summarise(min(ReportedDate), max(ReportedDate))
collisions_with_mo_labels %>% summarise(min(OccurredDate), max(OccurredDate))

# Distribution of OccurredTime
print(collisions_with_mo_labels %>% 
  group_by(OccurredTime = round(OccurredTime, -2)) %>% 
  summarise(Count = n()) %>%
  ggplot(aes(x=OccurredTime, y=Count)) +
  geom_bar(stat='identity'))

# Oddities w/ AreaName
  collisions_with_mo_labels %>% group_by(AreaName) %>% 
    summarise(Count = n()) %>% arrange(AreaName)
  
# Oddities w/ VictimSex
  collisions_with_mo_labels %>% group_by(VictimSex) %>% 
    summarise(Count = n()) %>% arrange(VictimSex)
  
# Oddities w/ VictimDescent
  collisions_with_mo_labels %>% group_by(VictimDescent) %>% 
    summarise(Count = n()) %>% arrange(VictimDescent)
  
# Oddities w/ PremiseDescription
  collisions_with_mo_labels %>% group_by(PremiseDescription) %>% 
    summarise(Count = n()) %>% arrange(PremiseDescription)
  
# Oddities w/ Address
  collisions_with_mo_labels %>% group_by(Address) %>% 
    summarise(Count = n()) %>% arrange(Address)
  
# Oddities w/ CrossStreet
  collisions_with_mo_labels %>% group_by(CrossStreet) %>% 
    summarise(Count = n()) %>% arrange(CrossStreet)
```

## Cleaning up the Address and CrossStreet values
This is a mess and I don't wanna do it
```{r}
street_names <- collisions_with_mo_labels %>% select(Address) %>%
  rbind(collisions_with_mo_labels %>% mutate(Address = CrossStreet) %>% 
          select(Address)) %>% distinct() %>% arrange(Address)

all_street_name_parts <- data.frame()

for (i in 1:nrow(street_names)) {
  street_name_parts <- str_split(street_names[i, 'Address'], ' ')[[1]]
  n_street_name_parts <- length(street_name_parts)
  new_street_name_part <- data.frame('Address' = street_names[i, 'Address'],
                                     'NumStreetNameParts' = n_street_name_parts)
  
  for (j in 1:8) {
    if (n_street_name_parts >= j) {
      new_street_name_part[paste('StreetNamePart', j, sep='')] = street_name_parts[j] 
    } else {
      new_street_name_part[paste('StreetNamePart', j, sep='')] = NA 
    }
  }
  new_street_name_part['LastStreetNamePart'] = street_name_parts %>% tail(n=1)
  all_street_name_parts <- rbind(all_street_name_parts, new_street_name_part)
}

```

```{r}
isolated_street_name_parts <- all_street_name_parts %>% 
  mutate(StreetNamePart = StreetNamePart1) %>% 
  select(Address, StreetNamePart) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart2) %>% 
      select(Address, StreetNamePart)
  ) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart3) %>% 
      select(Address, StreetNamePart)
  ) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart4) %>% 
      select(Address, StreetNamePart)
  ) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart5) %>% 
      select(Address, StreetNamePart)
  ) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart6) %>% 
      select(Address, StreetNamePart)
  ) %>%
  rbind(
    all_street_name_parts %>% 
      mutate(StreetNamePart = StreetNamePart7) %>% 
      select(Address, StreetNamePart)
  )

isolated_street_name_parts %>% 
  filter(!is.na(StreetNamePart)) %>%
  group_by(StreetNamePart) %>% 
  summarise(Count = n(), ExampleAddress = last(Address)) %>%
  arrange(desc(Count)) %>%
  filter(is.na(as.numeric(StreetNamePart)))

# Most street names that end in a number need that number
# Try removing initial numbers when that isn't the whole street name
# Then, remove cardinal directions from the beginning, street type from the end
# Then, convert "AVENUE", "AVUNUE", and "AVENEUE" to "AVE"
# Lastly, look again for weirdness and iterate

all_street_name_parts %>% 
  filter(!is.na(as.numeric(LastStreetNamePart)))

all_street_name_parts %>% 
  filter(is.na(as.numeric(LastStreetNamePart))) %>%
  group_by(LastStreetNamePart) %>% 
  summarise(Count = n(), ExampleAddress = last(Address)) %>%
  arrange(desc(Count)) %>%
  filter(is.na(as.numeric(LastStreetNamePart)))
```

```{r}
collisions_with_mo_labels %>% filter(
  grepl('WESTWOOD PA', Address)
)
```

```{r}
street_name_cleanup <- function(street_name) {
  return(
    street_name %>% str_to_upper() %>% 
      str_replace(' PLAZA', '') %>%
      str_replace(' PLAZ', '') %>%
      str_replace(' PL', '') %>%
      str_replace(' BL', '') %>%
      str_replace(' AVE', '') %>%
      str_replace(' AV', '') %>%
      str_replace('S ', '') %>%
      str_replace('N ', '') %>%
      str_replace('W ', '') %>%
      str_replace('E ', '') %>%
      str_replace('00 ', '') %>%
      str_replace(' RD', '') %>%
      str_replace(' ST', '') %>%
      str_replace(' PA', '') %>%
      str_replace(' CT', '') %>%
      str_replace(' CI', '') %>%
      str_replace(' LN', '') %>%
      str_replace(' WAY', '') %>%
      str_replace(' WY', '') %>%
      str_replace(' FWY', 'FREEWAY') %>%
      str_replace(' FRWY', 'FREEWAY') %>%
      str_replace(' E', '') %>%
      str_replace(' FY', '') %>%
      str_replace(' O', '') %>%
      str_replace(' ONRAMP', '') %>%
      str_replace(' ENTR', '') %>%
      str_replace(' OFF', '') %>%
      str_replace(' OFFRA', '') %>%
      str_replace(' EAST', '') %>%
      str_replace(' EXIT', '') %>%
      str_replace(' R', '') %>%

      str_replace(' DR', '')
  )
}

cleaned_street_names <- street_names %>%
  mutate(CleanedAddress = street_name_cleanup(Address)) %>%
  arrange(CleanedAddress)

cleaned_street_names %>% distinct(CleanedAddress) %>% 
cleaned_street_names %>% filter(CleanedAddress == '100 C')
# Order of operations
# Identify freeways by number, group them
# Keep weird names like C St

str_split('00 17TH AV', ' ')
```

```{r}
collisions_with_mo_labels %>% filter(CrossStreet %>% grepl('CLARENCST'))
```


## Cleaning up the coordinates that are listed as (0, 0)
```{r}
collisions_with_mo_labels %>% filter(Latitude == 0 | Longitude == 0) %>% 
  group_by(Address, CrossStreet) %>%
  summarise(Collisions = n()) %>%
  left_join(
    collisions_with_mo_labels %>% filter(Latitude != 0, Longitude != 0) %>%
      group_by(Address, CrossStreet) %>% 
      summarise(Latitude = mean(Latitude),
                Longitude = mean(Longitude),
                NumExamples = n()),
    by=c('Address', 'CrossStreet')
  )

print(
  collisions_with_mo_labels %>% filter(Latitude == 0 | Longitude == 0) %>% 
    group_by(Address, CrossStreet) %>%
    summarise(Collisions = n()) %>%
    left_join(
      collisions_with_mo_labels %>% filter(Latitude != 0, Longitude != 0) %>%
        group_by(Address, CrossStreet) %>% 
        summarise(Latitude = mean(Latitude),
                  Longitude = mean(Longitude),
                  NumExamples = n()),
      by=c('Address', 'CrossStreet')
    ) %>%
    ggplot(aes(x=Longitude, y=Latitude, size=Collisions)) +
    geom_point()
)
```

```{r}
collisions_with_mo_labels %>% 
  filter(Latitude == 0 | Longitude == 0) %>% 
  group_by(Year = year(OccurredDate)) %>%
  summarise(Count = n()) %>% 
  arrange(Year)
# These are largely from 2012, 2014, 2020, 2021, 2023
```


## Checking that coordinates have consistent AreaNames and PremiseDescriptions
I already know the Addresses and CrossStreets can differ, even beyond formatting; this just further indicates I should ignore them
```{r}
# Not super consistent, establish a hierarchy
collisions_with_mo_labels %>% 
  group_by(Latitude, Longitude, AreaName) %>%
  summarise(1) %>%
  group_by(Latitude, Longitude) %>%
  summarise(NumAreaNames = n_distinct(AreaName),
            AreaNames = paste(AreaName, collapse = '; ')) %>%
  arrange(desc(NumAreaNames))

# Also not super consistent; maybe the coordinates aren't accurate?
collisions_with_mo_labels %>% 
  group_by(Latitude, Longitude, PremiseDescription) %>%
  summarise(1) %>%
  group_by(Latitude, Longitude) %>%
  summarise(NumPremiseDescriptions = n_distinct(PremiseDescription),
            PremiseDescriptions = paste(PremiseDescription, collapse = '; ')) %>%
  arrange(desc(NumPremiseDescriptions))
```

