---
title: "COVID Cases Forecasting"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Shift in Collision Volume Before/After COVID
The COVID-19 pandemic and lockdown seems to have had a lasting effect on the number of traffic collisions occuring in LA. We'll explore in later analysis how this is driven by changes in car commuter volume as more workers began to work from home
```{r, fig.height=5, fig.width=7}
print(
  collisions_cleaned %>% 
    mutate(OccurredDateRounded = floor_date(OccurredDate, unit='weeks')) %>%
    group_by(OccurredDateRounded, COVIDPeriod) %>%
    summarise(Collisions = n(),
              NumDays = n_distinct(OccurredDate)) %>%
    filter(NumDays == 7) %>%
    ggplot(aes(x=OccurredDateRounded, y=Collisions, color=COVIDPeriod)) +
    geom_line() +
    labs(x='Week of Occurred Date', y='Number of Collisions', color='') +
    ggtitle('Influence of COVID Lockdown on Number of Collisions per Week')
)
```
## By area
```{r, fig.height=8, fig.width=10}
print(
  collisions_cleaned %>% 
    mutate(OccurredDateRounded = floor_date(OccurredDate, unit='months')) %>%
    group_by(OccurredDateRounded, COVIDPeriod, AreaName) %>%
    summarise(Collisions = n(),
              NumDays = n_distinct(OccurredDate)) %>%
    ggplot(aes(x=OccurredDateRounded, y=Collisions, color=COVIDPeriod)) +
    geom_line() +
    facet_wrap(~AreaName) +
    labs(x='Month of Occurred Date', y='Number of Collisions', color='') +
    ggtitle('Influence of COVID Lockdown on Number of Collisions by AreaName')
)
```


# OccurredTime - Time of Day of Collision
## Pre-COVID vs Post-COVID
```{r, fig.height=5, fig.width=7}
print(
  collisions_cleaned %>%
    filter(OccurredYear >= 2016) %>%
    group_by(COVIDPeriod) %>% mutate(TotalCollisions = n()) %>% ungroup() %>%
    group_by(OccurredTime = floor(OccurredTime/100)*100, COVIDPeriod) %>%
    summarise(CollisionProportion = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=OccurredTime, y=CollisionProportion, color=COVIDPeriod)) +
    geom_line() + ylim(0, NA) +
    labs(x='Collision Occurrence Time', y='Distribution of Collisions', color='') +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
    ggtitle('Distribution of Collision Occurrence Times')
)
```

# Demographic Data
## Age
```{r, fig.height=5, fig.width=7}
print(
  collisions_cleaned %>%
    filter(OccurredYear >= 2016) %>%
    group_by(COVIDPeriod) %>% mutate(TotalCollisions = n()) %>% ungroup() %>%
    group_by(VictimAge = floor(VictimAge/5)*5, COVIDPeriod) %>%
    summarise(CollisionProportion = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=VictimAge, y=CollisionProportion, color=COVIDPeriod)) +
    geom_line() + ylim(0, NA) +
    labs(x='Victim Age', y='Distribution of Collisions', color='') +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
    ggtitle('Distribution of Collision Victim Age')
)
```

## Sex
```{r, fig.height=5, fig.width=7}
print(
  collisions_cleaned %>% 
    filter(OccurredYear > 2016) %>%
    mutate(VictimSex = case_when(
      VictimSex %in% c('M', 'F') ~ VictimSex,
      TRUE ~ 'Other or Unknown'
    )) %>%
    group_by(COVIDPeriod) %>% mutate(TotalCollisions=n()) %>% ungroup() %>%
    group_by(VictimSex, COVIDPeriod) %>%
    summarise(CollisionProportion = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=VictimSex, y=CollisionProportion, fill=COVIDPeriod)) +
    geom_bar(stat='identity', position='dodge') +
    labs(x='Collision Victim Sex', y='Proportion of Collisions', fill='') +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Victim Sex')
)
```

## Race and Ethnicity
```{r, fig.height=5, fig.width=7}
print(
  collisions_cleaned %>% 
    filter(OccurredYear > 2016) %>%
    group_by(COVIDPeriod) %>% mutate(TotalCollisions=n()) %>% ungroup() %>%
    group_by(VictimEthnicity, COVIDPeriod) %>%
    summarise(CollisionProportion = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=VictimEthnicity, y=CollisionProportion, fill=COVIDPeriod)) +
    geom_bar(stat='identity', position='dodge') +
    labs(x='Collision Victim Race/Ethnicity', y='Proportion of Collisions', fill='') +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Victim Race/Ethnicity')
)
```

# By Location
## Scatterplot of Latitude, Longitude, 
```{r, fig.height=5, fig.width=7}
print(
  collisions_by_coordinates_and_year %>%
    arrange(TotalCollisions) %>% 
    filter(Latitude <= local(34.4),
           Latitude >= local(33.6),
           Longitude <= local(-118.125),
           Longitude >= local(-118.75),
           ) %>%
    ggplot(aes(x=Longitude, y=Latitude, color=TotalCollisions)) +
    geom_point(size=0.5) +
    facet_wrap(~OccurredYear) +
    ggtitle(paste0('Total Collisions by Coordinates in ', year, sep=''))
)
```

```{r, fig.width=6, fig.height=3.5}
for (year in c(2011:2025)) {
  print(
    collisions_by_coordinates_and_year %>%
      arrange(StanDevCollisions) %>%
      filter(OccurredYear == local(year),
             Latitude <= local(34.4),
             Latitude >= local(33.6),
             Longitude <= local(-118.125),
             Longitude >= local(-118.75),
             ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=StanDevCollisions)) +
      geom_point(size=2) +
      ggtitle(paste0('Avg Collisions Per Location by Coordinates in ', year, sep=''))
  )
}
```

## Check the spread of collision statistics
```{r}
print(
  collisions_by_location_2024 %>%
    ggplot(aes(x=NumCollisionRecords)) +
    geom_histogram()
)
```

```{r}
print(
  collisions_by_location_2024 %>% 
    ggplot(aes(x=RangeOfCollisions, y=StdCollisions, size=NumCollisionRecords)) +
    geom_point()
)
```

```{r}
collisions_by_location_2024 %>% arrange(desc(RangeOfCollisions))
```

```{r}
collisions_by_location_2024 %>% arrange(desc(StdCollisions))
```

```{r}
collisions_cleaned %>%
  mutate(OccurredYear = year(OccurredDate),
         RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
         RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor),
         ) %>%
  filter(
    OccurredYear == 2024,
    (RoundedLat == 33.99 & RoundedLon == -118.28) |
    (RoundedLat == 34.22 & RoundedLon == -118.45) |
    (RoundedLat == 34.03 & RoundedLon == -118.52) |
    (RoundedLat == 34.04 & RoundedLon == -118.56) 
  ) %>%
  group_by(Address, CrossStreet, Latitude, Longitude, RoundedLat, RoundedLon) %>% 
  summarise(Collisions = n()) %>% arrange(desc(RoundedLat), desc(RoundedLon))
```


