---
title: "Traffic Collisions MO Code Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Load Data
```{r}
collisions_with_injury_type <- read.csv('Cleaned Data/CollisionsWithMOLabels.csv') %>%
  mutate(
    ReportedDate = as.Date(ReportedDate, format='%Y-%m-%d'),
    OccurredDate = as.Date(OccurredDate, format='%Y-%m-%d'),
    MostSevereInjuryReported = case_when(
      FatalInjury ~ 'Fatal Injury',
      SevereInjury ~ 'Severe Injury',
      VisibleInjury ~ 'Visible Injury',
      ComplaintOfInjury ~ 'Complaint of Injury',
      NonInjury ~ 'No Injury',
      TRUE ~ 'Unspecified'
    )
  )

collisions_with_injury_type <- collisions_with_injury_type %>% mutate(
    CollisionAtIntersection = case_when(
      AtIntersection ~ 'At Intersection',
      NotAtIntersection ~ 'Not At Intersection',
      TRUE ~ 'Unspecified'
    ))
```

# Comparing Intersections
## First round of analysis - preliminary, no real leads here
```{r}
# histogram
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    ggplot(aes(x=CollisionAtIntersection, fill=CollisionAtIntersection)) +
    geom_histogram(stat='count') +
    ggtitle('Number of Collisions at Intersection in 2024')
)

# scatterplots
rounding_factor <- 1000
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>%
    summarise(Collisions = n(),
              ProportionAtIntersection = mean(AtIntersection)) %>%
    ggplot(aes(x=ProportionAtIntersection, y=Collisions)) +
    geom_point() +
    ggtitle('Number of Collisions at Locations by Proportion at Intersection, 2024')
)

rounding_factor <- 1000
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor),
             CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=Collisions, fill=CollisionAtIntersection)) +
    geom_histogram(stat='count') +
    facet_wrap(~CollisionAtIntersection) +
    ggtitle('Histogram of Number of Collisions at Intersection in 2024')
)

# map
rounding_factor <- 100
print(
  collisions_with_injury_type %>%
    filter(year(OccurredDate) == 2024,
           Latitude != 0, Longitude != 0) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>%
    summarise(Collisions = n(),
              ProportionAtIntersection = mean(AtIntersection)) %>%
    arrange(desc(Collisions)) %>%
    ggplot(aes(x=RoundedLon, y=RoundedLat, color=ProportionAtIntersection,
               size=Collisions)) +
    geom_point() +
    ggtitle('Number of Collisions at Location by Proportion At Intersection, 2024')
)

# bar chart of injury severity by intersection vs non-intersection
print(
  collisions_with_injury_type %>%
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude, 3), 
             RoundedLon = round(Longitude, 3)) %>%
    mutate(NumCollisionsAtIntersection = sum(as.numeric(AtIntersection)),
           NumCollisionsNotAtIntersection = sum(as.numeric(NotAtIntersection))) %>%
    ungroup() %>%
    filter(NumCollisionsAtIntersection >= 5 |
           NumCollisionsNotAtIntersection >= 1) %>%
    group_by(CollisionAtIntersection, MostSevereInjuryReported) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=MostSevereInjuryReported, y=Collisions, fill=CollisionAtIntersection)) +
    geom_bar(stat='identity') +
    facet_wrap(~CollisionAtIntersection) +
    ggtitle('Number of Collision Injuries by Severity, Intersection vs Non-Intersection, 2024')
)
```

## Second round - more directed, hopefully more fruitful
```{r}
# looking to see whether intersections in general are more "dangerous" than non-intersections
# in a given area with both intersection collisions and non-intersection collisions, which happen more frequently
## i.e. what proportion of the collisions in these areas occurs at intersections

# also, in those areas, are injuries more likely/severe at intersections or not

# first, can we identify the right coordinates per intersection
## Seems to be too much variance in what coordinates an address x cross street combo can be listed as, maybe by mistake
## Should group by coordinates, and pick the nearest intersection by most common listing; round to nearest 1000th place
collisions_with_injury_type %>% 
  filter(AtIntersection) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Count = 1) %>%
  group_by(Address, CrossStreet) %>% 
  summarise(Count=n()) %>%
  arrange(desc(Count))

# 33.942, -118.4095 is a problematic set of coordinates; it's LAX and encompasses a ton of nearby locations
# There are some other coordinates that seem wrong
collisions_with_injury_type %>%
  mutate(NearestIntersection = paste0(Address, ' and ', CrossStreet)) %>%
  group_by(RoundedLat = round(Latitude, 3),
           RoundedLon = round(Longitude, 3),
           NearestIntersection) %>%
  summarise(Count = n()) %>%
  group_by(RoundedLat, RoundedLon) %>%
  summarise(Count = n()) %>% arrange(desc(Count))

# trying again to round out the intersection areas by name and coordinates
collisions_at_intersection_areas <- collisions_with_injury_type %>%
  filter(Latitude != 0, Longitude != 0) %>%
  mutate(
    HasCrossStreet = as.numeric(!is.na(CrossStreet) & !(CrossStreet == '')),
    NearestIntersection = ifelse(
      HasCrossStreet,  paste0(Address, ' and ', CrossStreet), Address)) %>%
  group_by(
    HasCrossStreet, NearestIntersection,
    RoundedLat = round(Latitude, 4),
    RoundedLon = round(Longitude, 4)) %>%
  summarise(Count = n(), 
            ProportionListedAsAtIntersection = mean(AtIntersection),
            ProportionListedAsNotAtIntersection = mean(NotAtIntersection),
            NumCollisionsAtIntersection = sum(AtIntersection),
            NumCollisionsNotAtIntersection = sum(NotAtIntersection),
            ) %>%
  arrange(desc(Count)) %>%
  group_by(HasCrossStreet, RoundedLat, RoundedLon) %>%
  summarise(
    Collisions = sum(Count),
    NumIntersections = n(),
    NumCollisionsAtIntersection = sum(NumCollisionsAtIntersection),
    NumCollisionsNotAtIntersection = sum(NumCollisionsNotAtIntersection),
    ProportionListedAsAtIntersection = NumCollisionsAtIntersection/Collisions,
    ProportionListedAsNotAtIntersection = NumCollisionsNotAtIntersection/Collisions,
    NearestIntersections = paste(NearestIntersection, collapse=', ')
  ) %>% 
  mutate(
    ListedAsIntersection = as.numeric(ProportionListedAsAtIntersection > 0),
    ListedAsNotIntersection = as.numeric(ProportionListedAsNotAtIntersection > 0),
    ListedMoreAsAtIntersection = (ProportionListedAsAtIntersection > 
                                    ProportionListedAsNotAtIntersection)
  )
  
# besides a couple exceptions, it seems any location with a cross street and any collisions being listed 
# at intersection should be labeled an intersection, 

# comparing the possible outcomes
collisions_at_intersection_areas %>%
  group_by(ListedAsIntersection, ListedAsNotIntersection,
           ListedMoreAsAtIntersection, HasCrossStreet) %>% 
  summarise(Collisions = sum(Collisions))

collisions_at_intersection_areas %>% 
  filter(ListedAsIntersection == 1,
         ListedAsNotIntersection == 1,
         ListedMoreAsAtIntersection == 0,
         HasCrossStreet == 0) %>% 
  arrange(desc(Collisions))

```

```{r}
# tracking the intersection labeling over time
print(
  collisions_with_injury_type %>% 
    filter(OccurredDate >= '2013-01-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='weeks'),
             CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, color=CollisionAtIntersection)) +
    geom_line()
)

# tracking the intersection labeling by location over time
print(
  collisions_with_injury_type %>% 
    filter(OccurredDate >= '2013-01-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='weeks'),
             RoundedLat = round(Latitude, 4),
             RoundedLon = round(Longitude, 4)) %>%
    summarise(
      AtIntersection = max(AtIntersection),
      NotAtIntersection = max(NotAtIntersection),
    ) %>% 
    mutate(
      labeledAsIntersection = case_when(
        AtIntersection == 1 & NotAtIntersection == 1 ~ 'Both',
        AtIntersection == 1 ~ 'At Intersection',
        NotAtIntersection == 1 ~ 'Not At Intersection',
        TRUE ~ 'Unspecified'
      )) %>% 
    group_by(OccurredWeek, labeledAsIntersection) %>%
    summarise(NumLocations = n()) %>%
    ggplot(aes(x=OccurredWeek, y=NumLocations, color=labeledAsIntersection)) +
    geom_line()
)
```

```{r}
collisions_with_injury_type %>% 
  filter(round(Latitude, 3) == 33.989, round(Longitude, 3) == -118.309) %>%
  select(ReportingDistrict, AtIntersection, NotAtIntersection)

collisions_with_injury_type %>% distinct(Address)

collision_mo_codes %>% 
  filter(Latitude == 33.9470, Longitude == -118.2684) %>% 
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% 
  summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))
```
## Trying this again, without street names
```{r}
identifying_intersections <- collisions_with_injury_type %>%
  filter(Latitude != 0, Longitude != 0) %>%
  mutate(
    HasCrossStreet = as.numeric(!is.na(CrossStreet) & 
                                  !(CrossStreet == ''))) %>%
  group_by(
    RoundedLat = round(Latitude, 4),
    RoundedLon = round(Longitude, 4)) %>%
  summarise(Collisions = n(), 
            NumCollisionsWithCrossStreet = sum(HasCrossStreet),
            NumCollisionsAtIntersection = sum(AtIntersection),
            NumCollisionsNotAtIntersection = sum(NotAtIntersection),
            ProportionListedWithCrossStreet = mean(HasCrossStreet),
            ProportionListedAsAtIntersection = mean(AtIntersection),
            ProportionListedAsNotAtIntersection = mean(NotAtIntersection),
            HasCrossStreet = max(HasCrossStreet),
            ListedAsIntersection = max(AtIntersection),
            ListedAsNotIntersection = max(NotAtIntersection),
            ListedMoreAsAtIntersection = (ProportionListedAsAtIntersection >
                                            ProportionListedAsNotAtIntersection)
    ) %>% ungroup()

identifying_intersections %>% 
  filter(HasCrossStreet == 0, ListedAsIntersection == 1) %>%
  arrange(desc(Collisions))

identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0) %>%
  arrange(desc(Collisions))

identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0) %>%
  arrange(desc(NumCollisionsNotAtIntersection))

# Basically, if the coordinates are ever listed with a CrossStreet, they are at an intersection
```

```{r}
collisions_with_labeled_intersections <- collisions_with_injury_type %>%
  # filter(Latitude != 0, Longitude != 0) %>%
  select(-CollisionAtIntersection) %>%
  left_join(
    identifying_intersections %>% 
      mutate(CollisionAtIntersection = ifelse(HasCrossStreet %>% is.na(), 0, HasCrossStreet), 
             Latitude = RoundedLat, Longitude = RoundedLon) %>%
      select(Latitude, Longitude, CollisionAtIntersection, ListedAsIntersection),
    by = c('Latitude', 'Longitude')
  ) %>%
  mutate(
    CollisionAtIntersection = ifelse(is.na(CollisionAtIntersection), 
                                     as.numeric(CrossStreet != ''), 
                                     CollisionAtIntersection),
    OccurredYear = year(OccurredDate)
  )

```

## Finally examining the intersections vs non-intersections
```{r}
print(
  collisions_with_labeled_intersections %>% 
    filter(OccurredDate >= '2013-01-01',
           OccurredDate < '2025-03-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='months'),
             CollisionAtIntersection = ifelse(CollisionAtIntersection == 1, 
                                              'At Intersection', 'Not at Intersection')) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, 
               color=as.factor(CollisionAtIntersection))) +
    geom_line() +
    guides(
      colour = guide_legend(
        title = "Collision Location"
        #, theme = theme(legend.title = element_text(hjust = 0))
      )
    )
)
```

```{r}
print(
  collisions_with_labeled_intersections %>% 
    filter(OccurredDate >= '2013-01-01',
           OccurredDate < '2025-03-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='months'),
             CollisionAtIntersection = 
               case_when(CollisionAtIntersection == 1 & ListedAsIntersection == 1 ~ 'At Intersection', 
                         CollisionAtIntersection == 1 ~ 'Near Intersection',
                         TRUE ~ 'Not at Intersection')) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, 
               color=as.factor(CollisionAtIntersection))) +
    geom_line() +
    guides(
      colour = guide_legend(
        title = "Collision Location"
        #, theme = theme(legend.title = element_text(hjust = 0))
      )
    )
)
```

## Double-checking that I've re-labeled intersections correctly
```{r}
identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0, ListedAsNotIntersection == 1) %>%
  arrange(desc(NumCollisionsNotAtIntersection))
# Far as I can tell, the vast majority of coordinates listed with CrossStreets point to intersections
```

```{r}
# Check the MO Codes of one of these re-labeled intersections that looks iffy
collision_mo_codes %>% 
  filter(!is.na(MOCode), Latitude == 34.1589, Longitude ==	-118.4994) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))
```

## Back to looking into comparing intersections vs non-intersections
Seeing how often intersections have crashes in the vicinity that aren't at intersections - basically none
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(year(OccurredDate) == 2024,
           Latitude != 0, Longitude != 0) %>%
    group_by(RoundedLat = round(Latitude, 3), 
             RoundedLon = round(Longitude, 3)) %>%
    summarise(NumCollisionsAtIntersection = sum(as.numeric(CrossStreet != '')),
              NumCollisionsNotAtIntersection = sum(as.numeric(CrossStreet == '')),
              ProportionAtIntersection = round(NumCollisionsAtIntersection/
                (NumCollisionsAtIntersection + NumCollisionsNotAtIntersection), 2)) %>%
    ggplot(aes(x=ProportionAtIntersection)) +
    geom_histogram(stat='count')
)
```

Comparing the severity of crashes at intersections vs non-intersections
FINALLY getting somewhere
Maybe needs error bars
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(OccurredYear == 2024) %>%
    mutate(WasCollisionAtIntersection = ifelse(CollisionAtIntersection == 1, 
                                               'At Intersection', 'Not at Intersection')) %>%
    group_by(WasCollisionAtIntersection) %>%
    mutate(TotalCollisions = n()) %>%
    group_by(WasCollisionAtIntersection, 
             MostSevereInjuryReported = factor(
               MostSevereInjuryReported, levels = c('No Injury', 'Complaint of Injury', 'Visible Injury',
                                                    'Severe Injury', 'Fatal Injury', 'Unspecified'))) %>%
    summarise(InjuryRate = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=MostSevereInjuryReported, y=InjuryRate, fill=WasCollisionAtIntersection)) +
    geom_bar(stat='identity') +
    facet_wrap(~WasCollisionAtIntersection) +
    theme(axis.text.x = element_text(angle = 70, vjust = 1, hjust=1))
)
```
### Looking for dangerous intersections
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(OccurredYear == 2024,
           CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>% 
    mutate(TotalCollisions = n()) %>%
    ungroup() %>%
    filter(TotalCollisions > 5) %>%
    group_by(Latitude, Longitude, 
             MostSevereInjuryReported = factor(MostSevereInjuryReported, 
                                               levels = c('No Injury', 'Complaint of Injury', 'Visible Injury',
                                                          'Severe Injury', 'Fatal Injury', 'Unspecified'))) %>%
    summarise(InjuryRate = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=InjuryRate, fill=MostSevereInjuryReported)) +
    geom_histogram(binwidth = 0.1) +
    facet_wrap(~MostSevereInjuryReported)
)
```

```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(
      # OccurredYear == 2024,
      CollisionAtIntersection == 1,
      Latitude != 0, Longitude != 0) %>%
    group_by(
      RoundedLat = round(Latitude, 4), 
      RoundedLon = round(Longitude, 4)) %>%
    summarise(Collisions = n(), 
              NonInjuries = sum(as.numeric(NonInjury)),
              Injuries = sum(as.numeric(InjuryInvolved))) %>%
    ggplot(aes(x=Collisions, y=Injuries)) +
    geom_point()
)
```
```{r}
collisions_with_labeled_intersections %>%
  filter(
    # OccurredYear == 2024,
    CollisionAtIntersection == 1,
    Latitude != 0, Longitude != 0) %>%
  group_by(Latitude, Longitude) %>%
  summarise(Collisions = n(), 
            Injuries = sum(as.numeric(InjuryInvolved)),
            InjuryRate = mean(as.numeric(InjuryInvolved)),
            VisibleInjuryRate = mean(as.numeric(VisibleInjury)),
            SevereInjuryRate = mean(as.numeric(SevereInjury))
            ) %>%
  filter(Collisions > 50) %>%
  arrange(desc(InjuryRate))
```
## Looking into that bump in non-intersection collisions around COVID?
```{r}

```

# Collision Object
```{r}
# histogram, map of proportion, map of most common, map by type
# injury rates/severity
collisions_by_object <- collisions_with_labeled_intersections %>%
    filter(CollisionWithVehicle) %>%
    mutate(CollisionObject = 'Vehicle') %>% 
  rbind(
    collisions_with_labeled_intersections %>%
      filter(CollisionWithParkedVehicle) %>%
      mutate(CollisionObject = 'Parked Vehicle')
  ) %>% 
  rbind(
    collisions_with_labeled_intersections %>%
      filter(CollisionWithFixedObject) %>%
      mutate(CollisionObject = 'Fixed Object')
  ) %>% 
  rbind(
    collisions_with_labeled_intersections %>%
      filter(CollisionWithPedestrian) %>%
      mutate(CollisionObject = 'Pedestrian')
  ) %>% 
  rbind(
    collisions_with_labeled_intersections %>%
      filter(CollisionWithBike) %>%
      mutate(CollisionObject = 'Bike') 
  ) %>% 
  rbind(
    collisions_with_labeled_intersections %>%
      filter(CollisionWithOtherObject) %>%
      mutate(CollisionObject = 'Other') 
  ) %>% 
  select(CollisionObject, OccurredDate, OccurredYear, Latitude, Longitude, 
         CollisionAtIntersection, InjuryInvolved, MostSevereInjuryReported,
         NonInjury, ComplaintOfInjury, VisibleInjury, SevereInjury, FatalInjury)
```

```{r}
print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    group_by(CollisionObject) %>%
    summarise(InjuryRate = mean(InjuryInvolved, na.rm=TRUE)) %>%
    ggplot(aes(x=CollisionObject, y=InjuryRate, fill=CollisionObject)) +
    geom_bar(stat='identity')
)

print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    mutate(SevereOrFatalInjury = as.numeric(SevereInjury | FatalInjury)) %>%
    group_by(CollisionObject) %>%
    summarise(SevereOrFatalInjuryRate = mean(SevereOrFatalInjury, na.rm=TRUE)) %>%
    ggplot(aes(x=CollisionObject, y=SevereOrFatalInjuryRate, fill=CollisionObject)) +
    geom_bar(stat='identity') +
    scale_y_continuous(labels = scales::percent) +
    labs(x='Collision Object', y='Severe/Fatal Injury Rate') +
    theme(legend.position='none') +
    ggtitle('Severe/Fatal Injury Rate by Collision Object in LA, 2024')
)

print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    group_by(CollisionObject, CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=CollisionObject, y=Collisions, fill=CollisionObject)) +
    geom_bar(stat='identity') +
    facet_wrap(~CollisionAtIntersection)
)
```
## Finding the intersections with the most/highest rate of pedestrian or bike collisions
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisionRate = round(mean(CollisionWithPedestrian), 2)) %>%
    filter(Collisions > 50) %>%
    arrange(desc(PedestrianCollisionRate)) %>%
    ggplot(aes(x=PedestrianCollisionRate)) +
    geom_histogram()
)

print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisions = sum(CollisionWithPedestrian)) %>%
    filter(Collisions > 50) %>%
    arrange(desc(PedestrianCollisions)) %>%
    ggplot(aes(x=PedestrianCollisions)) +
    geom_histogram()
)

```
## Map of pedestrian collision hotspots
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisions = sum(CollisionWithPedestrian),
      PedestrianCollisionRate = mean(CollisionWithPedestrian)
    ) %>%
    filter(Collisions >= 50) %>%
    mutate(
      PedestrianCollisionHotspot = case_when(
        PedestrianCollisions >= 25 & PedestrianCollisionRate >= 0.15 ~ 'High Rate and Volume',
        PedestrianCollisions >= 25 ~ 'High Volume',
        PedestrianCollisionRate >= 0.15 ~ 'High Rate',
        TRUE ~ 'Not Pedestrian Collision Hotspot'
      )
    ) %>%
    group_by(PedestrianCollisionHotspot) %>% mutate(Total = n()) %>% ungroup() %>%
    arrange(desc(Total)) %>%
    ggplot(aes(x=Longitude, y=Latitude, color=PedestrianCollisionHotspot)) +
    geom_point(size=1.7) +
    labs(x='', y='', color='Is Hotspot for\nPedestrian Collisions') +
    ggtitle('Map of Pedestrian Collisions Hotspots, LA')
)
```


# Defining Functions for Future Analysis
```{r}
intro_graphs <- function(graph_data, field_name, label=field_name, year=2024, text_size=5) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))
  
  # histogram
  if (local(year) == 'all') {
    print(graph_data %>%
            filter(OccurredDate >= '2013-01-01',
                   OccurredDate < '2025-01-01') %>%
            group_by(TargetFeature) %>%
            summarise(Collisions = n()) %>%
            ggplot(aes(x=TargetFeature, y=Collisions, fill=TargetFeature)) +
            geom_bar(stat='identity') +
            theme_set(theme_gray(base_size = text_size)) +
            theme(axis.text.x = element_blank()) +
            labs(x=local(label), y='Collisions', fill=local(label)) +
            ggtitle(paste('Number of Collisions by ', label, ', 2013-2024', sep='')))
  } else {
    print(graph_data %>%
            filter(OccurredYear == local(year)) %>%
            group_by(TargetFeature) %>%
            summarise(Collisions = n()) %>%
            ggplot(aes(x=TargetFeature, y=Collisions, fill=TargetFeature)) +
            geom_bar(stat='identity') +
            theme_set(theme_gray(base_size = text_size)) +
            theme(axis.text.x = element_blank()) +
            labs(x=local(label), y='Collisions', fill=local(label)) +
            ggtitle(paste('Number of Collisions by ', label, ', ', year, sep='')))
  }
  
  # number over time
  print(graph_data %>%
          filter(OccurredDate >= '2013-01-01',
                 OccurredDate < '2025-03-01') %>%
          group_by(OccurredDate = floor_date(OccurredDate, unit='months'), TargetFeature) %>%
          summarise(Collisions = n()) %>%
          ggplot(aes(x=OccurredDate, y=Collisions, color=TargetFeature)) +
          geom_line() +
          labs(x='Date of Collisions', y='Collisions', color=local(label)) +
          theme_set(theme_gray(base_size = text_size)) +
          ggtitle(paste('Number of Collisions over Time by', label)))
  
  # rate over time
  print(graph_data %>%
          filter(OccurredDate >= '2013-01-01',
                 OccurredDate < '2025-03-01') %>%
          group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
          mutate(TotalCollisions = n()) %>% ungroup() %>%
          group_by(OccurredDate, TargetFeature) %>%
          summarise(Proportion = n()/first(TotalCollisions)) %>%
          ggplot(aes(x=OccurredDate, y=Proportion, color=TargetFeature)) +
          geom_line() +
          labs(x='Date of Collisions', y='Proportion of Collisions', color=local(label)) + 
          scale_y_continuous(labels = scales::percent) +
          theme_set(theme_gray(base_size = text_size)) +
          ggtitle(paste('Proportion of Collisions over Time by', label)))
}

mapping_hotspots <- function(graph_data, field_name, label=field_name,
                             year='all', high_percentile=95, text_size=5, dot_size=1) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))
  
  if (year != 'all') {
    graph_data <- graph_data %>% filter(OccurredYear == local(year))
    collision_threshold <- 10
    year_label <- year
  } else {
    graph_data <- graph_data %>% filter(OccurredDate >= '2013-01-01', 
                                        OccurredDate < '2025-01-01')
    collision_threshold <- 50
    year_label <- '2013-2024'
  }
  
  grouped_data <- graph_data %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      TargetFeatureRate = round(mean(TargetFeature), 2),
      TargetFeatureCount = sum(TargetFeature)) %>%
    filter(Collisions > local(collision_threshold))

  feature_percentiles <- grouped_data %>%
    group_by(1) %>%
    mutate(CountNTile = ntile(TargetFeatureCount, 100),
           RateNTile = ntile(TargetFeatureRate, 100)) %>%
    ungroup()
  
  count_percentile <- (feature_percentiles %>% 
                        filter(CountNTile == local(high_percentile)) %>% 
                        summarise(Threshold = min(TargetFeatureCount)))$Threshold[1]
  rate_percentile <- (feature_percentiles %>% 
                        filter(RateNTile == local(high_percentile)) %>% 
                        summarise(Threshold = min(TargetFeatureRate)))$Threshold[1]

  # histogram of collisions by intersection
  print(
    grouped_data %>%
      arrange(desc(TargetFeatureCount)) %>%
      ggplot(aes(x=TargetFeatureCount)) +
      geom_histogram(bins=25) +
      geom_vline(xintercept=local(count_percentile), linetype='dashed') +
      labs(x=paste('Collisions Involving ', label, sep=''), y='Number of Intersections') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Number of Collisions Involving ', label, ' by Intersection, ', year_label, sep=''),
              subtitle=paste(high_percentile, 'th Percentile: ', count_percentile, ' Collisions', sep=''))
  )
  
  # histogram of rates by intersection
  print(
    grouped_data %>%
      arrange(desc(TargetFeatureRate)) %>%
      ggplot(aes(x=TargetFeatureRate)) +
      geom_histogram(bins=25) +
      geom_vline(xintercept=local(rate_percentile), linetype='dashed') +
      scale_x_continuous(labels=scales::percent) +
      labs(x=paste('Rate of Collisions Involving ', label, sep=''), y='Number of Intersections') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Rate of Collisions Involving ', label, ' by Intersection, ', year_label, sep=''),
              subtitle=paste(high_percentile, 'th Percentile: ', rate_percentile*100, '%', sep=''))
  )

  # mapping hotspots
  print(
    grouped_data %>%
      mutate(
        Hotspot = factor(case_when(
          TargetFeatureCount >= local(count_percentile) & 
            TargetFeatureRate >= local(rate_percentile) ~ 'High Rate and Volume',
          TargetFeatureCount >= local(count_percentile) ~ 'High Volume',
          TargetFeatureRate >= local(rate_percentile) ~ 'High Rate',
          TRUE ~ 'Not a Hotspot'
        ), levels=c('High Rate and Volume', 'High Volume', 'High Rate', 'Not a Hotspot'))
      ) %>%
      arrange(desc(Hotspot)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Hotspot)) +
      geom_point(size=dot_size) +
      labs(x='', y='', color='') +
      theme_set(theme_gray(base_size = text_size)) +
      ggtitle(paste('Map of Hotspots for Collisions Involving ', label, ', ', year_label, sep=''))
  )
}

graphing_injury_rates <- function(graph_data, field_name, label1=field_name,
                                  label2=label1, year=2024, text_size=5) {
  # rename column
  graph_data <- graph_data %>% rename(TargetFeature = as.symbol(local(field_name)))

  if (year != 'all') {
    graph_data <- graph_data %>% filter(OccurredYear == local(year))
    year_label <- year
  } else {
    graph_data <- graph_data %>% filter(OccurredDate >= '2013-01-01', 
                                        OccurredDate < '2025-01-01')
    year_label <- '2013-2024'
  }
  
  ## proportion of hit and run severity by injury outcome
  print(
    graph_data %>%
      group_by(TargetFeature) %>%
      mutate(TotalCollisions = n()) %>% ungroup() %>%
      group_by(TargetFeature, MostSevereInjuryReported) %>%
      summarise(Proportion = n()/first(TotalCollisions)) %>%
      ggplot(aes(x=TargetFeature, y=Proportion, fill=TargetFeature)) +
      geom_bar(stat='identity', position='dodge') +
      labs(x='', y='Proportion of Collisions', fill=label1) +
      scale_y_continuous(labels = scales::percent) +
      facet_wrap(~MostSevereInjuryReported) +
      theme_set(theme_gray(base_size = text_size)) +
      theme(axis.text.x = element_blank()) +
      ggtitle(paste('Distribution of Injury Outcomes by ', label1, ', ', year_label, sep=''))
  )
  
  ## proportion of injury outcome by hit and run severity
  print(
    graph_data %>%
      group_by(MostSevereInjuryReported) %>%
      mutate(TotalCollisions = n()) %>% ungroup() %>%
      group_by(TargetFeature, MostSevereInjuryReported) %>%
      summarise(Proportion = n()/first(TotalCollisions)) %>%
      ggplot(aes(x=TargetFeature, y=Proportion, fill=TargetFeature)) +
      geom_bar(stat='identity', position = position_dodge2(width = 0.9, preserve = "single")) +
      labs(x='', y='Proportion of Collisions', fill=label1) +
      scale_y_continuous(labels = scales::percent) +
      facet_wrap(~MostSevereInjuryReported) +
      theme_set(theme_gray(base_size = text_size)) +
      theme(axis.text.x = element_blank()) +
      ggtitle(paste('Distribution of ', label2, ' Causes by Injury Severity, ', year_label, sep=''))
  )
}
```

# Injury Type
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'MostSevereInjuryReported', 
             label='Injury Severity', year=2024)
```


## Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'FatalInjury', label='Fatal Injuries')

mapping_hotspots(collisions_with_labeled_intersections, 
                 'SevereInjury', label='Severe Injuries')
```

# Collision Object
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
        mutate(MostVulnerableCollisionObject = factor(case_when(
          CollisionWithPedestrian ~ 'Pedestrian', 
          CollisionWithBike ~ 'Bike',
          CollisionWithOtherObject ~ 'Other Object',
          CollisionWithVehicle ~ 'Vehicle',
          CollisionWithFixedObject ~ 'Fixed Object',
          CollisionWithParkedVehicle ~ 'Parked Vehicle',
          TRUE ~ 'Unknown'),
          levels = c('Pedestrian', 'Bike', 'Vehicle', 'Fixed Object',
                     'Parked Vehicle', 'Other Object')))

intro_graphs(collisions_with_labeled_intersections, 'MostVulnerableCollisionObject', 
             label='Collision Object', year=2024)
```

## Finding the intersections with the most/highest rate of pedestrian/bike collisions
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'CollisionWithPedestrian', label='Collision w/ Pedestrians')

mapping_hotspots(collisions_with_labeled_intersections, 
                 'CollisionWithBike', label='Collision w/ Bikes')
```

## Injuries by collision object
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'MostVulnerableCollisionObject', 
                      label1='Collision Object', label2='Collision Object')

graphing_injury_rates(collisions_with_labeled_intersections, 'CollisionWithPedestrian', 
                      label1='Collision w/ Pedestrians', label2='Collision w/ Pedestrians')
```

# Hit and Runs
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
        mutate(HitAndRunSeverity = factor(case_when(
          HitAndRunFelony ~ 'Hit & Run Felony', 
          HitAndRunMisdemeanor ~ 'Hit & Run Misdemeanor',
          HitAndRun ~ 'Other Hit & Run',
          TRUE ~ 'Not Labelled Hit & Run'),
          levels = c('Hit & Run Felony', 'Hit & Run Misdemeanor',
                     'Other Hit & Run', 'Not Labelled Hit & Run')))

intro_graphs(collisions_with_labeled_intersections, 'HitAndRunSeverity', 
             label='Hit & Run Severity', year=2024)

intro_graphs(collisions_with_labeled_intersections, 'HitAndRun', 
             label='Hit & Run', year=2024)
```


## Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'HitAndRun', label='Hit & Runs')
```

## Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'HitAndRunSeverity', 
                      label1='Hit & Run Severity', label2='Hit & Run')

graphing_injury_rates(collisions_with_labeled_intersections, 'HitAndRun', 
                      label1='Hit & Run', label2='Hit & Run')
```

# DUIs
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
        mutate(DUI = (DUIMisdemeanor | DUIFelony),
               DUISeverity = factor(case_when(
                 DUIFelony ~ 'DUI Felony', 
                 DUIMisdemeanor ~ 'DUI Misdemeanor',
                 Sobriety ~ 'Sobriety',
                 TRUE ~ 'Unspecified'),
                 levels = c('DUI Felony', 'DUI Misdemeanor',
                            'Sobriety', 'Unspecified')))

intro_graphs(collisions_with_labeled_intersections, 'DUISeverity', 
             label='DUI Severity', year=2024)
```


## Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'DUI', label='DUIs')
```

## Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'DUISeverity', 
                      label1='DUI Severity', label2='DUI')

graphing_injury_rates(collisions_with_labeled_intersections, 'DUI', 
                      label1='DUI', label2='DUI')
```

# Private Property Involved
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
        mutate(PrivatePropertyInvolvement = factor(case_when(
                 PrivatePropertyInvolved ~ 'Involved', 
                 PrivatePropertyNotInvolved ~ 'Not Involved',
                 TRUE ~ 'Unspecified'),
                 levels = c('Involved', 'Not Involved', 'Unspecified')))

intro_graphs(collisions_with_labeled_intersections, 'PrivatePropertyInvolvement', 
             label='Private Property Involved', year=2024)
```


## Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'PrivatePropertyInvolved', label='Private Property')
```

## Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'PrivatePropertyInvolved', 
                      label1='Private Property Involvement', label2='Private Property Collision')
```

# Public Property Involved
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
        mutate(CityePropertyInvolvement = factor(case_when(
                 CityPropertyInvolved ~ 'Involved', 
                 CityPropertyNotInvolved ~ 'Not Involved',
                 TRUE ~ 'Unspecified'),
                 levels = c('Involved', 'Not Involved', 'Unspecified')))

intro_graphs(collisions_with_labeled_intersections, 'CityePropertyInvolvement', 
             label='City Property Involvement', year=2024)
```


## Finding the intersections with the most/highest rate of public property involvement
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'CityPropertyInvolved', label='City Property')
```

## Injuries from public property collisions
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'CityePropertyInvolvement', 
                      label1='City Property Involvement', label2='City Property Collision')
```

# Unlicensed Motorists
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'UnlicensedMotorist', 
             label='Unlicenced Motorist', year=2024)
```


## Finding the intersections with the most/highest rate of unlicensed motorists
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'UnlicensedMotorist', label='Unlicenced Motorists')
```

## Injuries from unlicensed motorists
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'UnlicensedMotorist', 
                      label1='Unlicensed Motorist')
```

# Homeless Victim
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'HomelessVictim', 
             label='Homeless Victim', year=2024)
```


## Finding the intersections with the most/highest rate of homeless victimes
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'HomelessVictim', label='Homeless Victims')
```

## Injuries to homeless victims
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'HomelessVictim', 
                      label1='Homeless Victim')
```

# Victim Descent
```{r, fig.height=2, fig.width=4}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
  mutate(
    VictimEthnicity = case_when(
      VictimDescent == 'H' ~ 'Hispanic/Latino',
      VictimDescent %in% c('A', 'C', 'D', 'F', 'G', 'J', 'K', 'L', 'V', 'Z') ~ 'Asian',
      VictimDescent %in% c('P', 'S', 'U') ~ 'Pacific Islander',
      VictimDescent == 'B' ~ 'Black',
      VictimDescent == 'W' ~ 'White',
      VictimDescent %in% c('O', 'X', '', '-') ~ 'Other',
      VictimDescent == 'I' ~ 'Native American/Alaskan',
      TRUE ~ '?'
      ),
    HispanicOrLatinoVictim = (VictimEthnicity == 'Hispanic/Latino')
    )

intro_graphs(collisions_with_labeled_intersections, 'VictimEthnicity', 
             label='Victim Ethnicity', year=2024)

# Descent Code: A - Other Asian B - Black C - Chinese D - Cambodian F - Filipino G - Guamanian H - Hispanic/Latin/Mexican I - American Indian/Alaskan Native J - Japanese K - Korean L - Laotian O - Other P - Pacific Islander S - Samoan U - Hawaiian V - Vietnamese W - White X - Unknown Z - Asian Indian
```


## Finding the intersections with the most/highest rate of Hispanic victims
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'HispanicOrLatinoVictim', label='Hispanic/Latino Victims')
```

## Injuries by Victim Ethnicity
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'VictimEthnicity', 
                      label1='Victim Ethnicity')
```

# Tracking Number of MOCodes over Time
```{r}
collisions_with_labeled_intersections <- collisions_with_labeled_intersections %>%
  # select(-NumMOCodes) %>%
  left_join(collisions_with_labeled_intersections %>% 
              mutate(MOCode = str_replace(MOCodes, ' ', ', ')) %>%
              separate_rows(MOCode, convert=TRUE) %>%
              group_by(DRKey) %>% summarise(NumMOCodes = n_distinct(MOCode)), by='DRKey')
```

```{r, fig.height=2, fig.width=4}
print(collisions_with_labeled_intersections %>%
        filter(OccurredDate >= '2013-01-01',
               OccurredDate < '2025-03-01') %>%
        group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
        summarise(AvgNumMOCodes = mean(NumMOCodes, na.rm=TRUE)) %>%
        ggplot(aes(x=OccurredDate, y=AvgNumMOCodes)) +
        geom_line() +
        labs(x='Date of Collisions', y='Number of MO Codes') +
        theme_set(theme_gray(base_size = 5)) +
        ggtitle('Avg Number of MO Codes over Time'))
```

```{r}
collision_mo_codes_labeled <- collision_mo_codes %>% 
  filter(!is.na(MOCode)) %>%
  left_join(
    collision_mo_codes %>% group_by(MOCode) %>%
      summarise(TotalCollisions = n()) %>%
      arrange(desc(TotalCollisions)) %>%
      mutate(Rank=row_number()) %>% select(-TotalCollisions), by='MOCode') %>%
  left_join(mo_code_descriptions, by='MOCode')

print(
  collision_mo_codes_labeled %>%
    mutate(Description = ifelse(Rank > 10, 'Other', Description)) %>%
    group_by(Description, OccurredDate = floor_date(OccurredDate, unit='months')) %>%
    summarise(TotalCollisions = n_distinct(DRKey)) %>%
    ggplot(aes(x=OccurredDate, y=TotalCollisions, color=Description)) +
    geom_line()
)

collision_mo_codes_pre_post_covid <- collision_mo_codes_labeled %>%
  filter(OccurredDate < '2020-01-01' | OccurredDate >= '2021-01-01') %>%
  mutate(TimePeriod = ifelse(OccurredDate < '2020-01-01', 'Pre-COVID', 'Post-COVID'),
         PreCOVIDCollision = ifelse(TimePeriod == 'Pre-COVID', DRKey, NA),
         PostCOVIDCollision = ifelse(TimePeriod == 'Post-COVID', DRKey, NA)) %>%
  group_by(1) %>%
  mutate(
    TotalPreCOVIDCollisions = n_distinct(PreCOVIDCollision),
    TotalPostCOVIDCollisions = n_distinct(PostCOVIDCollision)
  ) %>% ungroup() %>%
  mutate(Description = ifelse(Rank > 20, 'Other', Description)) %>%
  group_by(Description) %>%
  summarise(
    PreCOVIDCollisions = n_distinct(PreCOVIDCollision),
    PostCOVIDCollisions = n_distinct(PostCOVIDCollision),
    ProportionPreCOVID = PreCOVIDCollisions/first(TotalPreCOVIDCollisions),
    ProportionPostCOVID = PostCOVIDCollisions/first(TotalPostCOVIDCollisions),
    ChangeInProportion = ProportionPostCOVID - ProportionPreCOVID
  )

print(
  collision_mo_codes_pre_post_covid %>%
    ggplot(aes(x=ChangeInProportion, y=Description, fill=Description)) +
    geom_bar(stat='identity') +
    scale_x_continuous(label=scales::percent) +
    theme_set(theme_gray(base_size = 8)) +
    theme(legend.position='none') +
    labs(y='', x='Change in Proportion') +
    ggtitle('Shift in Proportion of Collisions Labelled, Pre- and Post-COVID')
)

```

## Proportion of collisions w/ conflicting MO codes over time
```{r}
collision_mo_codes_labeled %>%
  group_by(MOCode, Description) %>%
  summarise(TotalCollisions = n_distinct(DRKey)) %>%
  arrange(desc(MOCode))

# 3037 vs 3036 - intersection
# 3028 vs 3027, 3026, 3025, 3024 - injury type
# 3035 vs 3034 - city property involved
# 3033 vs 3032 - private property involved
# 3039, 3038 - DUI level
# 3030 vs 3029 - Hit & Run
# 4026, 4027, 4025, 4024 - region
# 4021, 4020, 4019, 4018, 4017 ... 4001 - location
# 3104, 3103, 3102 - PCF = Primary Collision Factor!

check_multi_substring <- function(patterns, text) {
  count <- 0
  
  for (pattern in unique(patterns)) {
    count <- count + as.numeric(grepl(as.character(pattern), text))
  }
  
  return(as.numeric(count > 1))
}

collisions_mo_code_conflicts <- collisions_with_labeled_intersections %>% 
  mutate(
    ConflictingIntersection = check_multi_substring(c(3037, 3036), MOCodes),
    ConflictingCityProperty = check_multi_substring(c(3035, 3034), MOCodes),
    ConflictingPrivateProperty = check_multi_substring(c(3033, 3032), MOCodes),
    ConflictingDUI = check_multi_substring(c(3039, 3038), MOCodes),
    ConflictingHitAndRun = check_multi_substring(c(3030, 3029), MOCodes),
    ConflictingInjury = check_multi_substring(c(3028, 3027, 3026, 3025, 3024), MOCodes),
    ConflictingRegion = check_multi_substring(c(4027, 4026, 4025, 4024), MOCodes),
    ConflictingPCF = check_multi_substring(c(3104, 3103, 3102), MOCodes),
    ConflictingLocation = check_multi_substring(c(4001:4021), MOCodes),
    ConflictingMOCodes = ConflictingIntersection == 1 |
      ConflictingCityProperty == 1 | ConflictingPrivateProperty == 1 |
      ConflictingDUI == 1 | ConflictingHitAndRun == 1 | 
      ConflictingInjury == 1 | ConflictingRegion == 1 |
      ConflictingPCF == 1 | ConflictingLocation == 1
  )

print(
  collisions_mo_code_conflicts %>% 
    group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
    summarise(ConflictingCodeRate = mean(ConflictingMOCodes, na.rm=TRUE)) %>%
    ggplot(aes(x=OccurredDate, y=ConflictingCodeRate)) +
    geom_line() +
    scale_y_continuous(label=scales::percent)
)
```

