---
title: "Traffic Collisions Data Validation"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Validating Data for File 3. Data Exploration
# Checking for intersections with multiple sets of coordinates
```{r}
collisions_cleaned %>% 
  filter(Latitude != 0 | Longitude != 0) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Collisions = n()) %>% 
  group_by(Address, CrossStreet) %>% 
  summarise(
    NumCoordinates = n(),
    MaxLat = max(Latitude),
    MinLat = min(Latitude),
    MaxLon = max(Longitude),
    MinLon = min(Longitude),
    LatRange = MaxLat - MinLat,
    LonRange = MaxLon - MinLon,
    LatStd = sd(Latitude),
    LonStd = sd(Longitude),
    AvgLat = mean(Latitude),
    AvgLon = mean(Longitude),
    ModeLat = mode(Latitude),
    ModeLon = mode(Longitude)
  ) %>%
  arrange(desc(NumCoordinates))
```

```{r}
print(collisions_cleaned %>% filter(Address == 'PACIFIC COAST HY', CrossStreet == 'FIGUEROA ST') %>%
        select(Address, CrossStreet, Latitude, Longitude, PremiseDescription) %>%
        group_by(Latitude, Longitude) %>% summarise(Count = n()) %>%
        ggplot(aes(x=Longitude, y=Latitude, size=Count)) +
        geom_point())
```

```{r}
collisions_cleaned %>% filter(Latitude == 33.9892, Longitude ==	-118.2827)
```

## Classifying the MO Codes, based on the PDF I found
```{r}
collision_mo_codes %>% 
  filter(!is.na(MOCode)) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

# Select the MOCodes of interest
collision_mo_codes %>% 
  filter(MOCode %in% c(3004, 3006, 3011, 3003, 3008, 3012,
                       3037, 3036, 3030, 3029, 3028, 3026, 3025, 3035, 3034, 3602, 3039, 3027, 3038, 1218
                       )) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

# 3004, 3006, 3011, 3003, 3008, 3012
# 3037 vs 3036, 3030/3029, 3028 vs 3026/3025, 3035 vs 3034, 3602, 3039, 3027, 3038, 1218
```

```{r}
collision_mo_codes %>% 
  filter(MOCode %in% c(0947, 3025,			
3101, 3401, 3701, 3006, 3030, 3032,			
0605, 3101, 3401, 3701, 3011, 3030, 3036,		
0605, 3101, 3401, 3701, 3006, 3030,			
3101, 3401, 3701, 3006, 3034, 3032,			
3101, 3401, 3701, 3006, 3030, 0605,			
3028, 4011, 3030, 3034, 0605, 3101, 3401, 3701, 3011, 3104)) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))

collisions_with_mo_labels %>% 
  filter(CrossStreet == '') %>%
  group_by(Latitude, Longitude) %>%
  summarise(AtIntersection = max(AtIntersection),
            PrivatePropertyInvolved = max(PrivatePropertyInvolved)) %>%
  filter(AtIntersection == 1, PrivatePropertyInvolved == 1) %>%
  select(Latitude, Longitude)

collisions_with_mo_labels %>%
  filter(Latitude == 34.1367	, Longitude == -118.2713) %>%
  select(AtIntersection, PrivatePropertyInvolved, CrossStreet)
```

```{r}
# checking for MO Code overlap
## at intersection y/n - rare overlap seems to be more at intersections
collisions_cleaned %>% filter(grepl('3037', MOCodes), grepl('3036', MOCodes))

## collision w/; rare overlap, should probably classify as multiple and also make individual columns
collisions_cleaned %>% filter(grepl('3004', MOCodes) + grepl('3006', MOCodes) + grepl('3011', MOCodes) + 
                              grepl('3003', MOCodes) + grepl('3008', MOCodes) + grepl('3012', MOCodes) > 1)

## hit and run misd vs felony; rare overlap, should probably classify as felony
collisions_cleaned %>% filter(grepl('3030', MOCodes), grepl('3029', MOCodes))

## injury y/n; rare overlap; should probably classify as complaint of injury
collisions_cleaned %>% filter(grepl('3028', MOCodes),  grepl('3026', MOCodes))
# also 3024, 3027

## city property involved y/n; rare overlap; should probably classify as y
collisions_cleaned %>% filter(grepl('3035', MOCodes), grepl('3034', MOCodes))
# misc - 3025, 3602, 3027, 3038, 1218

```

# Cleaning Up Some More
```{r}
collisions_with_mo_labels
```

```{r}
# Check that DRKey values are unique
collisions_with_mo_labels %>% nrow()
collisions_with_mo_labels %>% distinct(DRKey) %>% nrow()

# Range of ReportedDate and OccurredDate
collisions_with_mo_labels %>% summarise(min(ReportedDate), max(ReportedDate))
collisions_with_mo_labels %>% summarise(min(OccurredDate), max(OccurredDate))

# Distribution of OccurredTime
print(collisions_with_mo_labels %>% 
  group_by(OccurredTime = round(OccurredTime, -2)) %>% 
  summarise(Count = n()) %>%
  ggplot(aes(x=OccurredTime, y=Count)) +
  geom_bar(stat='identity'))

# Oddities w/ AreaName
  collisions_with_mo_labels %>% group_by(AreaName) %>% 
    summarise(Count = n()) %>% arrange(AreaName)
  
# Oddities w/ VictimSex
  collisions_with_mo_labels %>% group_by(VictimSex) %>% 
    summarise(Count = n()) %>% arrange(VictimSex)
  
# Oddities w/ VictimDescent
  collisions_with_mo_labels %>% group_by(VictimDescent) %>% 
    summarise(Count = n()) %>% arrange(VictimDescent)
  
# Oddities w/ PremiseDescription
  collisions_with_mo_labels %>% group_by(PremiseDescription) %>% 
    summarise(Count = n()) %>% arrange(PremiseDescription)
  
# Oddities w/ Address
  collisions_with_mo_labels %>% group_by(Address) %>% 
    summarise(Count = n()) %>% arrange(Address)
  
# Oddities w/ CrossStreet
  collisions_with_mo_labels %>% group_by(CrossStreet) %>% 
    summarise(Count = n()) %>% arrange(CrossStreet)
```

# Cleaning up the coordinates that are listed as (0, 0)
```{r}
collisions_with_mo_labels %>% filter(Latitude == 0 | Longitude == 0) %>% 
  group_by(Address, CrossStreet) %>%
  summarise(Collisions = n()) %>%
  left_join(
    collisions_with_mo_labels %>% filter(Latitude != 0, Longitude != 0) %>%
      group_by(Address, CrossStreet) %>% 
      summarise(Latitude = mean(Latitude),
                Longitude = mean(Longitude),
                NumExamples = n()),
    by=c('Address', 'CrossStreet')
  )

print(
  collisions_with_mo_labels %>% filter(Latitude == 0 | Longitude == 0) %>% 
    group_by(Address, CrossStreet) %>%
    summarise(Collisions = n()) %>%
    left_join(
      collisions_with_mo_labels %>% filter(Latitude != 0, Longitude != 0) %>%
        group_by(Address, CrossStreet) %>% 
        summarise(Latitude = mean(Latitude),
                  Longitude = mean(Longitude),
                  NumExamples = n()),
      by=c('Address', 'CrossStreet')
    ) %>%
    ggplot(aes(x=Longitude, y=Latitude, size=Collisions)) +
    geom_point()
)
```

```{r}
collisions_with_mo_labels %>% 
  filter(Latitude == 0 | Longitude == 0) %>% 
  group_by(Year = year(OccurredDate)) %>%
  summarise(Count = n()) %>% 
  arrange(Year)
# These are largely from 2012, 2014, 2020, 2021, 2023
```

# Checking that coordinates have consistent AreaNames and PremiseDescriptions
I already know the Addresses and CrossStreets can differ, even beyond formatting; this just further indicates I should ignore them
```{r}
# Not super consistent, establish a hierarchy
collisions_with_mo_labels %>% 
  group_by(Latitude, Longitude, AreaName) %>%
  summarise(1) %>%
  group_by(Latitude, Longitude) %>%
  summarise(NumAreaNames = n_distinct(AreaName),
            AreaNames = paste(AreaName, collapse = '; ')) %>%
  arrange(desc(NumAreaNames))

# Also not super consistent; maybe the coordinates aren't accurate?
collisions_with_mo_labels %>% 
  group_by(Latitude, Longitude, PremiseDescription) %>%
  summarise(1) %>%
  group_by(Latitude, Longitude) %>%
  summarise(NumPremiseDescriptions = n_distinct(PremiseDescription),
            PremiseDescriptions = paste(PremiseDescription, collapse = '; ')) %>%
  arrange(desc(NumPremiseDescriptions))
```

# Validating Data for File 4. MO Code Analysis
# Tracking Number of MOCodes over Time
```{r, fig.height=2, fig.width=4}
print(collisions_with_labeled_intersections %>%
        filter(OccurredDate >= '2013-01-01',
               OccurredDate < '2025-03-01') %>%
        group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
        summarise(AvgNumMOCodes = mean(NumMOCodes, na.rm=TRUE)) %>%
        ggplot(aes(x=OccurredDate, y=AvgNumMOCodes)) +
        geom_line() +
        labs(x='Date of Collisions', y='Number of MO Codes') +
        theme_set(theme_gray(base_size = 5)) +
        ggtitle('Avg Number of MO Codes over Time'))
```

```{r}
print(
  collision_mo_codes_labeled %>%
    mutate(Description = ifelse(Rank > 10, 'Other', Description)) %>%
    group_by(Description, OccurredDate = floor_date(OccurredDate, unit='months')) %>%
    summarise(TotalCollisions = n_distinct(DRKey)) %>%
    ggplot(aes(x=OccurredDate, y=TotalCollisions, color=Description)) +
    geom_line()
)

print(
  collision_mo_codes_pre_post_covid %>%
    ggplot(aes(x=ChangeInProportion, y=Description, fill=Description)) +
    geom_bar(stat='identity') +
    scale_x_continuous(label=scales::percent) +
    theme_set(theme_gray(base_size = 8)) +
    theme(legend.position='none') +
    labs(y='', x='Change in Proportion') +
    ggtitle('Shift in Proportion of Collisions Labelled, Pre- and Post-COVID')
)

```

## Proportion of collisions w/ conflicting MO codes over time
```{r}
collision_mo_codes_labeled %>%
  group_by(MOCode, Description) %>%
  summarise(TotalCollisions = n_distinct(DRKey)) %>%
  arrange(desc(MOCode))

print(
  collisions_mo_code_conflicts %>% 
    group_by(OccurredDate = floor_date(OccurredDate, unit='months')) %>%
    summarise(ConflictingCodeRate = mean(ConflictingMOCodes, na.rm=TRUE)) %>%
    ggplot(aes(x=OccurredDate, y=ConflictingCodeRate)) +
    geom_line() +
    scale_y_continuous(label=scales::percent)
)
```

```{r}
# histogram
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    ggplot(aes(x=CollisionAtIntersection, fill=CollisionAtIntersection)) +
    geom_histogram(stat='count') +
    ggtitle('Number of Collisions at Intersection in 2024')
)

# scatterplots
rounding_factor <- 1000
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>%
    summarise(Collisions = n(),
              ProportionAtIntersection = mean(AtIntersection)) %>%
    ggplot(aes(x=ProportionAtIntersection, y=Collisions)) +
    geom_point() +
    ggtitle('Number of Collisions at Locations by Proportion at Intersection, 2024')
)

rounding_factor <- 1000
print(
  collisions_with_injury_type %>% 
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor),
             CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=Collisions, fill=CollisionAtIntersection)) +
    geom_histogram(stat='count') +
    facet_wrap(~CollisionAtIntersection) +
    ggtitle('Histogram of Number of Collisions at Intersection in 2024')
)

# map
rounding_factor <- 100
print(
  collisions_with_injury_type %>%
    filter(year(OccurredDate) == 2024,
           Latitude != 0, Longitude != 0) %>%
    group_by(RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
             RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor)) %>%
    summarise(Collisions = n(),
              ProportionAtIntersection = mean(AtIntersection)) %>%
    arrange(desc(Collisions)) %>%
    ggplot(aes(x=RoundedLon, y=RoundedLat, color=ProportionAtIntersection,
               size=Collisions)) +
    geom_point() +
    ggtitle('Number of Collisions at Location by Proportion At Intersection, 2024')
)

# bar chart of injury severity by intersection vs non-intersection
print(
  collisions_with_injury_type %>%
    filter(year(OccurredDate) == 2024) %>%
    group_by(RoundedLat = round(Latitude, 3), 
             RoundedLon = round(Longitude, 3)) %>%
    mutate(NumCollisionsAtIntersection = sum(as.numeric(AtIntersection)),
           NumCollisionsNotAtIntersection = sum(as.numeric(NotAtIntersection))) %>%
    ungroup() %>%
    filter(NumCollisionsAtIntersection >= 5 |
           NumCollisionsNotAtIntersection >= 1) %>%
    group_by(CollisionAtIntersection, MostSevereInjuryReported) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=MostSevereInjuryReported, y=Collisions, fill=CollisionAtIntersection)) +
    geom_bar(stat='identity') +
    facet_wrap(~CollisionAtIntersection) +
    ggtitle('Number of Collision Injuries by Severity, Intersection vs Non-Intersection, 2024')
)
```

## Second round - more directed, hopefully more fruitful
```{r}
# looking to see whether intersections in general are more "dangerous" than non-intersections
# in a given area with both intersection collisions and non-intersection collisions, which happen more frequently
## i.e. what proportion of the collisions in these areas occurs at intersections

# also, in those areas, are injuries more likely/severe at intersections or not

# first, can we identify the right coordinates per intersection
## Seems to be too much variance in what coordinates an address x cross street combo can be listed as, maybe by mistake
## Should group by coordinates, and pick the nearest intersection by most common listing; round to nearest 1000th place
collisions_with_injury_type %>% 
  filter(AtIntersection) %>%
  group_by(Address, CrossStreet, Latitude, Longitude) %>% 
  summarise(Count = 1) %>%
  group_by(Address, CrossStreet) %>% 
  summarise(Count=n()) %>%
  arrange(desc(Count))

# 33.942, -118.4095 is a problematic set of coordinates; it's LAX and encompasses a ton of nearby locations
# There are some other coordinates that seem wrong
collisions_with_injury_type %>%
  mutate(NearestIntersection = paste0(Address, ' and ', CrossStreet)) %>%
  group_by(RoundedLat = round(Latitude, 3),
           RoundedLon = round(Longitude, 3),
           NearestIntersection) %>%
  summarise(Count = n()) %>%
  group_by(RoundedLat, RoundedLon) %>%
  summarise(Count = n()) %>% arrange(desc(Count))

# trying again to round out the intersection areas by name and coordinates
# besides a couple exceptions, it seems any location with a cross street and any collisions being listed 
# at intersection should be labeled an intersection, 

# comparing the possible outcomes
collisions_at_intersection_areas %>%
  group_by(ListedAsIntersection, ListedAsNotIntersection,
           ListedMoreAsAtIntersection, HasCrossStreet) %>% 
  summarise(Collisions = sum(Collisions))

collisions_at_intersection_areas %>% 
  filter(ListedAsIntersection == 1,
         ListedAsNotIntersection == 1,
         ListedMoreAsAtIntersection == 0,
         HasCrossStreet == 0) %>% 
  arrange(desc(Collisions))

```

```{r}
# tracking the intersection labeling over time
print(
  collisions_with_injury_type %>% 
    filter(OccurredDate >= '2013-01-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='weeks'),
             CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, color=CollisionAtIntersection)) +
    geom_line()
)

# tracking the intersection labeling by location over time
print(
  collisions_with_injury_type %>% 
    filter(OccurredDate >= '2013-01-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='weeks'),
             RoundedLat = round(Latitude, 4),
             RoundedLon = round(Longitude, 4)) %>%
    summarise(
      AtIntersection = max(AtIntersection),
      NotAtIntersection = max(NotAtIntersection),
    ) %>% 
    mutate(
      labeledAsIntersection = case_when(
        AtIntersection == 1 & NotAtIntersection == 1 ~ 'Both',
        AtIntersection == 1 ~ 'At Intersection',
        NotAtIntersection == 1 ~ 'Not At Intersection',
        TRUE ~ 'Unspecified'
      )) %>% 
    group_by(OccurredWeek, labeledAsIntersection) %>%
    summarise(NumLocations = n()) %>%
    ggplot(aes(x=OccurredWeek, y=NumLocations, color=labeledAsIntersection)) +
    geom_line()
)
```

```{r}
collisions_with_injury_type %>% 
  filter(round(Latitude, 3) == 33.989, round(Longitude, 3) == -118.309) %>%
  select(ReportingDistrict, AtIntersection, NotAtIntersection)

collisions_with_injury_type %>% distinct(Address)

collision_mo_codes %>% 
  filter(Latitude == 33.9470, Longitude == -118.2684) %>% 
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% 
  summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))
```
## Trying this again, without street names
```{r}
identifying_intersections %>% 
  filter(HasCrossStreet == 0, ListedAsIntersection == 1) %>%
  arrange(desc(Collisions))

identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0) %>%
  arrange(desc(Collisions))

identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0) %>%
  arrange(desc(NumCollisionsNotAtIntersection))

# Basically, if the coordinates are ever listed with a CrossStreet, they are at an intersection
```

## Double-checking that I've re-labeled intersections correctly
```{r}
identifying_intersections %>% 
  filter(HasCrossStreet == 1, ListedAsIntersection == 0, ListedAsNotIntersection == 1) %>%
  arrange(desc(NumCollisionsNotAtIntersection))
# Far as I can tell, the vast majority of coordinates listed with CrossStreets point to intersections
```

```{r}
# Check the MO Codes of one of these re-labeled intersections that looks iffy
collision_mo_codes %>% 
  filter(!is.na(MOCode), Latitude == 34.1589, Longitude ==	-118.4994) %>%
  left_join(mo_code_descriptions, by='MOCode') %>%
  group_by(MOCode, Description) %>% summarise(Collisions = n_distinct(DRKey)) %>% 
  arrange(desc(Collisions))
```

# Validating Data for File 5. Demographic Trends
# Parsing ZIP Codes
```{r}
print(
  zip_codes_indexed %>%
    ggplot(aes(x=DistanceToNext)) +
    geom_histogram(binwidth = 0.001)
)

zip_codes_indexed %>% arrange(desc(DistanceToNext))
zip_codes_indexed %>% group_by(DistanceToNext = round(DistanceToNext, 4)) %>%
  summarise(Count = n()) %>% arrange(desc(DistanceToNext))
```

## Labeling the collision coordinates ZIP codes using a Latitude-wise method
```{r}
for (zip_code in relevant_zip_codes) {
  print(
    all_coordinates %>% filter(ZIPCode == local(zip_code),
                               Backfilled == Frontfilled | CoordinateType == 'Boundary'
                               ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=CoordinateType)) +
      geom_point(size=1) +
      ggtitle(local(zip_code))
  )
}
```

Looking at coordinates with conflicting ZIP code labels, the vast majority of these are along boundaries and could reasonably be attributed to either ZIP code. ZIP code 91330 lies within another ZIP code, so the method used gives conflicting results for coordinates well outside its bounds; this has been fixed above
```{r}
for (i in c(1:nrow(conflicted_coordinates))) {
  print(
    all_coordinates %>% 
      filter(ZIPCode %in% c(conflicted_coordinates[i, 'Backfilled'],
                            conflicted_coordinates[i, 'Frontfilled'])) %>%
      mutate(IsConflicted = ifelse(Backfilled == conflicted_coordinates[i, 'Backfilled'][[1]] &
                                   Frontfilled == conflicted_coordinates[i, 'Frontfilled'][[1]] &
                                   CoordinateType == 'Collision', 1, 0)) %>%
      arrange(IsConflicted) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=as.factor(IsConflicted))) +
      geom_point(size=1) +
      ggtitle(local(zip_code))
  )
}
```


```{r}
census_by_zip %>% group_by(LabelGroup1, LabelGroup2, LabelGroup3) %>% summarise(n())
```