---
title: "Traffic Collisions Location Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
library(Morpho)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Census data
```{r}
parse_census_data <- function(year) {
  
  census_overall <- read.csv(paste('Data/Census Data/demographics_',
                                 year, '.csv', sep='')) %>%
      mutate(Year = local(year),
             Demographic = str_trim(`Label..Grouping.`),
             Demographic = case_when(
               Demographic %in% c('White Alone') ~ 'White',
               Demographic %in% c('Black or African American alone') ~ 'Black',
               Demographic %in% c('Asian alone') ~ 'Asian',
               Demographic %in% c('Some Other Race alone') ~ 'Other',
               Demographic %in% c('Two or More Races') ~ 'Two or More Races',
               Demographic %in% c('Hispanic or Latino (of any race)') ~ 'Hispanic',
               TRUE ~ Demographic),
             DemographicType = case_when(
               Demographic %in% c('Male', 'Female') ~ 'Sex',
               Demographic %in% c('Under 5 years', '5 to 9 years', '10 to 14 years', 
                                  '15 to 19 years', '20 to 24 years', '25 to 34 years', 
                                  '35 to 44 years', '45 to 54 years', '55 to 59 years', 
                                  '60 to 64 years', '65 to 74 years', '75 to 84 years', 
                                  '85 years and over') ~ 'Age',
               Demographic %in% c('White', 'Black', 'Asian', 'Other', 
                                  'Two or More Races', 'Hispanic') ~ 'RaceEthnicity',
               TRUE ~ NA),
             Population = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ',')),
             TotalPopulation = ifelse(Demographic %in% c('Total population'), Population, NA)) %>%
      group_by(Year) %>% mutate(TotalPopulation = max(TotalPopulation, na.rm=TRUE)) %>% ungroup() %>%
      filter(!is.na(DemographicType)) %>%
      select(Year, DemographicType, Demographic, Population, TotalPopulation)

  return(census_overall)
}

census_data <- data.frame()

for (year in c(2013:2024)) {census_data <- rbind(census_data, parse_census_data(year))}

census_data %>% write.csv('Cleaned Data/LACensusDataOverall.csv', row.names=FALSE)
```

## Age trends over time
```{r}
print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    ggplot(aes(x=Year, y=Population, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group')
)

print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    mutate(Proportion = Population/Total) %>%
    ggplot(aes(x=Year, y=Proportion, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group') +
    scale_y_continuous(labels=scales::percent)
)
```

## Breakdown of collisions by race/ethnicity
```{r, fig.height=4, fig.width=6}
for (race in c('Asian', 'Black American', 'Hispanic/Latino', 'White')) {
  print(
    collisions_with_labeled_intersections %>%
      filter(Latitude != 0, Longitude != 0, VictimEthnicity == local(race)) %>%
      group_by(Latitude = round(Latitude, 3), Longitude = round(Longitude, 3), VictimEthnicity) %>%
      summarise(Collisions = n()) %>%
      arrange(Collisions) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point() +
      xlim(-118.7, -118.1) +
      ylim(33.6, 34.4) +
      facet_wrap(~VictimEthnicity)
  )
}
```

## Collision rates over time
```{r}
print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    mutate(CollisionRate = Collisions/TotalPopulation) %>%
    ggplot(aes(x=Year, y=CollisionRate)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA))
)

print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    left_join(commuting_data_overall, by='Year') %>%
    mutate(CollisionRate = Collisions/TotalPopulation,
           CarCommutersByPopulation = (Driving + Carpooling)/TotalPopulation
           ) %>%
    ggplot(aes(x=CarCommutersByPopulation, y=CollisionRate, color=as.factor(Year))) +
    geom_point() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA)) +
    scale_x_continuous(labels=scales::percent) +
    labs(x='Number of Car Commuters per Population', y='Collisions per Population', color='Year')
)
```

# Commutes over Time
## Loading and cleaning data
```{r}
pivot_commuting_overall <- function(year) {
  commuting_overall <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                               year, '.csv', sep='')) %>%
    mutate(Year = local(year),
           Labels = str_trim(`Label..Grouping.`),
           Labels = case_when(
             Labels == 'Taxi or ride-hailing services' ~ 'Taxicab', 
             Labels == 'Public transportation (excluding taxicab):' ~ 'Public transportation:',
             Labels == 'Worked at home' ~ 'Worked from home',
             TRUE ~ Labels),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, Labels, Values) %>%
    filter(Labels %in% c('Total:', 'Drove alone', 'Carpooled:', 'Public transportation:', 
                         'Walked', 'Taxicab', 'Motorcycle', 'Bicycle', 'Other means', 'Worked from home')) %>%
    pivot_wider(names_from=Labels, values_from=Values) %>%
    mutate(TaxiBikeAndOther = Taxicab + Motorcycle + Bicycle + `Other means`) %>% 
    select(Year, `Total:`, `Drove alone`, `Carpooled:`, `Public transportation:`, 
           Walked, TaxiBikeAndOther, `Worked from home`)
  
  colnames(commuting_overall) <- c('Year', 'TotalWorkers', 'Driving', 'Carpooling', 'PublicTransit', 
                                   'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(commuting_overall)
}


pivot_commuting_data_by_ethn <- function(year, file_subname, race_ethnicity_label) {
  
  ethn_df <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                            file_subname, '_', year, '.csv', sep='')) %>% 
    mutate(RaceEthnicity = local(race_ethnicity_label), Year = local(year),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, RaceEthnicity, `Label..Grouping.`, Values) %>%
    pivot_wider(names_from=`Label..Grouping.`, values_from=Values)
  
  colnames(ethn_df) <- c('Year', 'RaceEthnicity', 'TotalWorkers', 'Driving', 'Carpooling',
                         'PublicTransit', 'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(ethn_df)
}

parse_commute_data <- function(years) {
  overall <- data.frame()
  by_ethn <- data.frame()
  by_lang <- data.frame()
  
  for (year in years) {
    overall <- rbind(overall, pivot_commuting_overall(year))
    by_ethn <- rbind(by_ethn, pivot_commuting_data_by_ethn(year, 'asian', 'Asian') %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'black', 'Black')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'hispanic', 'Hispanic')) %>%
                       # rbind(pivot_commuting_data_by_ethn(year, 'white', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'white_nonhispanic', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'multiple_races', 'Two or More Races')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'other', 'Other')))
    # by_lang <- rbind(by_lang, read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_by_language_', year, '.csv', sep='')))
  }
  
  return(list(overall, by_ethn, by_lang))
}

commuting_data <- parse_commute_data(c(2013:2024))
commuting_data_overall <- commuting_data[1] %>% as.data.frame()
commuting_data_by_ethnicity <- commuting_data[2] %>% as.data.frame()

commuting_data_overall %>% write.csv('Cleaned Data/LACommutingDataOverall.csv', row.names=FALSE)
commuting_data_by_ethnicity %>% write.csv('Cleaned Data/LACommutingDataByEthnicity.csv', row.names=FALSE)
```

```{r}
commuting_data_overall <- read.csv('Cleaned Data/LACommutingDataOverall.csv')
commuting_data_by_ethnicity <- read.csv('Cleaned Data/LACommutingDataByEthnicity.csv')
```

## Commuting trends over time
```{r}
# Proportion of each mode of transportation over time
print(
  commuting_data_overall %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)

print(
  commuting_data_by_ethnicity %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    facet_wrap(~RaceEthnicity) +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)
```

## Commuting shifts after COVID
```{r}
commuting_by_covid_period <- commuting_data_by_ethnicity %>%
  filter(Year >= 2016, Year != 2020) %>%
  mutate(Period = ifelse(Year < 2020, 'Pre-COVID', 'Post-COVID') %>%
           factor(levels = c('Pre-COVID', 'Post-COVID'))) %>%
  group_by(Period, RaceEthnicity) %>%
  summarise(
    AvgWorkersPerYear = mean(TotalWorkers) %>% round(),
    AvgCommutingByCar = mean(Driving + Carpooling) %>% round(),
    AvgCommutingByPublicTransit = mean(PublicTransit) %>% round(),
    AvgCommutingByWalking = mean(Walking) %>% round(),
    AvgCommutingByTaxiBikeAndOther = mean(TaxiBikeAndOther) %>% round(),
    AvgWorkFromHome = mean(WorkingFromHome) %>% round(),
    
    CommutingByCarRate = round(AvgCommutingByCar/AvgWorkersPerYear, 2),
    CommutingByPublicTransitRate = round(AvgCommutingByPublicTransit/AvgWorkersPerYear, 2),
    CommutingByWalkingRate = round(AvgCommutingByWalking/AvgWorkersPerYear, 2),
    CommutingByTaxiBikeAndOtherRate = round(AvgCommutingByTaxiBikeAndOther/AvgWorkersPerYear, 2),
    WFHRate = round(AvgWorkFromHome/AvgWorkersPerYear, 2),
  )

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=CommutingByCarRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Commuting by Car Rate', fill='') +
    ggtitle('Proportion of Workers Commuting by Car, Pre- and Post-COVID')
)

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=WFHRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Working from Home Rate', fill='') +
    ggtitle('Proportion of Workers Working from Home, Pre- and Post-COVID')
)
```

## Comparing number of collision victims by ethnicity, compared to WFH rate
Too correlated w/ decrease in collisions after COVID
```{r}
collisions_by_ethnicity <- collisions_with_labeled_intersections %>%
  mutate(
    VictimEthnicity = case_when(
      VictimDescent == 'H' ~ 'Hispanic',
      VictimDescent %in% c('A', 'C', 'D', 'F', 'G', 'J', 'K', 'L', 'V', 'Z') ~ 'Asian',
      VictimDescent %in% c('P', 'S', 'U') ~ 'Pacific Islander',
      VictimDescent == 'B' ~ 'Black',
      VictimDescent == 'W' ~ 'White',
      VictimDescent %in% c('O', 'X', '', '-') ~ 'Other',
      VictimDescent == 'I' ~ 'Native American/Alaskan',
      TRUE ~ '?'    )
  ) %>%
  group_by(OccurredYear, VictimEthnicity) %>%
  summarise(Collisions = n()) %>%
  right_join(commuting_data_by_ethnicity, 
             by=c('VictimEthnicity'='RaceEthnicity', 'OccurredYear'='Year')) %>%
  right_join(census_data %>% filter(DemographicType == 'RaceEthnicity'),
             by=c('VictimEthnicity'='Demographic', 'OccurredYear'='Year')) %>%
  mutate(
    Car = Driving + Carpooling,
    DrivingProportion = Car/TotalWorkers,
    WalkingProportion = Walking/TotalWorkers,
    PublicTransitProportion = PublicTransit/TotalWorkers,
    WFHProportion = WorkingFromHome/TotalWorkers,
    CollisionRate = Collisions/Population)
```

```{r}
print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=DrivingProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Driving to Work vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)

print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=WFHProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Working from Home vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)
```

## Distributions of collision time over the years
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(!is.na(OccurredTime), OccurredYear >= 2016, 
           OccurredYear < 2025, OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID'),) %>%
    group_by(Period) %>% mutate(Total = n()) %>% ungroup() %>%
    group_by(OccurredTime = round(OccurredTime, -2), Period) %>%
    summarise(Distribution = n()/first(Total)) %>%
    ggplot(aes(x=OccurredTime, y=Distribution, color=Period)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Occurrence Time, Pre- and Post-COVID')
)
```

# Parsing ZIP Codes
```{r}
zip_code_boundaries <- read_json('Data/ZIP Codes/Zipcodes.geojson')

# zip_codes_df <- data.frame()

# for (i in c(1:length(zip_code_boundaries$features))) {
#   row <- zip_code_boundaries$features[[i]]
# 
#   object_id <- row$properties$OBJECTID
#   zip_code <- row$properties$ZIPCODE
#   shape_type <- row$geometry$type
# 
#   if (shape_type == 'Polygon') {
#     coordinates <- row$geometry$coordinates[[1]]
#   } else {
#     coordinates <- row$geometry$coordinates[[1]][[1]]
#   }
#   for (coord_set in coordinates) {
#     zip_codes_df <- rbind(zip_codes_df,
#                           list('ObjectID'=object_id, 'ZIPCode'=zip_code,
#                                'Longitude'=coord_set[[1]],
#                                'Latitude'=coord_set[[2]]))
#   }
# }
# 
# zip_codes_df <- zip_codes_df %>% distinct()
# zip_codes_df %>% group_by(ObjectID) %>% summarise(n())

# zip_codes_df %>% write.csv('Cleaned Data/LAZipCodes.csv', row.names=FALSE)
zip_codes_df <- read.csv('Cleaned Data/LAZipCodes.csv')
```

## Interpolating the ZIP code boundary coordinates
```{r}
zip_codes_indexed <- zip_codes_df %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  group_by(ZIPCode) %>% 
  mutate(Index =  row_number()) %>% 
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2))
```

```{r}
print(
  zip_codes_indexed %>%
    ggplot(aes(x=DistanceToNext)) +
    geom_histogram(binwidth = 0.001)
)

zip_codes_indexed %>% arrange(desc(DistanceToNext))
zip_codes_indexed %>% group_by(DistanceToNext = round(DistanceToNext, 4)) %>%
  summarise(Count = n()) %>% arrange(desc(DistanceToNext))
```

```{r}
interpolate_coordinates <- function(distance_to_next, boundary_latitude, boundary_longitude, 
                                    next_boundary_latitude, next_boundary_longitude, interpol_distance) {
    num_steps <- floor(distance_to_next/interpol_distance)
    latitude_diff <- next_boundary_latitude - boundary_latitude
    longitude_diff <- next_boundary_longitude - boundary_longitude
    
    new_latitudes <- ''
    new_longitudes <- ''
    
    for (i in c(1:num_steps)) {
      new_latitudes <- paste(new_latitudes, ', ', boundary_latitude + latitude_diff*i/(1+num_steps))
      new_longitudes <- paste(new_longitudes, ', ', boundary_longitude + longitude_diff*i/(1+num_steps))
    }
    
    return(paste(new_latitudes, new_longitudes, sep=';'))
  }

start_time <- Sys.time()
interpolation_distance <- 0.00005

zip_codes_interpolated <- zip_codes_indexed %>%
  rowwise() %>%
  mutate(
    InterpolatedCoordinates = ifelse(
      DistanceToNext > local(interpolation_distance) & (DistanceToNext < 0.0095 | 
                                                          !(ZIPCode %in% c('90803', '90732'))), # 90803 has some non-contiguous areas 
      interpolate_coordinates(DistanceToNext, Latitude, Longitude,
                              NextLatitude, NextLongitude, local(interpolation_distance)), NA)
  ) %>%
  separate(InterpolatedCoordinates, sep=';', 
           into=c('InterpolatedLatitude', 'InterpolatedLongitude')) %>%
  separate_longer_delim(c(InterpolatedLatitude, InterpolatedLongitude), delim=', ') %>%
  mutate(
    InterpolatedLongitude = as.numeric(InterpolatedLongitude),
    InterpolatedLatitude = as.numeric(InterpolatedLatitude),
    Longitude = ifelse(is.na(InterpolatedLongitude), Longitude, InterpolatedLongitude),
    Latitude = ifelse(is.na(InterpolatedLatitude), Latitude, InterpolatedLatitude),
  ) %>% select(-InterpolatedLatitude, -InterpolatedLongitude)

zip_codes_reindexed <- zip_codes_interpolated %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  mutate(Latitude = round(Latitude, 4), Longitude = round(Longitude, 4)) %>%
  distinct() %>%
  group_by(ZIPCode) %>%
  mutate(Index = row_number()) %>%
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2 ))

print(Sys.time() - start_time)

zip_codes_reindexed %>% filter(DistanceToNext > 0.0001) # 0.0001414214
zip_codes_reindexed %>% mutate(LatitudeDiff = NextLatitude - Latitude) %>% filter(LatitudeDiff > 0.0001)
zip_codes_reindexed %>% mutate(LongitudeDiff = NextLongitude - Longitude) %>% filter(LongitudeDiff > 0.0001)

# zip_codes_reindexed %>% write.csv('Cleaned Data/LAZipCodesInterpolated.csv', row.names=FALSE)
# zip_codes_reindexed <- read.csv('Cleaned Data/LAZipCodesInterpolated.csv')
```

## Labeling the collision coordinates ZIP codes using a Latitude-wise method
```{r}
all_coordinates <- zip_codes_reindexed %>% mutate(CoordinateType = 'Boundary') %>%
  mutate(Latitude = round(Latitude, 4), Longitude = round(Longitude, 4),
         LateralDirection = case_when(
           NextLatitude > Latitude ~ 1,
           NextLatitude < Latitude ~ -1,
           TRUE ~ 0),
         Longitude = case_when(
           LateralDirection == 1 ~ Longitude - 0.00009,
           LateralDirection == -1 ~ Longitude + 0.00009,
           TRUE ~ Longitude)) %>%
  select(Latitude, Longitude, ZIPCode, CoordinateType) %>%
  rbind(collisions_with_labeled_intersections %>% 
          filter(Latitude != 0, Longitude != 0) %>% 
          mutate(ZIPCode = NA, CoordinateType = 'Collision') %>% 
          select(Latitude, Longitude, ZIPCode, CoordinateType) %>% distinct()) %>%
  arrange(Latitude, Longitude, ZIPCode) %>%
  mutate(IsCollisionPoint = ifelse(CoordinateType == 'Collision', 1, 0)) %>%
  group_by(Latitude, ZIPCode) %>%
  mutate(MinLongitude = min(Longitude), MaxLongitude = max(Longitude)) %>% ungroup() %>%
  mutate(Backfilled = case_when(CoordinateType == 'Collision' ~ 0, 
                                Longitude == MaxLongitude ~ NA, TRUE ~ ZIPCode),
         Frontfilled = case_when(CoordinateType == 'Collision' ~ 0,
                                 Longitude == MinLongitude ~ NA, TRUE ~ ZIPCode),) %>%
  fill(Backfilled, .direction='up') %>%
  fill(Frontfilled, .direction='down') %>%
  mutate(
    Backfilled = ifelse(Backfilled == 0, NA, Backfilled),
    Frontfilled = ifelse(Frontfilled == 0, NA, Frontfilled),
  )  %>%
  fill(Backfilled, .direction='down') %>%
  fill(Frontfilled, .direction='up') %>%
  mutate(ZIPCode = ifelse(CoordinateType == 'Collision', 
                          case_when(
                            Backfilled == 91330 ~ Frontfilled,
                            TRUE ~ Backfilled), ZIPCode))
```

```{r}
relevant_zip_codes <- (all_coordinates %>% group_by(ZIPCode) %>% 
  summarise(NumCollisionPoints = sum(IsCollisionPoint)) %>%
  filter(NumCollisionPoints > 0) %>% ungroup() %>%
  select(ZIPCode))$ZIPCode

for (zip_code in relevant_zip_codes) {
  print(
    all_coordinates %>% filter(ZIPCode == local(zip_code),
                               Backfilled == Frontfilled | CoordinateType == 'Boundary'
                               ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=CoordinateType)) +
      geom_point(size=1) +
      ggtitle(local(zip_code))
  )
}
```

Looking at coordinates with conflicting ZIP code labels, the vast majority of these are along boundaries and could reasonably be attributed to either ZIP code. ZIP code 91330 lies within another ZIP code, so the method used gives conflicting results for coordinates well outside its bounds; this has been fixed above
```{r}
conflicted_coordinates <- all_coordinates %>%
  filter(CoordinateType == 'Collision',
         Backfilled != Frontfilled) %>%
  group_by(Backfilled, Frontfilled) %>% 
  summarise(1)

for (i in c(1:nrow(conflicted_coordinates))) {
  print(
    all_coordinates %>% 
      filter(ZIPCode %in% c(conflicted_coordinates[i, 'Backfilled'],
                            conflicted_coordinates[i, 'Frontfilled'])) %>%
      mutate(IsConflicted = ifelse(Backfilled == conflicted_coordinates[i, 'Backfilled'][[1]] &
                                   Frontfilled == conflicted_coordinates[i, 'Frontfilled'][[1]] &
                                   CoordinateType == 'Collision', 1, 0)) %>%
      arrange(IsConflicted) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=as.factor(IsConflicted))) +
      geom_point(size=1) +
      ggtitle(local(zip_code))
  )
}
```

# Breaking Down Demographics by ZIP Code
I used (all_coordinates %>% filter(CoordinateType == 'Collision') %>% select(ZIPCode) %>% distinct() %>% arrange(ZIPCode))$ZIPCode to list all the relevant ZIP codes and collected the overall census and commuter data for 2013-2024 for these ZIP codes. Each CSV contains the data for 20-40 of the ZIP codes due to filesize limits
```{r}
census_filenames <-list.files('Data/Census Data/By ZIP Code')
commuter_filenames <-list.files('Data/Commuting Data/By ZIP Code')
```

## Parsing census data
First, we need to standardize the Label Groupings, as they vary year-to-year. I've used the following script to narrow them down to the most important characteristics that are observed each year
```{r}
label_groupings <- data.frame()

parse_label_groupings <- function(label_grouping) {
  stripped_label_grouping <- str_trim(label_grouping)
  num_indents <- floor((str_length(label_grouping) - str_length(stripped_label_grouping))/4)
  
  if (num_indents == 0) {
    return(paste(stripped_label_grouping, 'None', 'None', 'None', 'None', sep=';'))
  } else if (num_indents == 1) {
    return(paste('', stripped_label_grouping, 'None', 'None', 'None', sep=';'))
  } else if (num_indents == 2) {
    return(paste('', '',  stripped_label_grouping, 'None', 'None', sep=';'))
  } else if (num_indents == 3) {
    return(paste('', '', '', stripped_label_grouping, 'None', sep=';'))
  } else if (num_indents == 4) {
    return(paste('', '', '', '', stripped_label_grouping, sep=';'))
  }
}

for (filename in census_filenames) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()
  
  label_groupings <- rbind(label_groupings, 
    read.csv(paste('Data/Census Data/By ZIP Code/', filename, sep='')) %>%
      rowwise() %>%
      mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
             Year = local(year_of_data)) %>%
      separate(LabelGroupings, sep=';', into=c('LabelGroup1',
                                               'LabelGroup2',
                                               'LabelGroup3',
                                               'LabelGroup4',
                                               'LabelGroup5')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~na_if(.,'')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5), ~toupper(.)) %>%
      fill(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
      mutate(
        LabelGroup1 = ifelse(LabelGroup1 == 'TOTAL POPULATION', 
                             LabelGroup2, LabelGroup1),
        LabelGroup2 = case_when(
          LabelGroup2 == 'TOTAL POPULATION' | 
            LabelGroup2 == LabelGroup1 ~ LabelGroup3,
          TRUE ~ LabelGroup2),
        LabelGroup3 = case_when(
          LabelGroup3 == 'TOTAL POPULATION' | 
            LabelGroup3 == LabelGroup2 ~ LabelGroup4,
          TRUE ~ LabelGroup3),
        LabelGroup4 = case_when(
          LabelGroup4 == 'TOTAL POPULATION' | 
            LabelGroup4 == LabelGroup3 ~ LabelGroup5,
          TRUE ~ LabelGroup4),
        LabelGroup5 = case_when(
          LabelGroup5 == 'TOTAL POPULATION' | 
            LabelGroup5 == LabelGroup4 ~ 'NONE',
          TRUE ~ LabelGroup5),
      ) %>%
      select(Year, LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5))
}

label_groupings_summarised <- label_groupings %>% 
  group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5) %>%
  summarise(Count= n_distinct(Year)) %>% arrange(desc(Count)) %>% ungroup() %>%
  filter(LabelGroup4 == 'NONE', LabelGroup5 == 'NONE', LabelGroup2 != 'NONE',
         LabelGroup3 == 'NONE' | (LabelGroup1 == 'RACE' & LabelGroup2 == 'ONE RACE'),
         LabelGroup3 != 'NONE' | (LabelGroup1 != 'RACE' & LabelGroup2 != 'ONE RACE'),
         Count == 12, LabelGroup2 != 'NOT HISPANIC OR LATINO', LabelGroup2 != 'MEDIAN AGE (YEARS)',
         !(grepl('YEARS AND OVER', LabelGroup2)) | LabelGroup2 == '85 YEARS AND OVER',
         ) %>% select(LabelGroup1, LabelGroup2, LabelGroup3)
```

Now, we can use label_groupings_summarised to only pull the most important data from each CSV
```{r}
parse_census_data_by_zip_code <- function(filename) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()

  census_data_by_zip <- read.csv(paste('Data/Census Data/By ZIP Code/', filename, sep=''))  %>%
    rowwise() %>%
    mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
           Year = local(year_of_data)) %>%
    separate(LabelGroupings, sep=';', into=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3',
                                             'LabelGroup4', 'LabelGroup5')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~na_if(.,'')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4, LabelGroup5), ~toupper(.)) %>%
    fill(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
    mutate(
      LabelGroup1 = ifelse(LabelGroup1 == 'TOTAL POPULATION', 
                           LabelGroup2, LabelGroup1),
      LabelGroup2 = case_when(
        LabelGroup2 == 'TOTAL POPULATION' | 
          LabelGroup2 == LabelGroup1 ~ LabelGroup3,
        TRUE ~ LabelGroup2),
      LabelGroup3 = case_when(
        LabelGroup3 == 'TOTAL POPULATION' | 
          LabelGroup3 == LabelGroup2 ~ LabelGroup4,
        TRUE ~ LabelGroup3),
      LabelGroup4 = case_when(
        LabelGroup4 == 'TOTAL POPULATION' | 
          LabelGroup4 == LabelGroup3 ~ LabelGroup5,
        TRUE ~ LabelGroup4),
      LabelGroup5 = case_when(
        LabelGroup5 == 'TOTAL POPULATION' | 
          LabelGroup5 == LabelGroup4 ~ 'NONE',
        TRUE ~ LabelGroup5),
    )  %>% 
    filter(LabelGroup4 == 'NONE', LabelGroup5 == 'NONE') %>%
    select(-`Label..Grouping.`, -LabelGroup4, -LabelGroup5) %>%
    select(Year, LabelGroup1, LabelGroup2, LabelGroup3, everything()) %>% 
    pivot_longer(cols=starts_with('ZCTA'), names_to = 'ZIP', values_to = 'Population', values_transform = as.character) %>%
    filter(grepl('..Estimate', ZIP)) %>%
    mutate(ZIP = substring(ZIP, 7, 11),
           Population = str_remove(Population, ',') %>% as.numeric()) %>%
    inner_join(label_groupings_summarised, by=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3'))
  
  return(census_data_by_zip)
}
```

```{r}
census_by_zip <- data.frame()

for (filename in census_filenames) {
  census_by_zip <- rbind(census_by_zip, parse_census_data_by_zip_code(filename))
}

# checking the completeness of the data; ZIP code 90090 is missing data for 2021-2024, but these years are simply not available in the census data
for (year in c(2013:2024)) {
  census_by_zip %>% 
    # group_by(ZIP) %>% summarise(Count = n()) %>% arrange(desc(Count))
    group_by(ZIP) %>% mutate(Count = n()) %>% filter(Count < 264, Year == local(year)) %>% ungroup() %>%
    arrange(ZIP, LabelGroup1, LabelGroup2, LabelGroup3) %>%
    print()
}

census_by_zip %>% write.csv('Cleaned Data/LACensusDataByZIP.csv', row.names=FALSE)
```

```{r}
census_by_zip <- read.csv('Cleaned Data/LACensusDataByZIP.csv')
census_by_zip %>% group_by(LabelGroup1, LabelGroup2, LabelGroup3) %>% summarise(n())

census_by_zip_pivoted <- census_by_zip %>%
  mutate(LabelGroup = case_when(
      LabelGroup1 == 'RACE' & LabelGroup2 == 'ONE RACE' ~ case_when(
        LabelGroup3 == 'BLACK OR AFRICAN AMERICAN' ~ 'BLACK',
        LabelGroup3 == 'AMERICAN INDIAN AND ALASKA NATIVE' ~ 'INDIGENOUS',
        LabelGroup3 == 'NATIVE HAWAIIAN AND OTHER PACIFIC ISLANDER' ~ 'PACIFIC ISLANDER',
        LabelGroup3 == 'SOME OTHER RACE' ~ 'OTHER',
        TRUE ~ LabelGroup3),
      LabelGroup1 == 'HISPANIC OR LATINO AND RACE' ~ 'HISPANIC',
      TRUE ~ LabelGroup2) %>% str_to_title() %>% 
        str_replace_all(' To ', ' to ') %>% str_replace_all(' And ', ' and ') %>%
        str_replace_all(' ', '')
  ) %>% select(Year, ZIP, LabelGroup, Population) %>%
  pivot_wider(names_from=LabelGroup, values_from=Population)
```

## Parsing commuting data
```{r}
label_groupings <- data.frame()
filename <- "ACSDT5Y2013.B08301-2026-07-28T222015.csv"

for (filename in commuter_filenames) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()
  
  label_groupings <- rbind(label_groupings, 
    read.csv(paste('Data/Commuting Data/By ZIP Code/', filename, sep='')) %>%
      rowwise() %>%
      mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
             LabelGroupings = str_replace(LabelGroupings, ':', ''),
             Year = local(year_of_data)) %>%
      separate(LabelGroupings, sep=';', into=c('LabelGroup1',
                                               'LabelGroup2',
                                               'LabelGroup3',
                                               'LabelGroup4')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3), ~na_if(.,'')) %>%
      mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~toupper(.)) %>%
      fill(LabelGroup1, LabelGroup2, LabelGroup3) %>%
      select(Year, LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4))
}

label_groupings %>% group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>% 
  summarise(Count = n()) %>% arrange(desc(Count))

label_groupings_summarised <- label_groupings %>% 
  group_by(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4) %>%
  summarise(Count= n_distinct(Year)) %>% arrange(desc(Count)) %>% ungroup() %>%
  filter(LabelGroup4 == 'NONE', LabelGroup2 != 'NONE',
         LabelGroup3 == 'NONE' | (LabelGroup3 %in% c('DROVE ALONE', 'CARPOOLED'))
         ) %>% select(LabelGroup1, LabelGroup2, LabelGroup3)
```

```{r}
parse_commuting_data_by_zip_code <- function(filename) {
  year_of_data <- filename %>% substring(8, 11) %>% as.numeric()

  commuting_data_by_zip <- read.csv(paste('Data/Commuting Data/By ZIP Code/', filename, sep=''))  %>%
    rowwise() %>%
    mutate(LabelGroupings = parse_label_groupings(`Label..Grouping.`),
           LabelGroupings = str_replace_all(LabelGroupings, ':', ''),
           Year = local(year_of_data)) %>%
    separate(LabelGroupings, sep=';', into=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3', 'LabelGroup4')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3), ~na_if(.,'')) %>%
    mutate_at(vars(LabelGroup1, LabelGroup2, LabelGroup3, LabelGroup4), ~toupper(.)) %>%
    fill(LabelGroup1, LabelGroup2, LabelGroup3) %>%
    filter(LabelGroup4 == 'NONE', LabelGroup2 != 'NONE') %>%
    select(-`Label..Grouping.`, -LabelGroup4) %>%
    select(Year, LabelGroup1, LabelGroup2, LabelGroup3, everything()) %>% 
    pivot_longer(cols=starts_with('ZCTA'), names_to = 'ZIP', values_to = 'Population', values_transform = as.character) %>%
    filter(grepl('..Estimate', ZIP)) %>%
    mutate(ZIP = substring(ZIP, 7, 11),
           Population = str_remove(Population, ',') %>% as.numeric()) %>%
    inner_join(label_groupings_summarised, by=c('LabelGroup1', 'LabelGroup2', 'LabelGroup3'))
  
  return(commuting_data_by_zip)
}
```

```{r}
commuting_by_zip <- data.frame()

for (filename in commuter_filenames) {
  commuting_by_zip <- rbind(commuting_by_zip, parse_commuting_data_by_zip_code(filename))
}

# checking the completeness of the data; ZIP code 90090 is missing data for 2021-2024, but these years are simply not available in the census data
for (year in c(2013:2024)) {
  commuting_by_zip %>% 
    # group_by(ZIP) %>% summarise(Count = n()) %>% arrange(desc(Count))
    group_by(ZIP) %>% mutate(Count = n()) %>% filter(Count < 264, Year == local(year)) %>% ungroup() %>%
    arrange(ZIP, LabelGroup1, LabelGroup2, LabelGroup3) %>%
    print()
}

commuting_by_zip %>% write.csv('Cleaned Data/LACommutingDataByZIP.csv', row.names=FALSE)
```

```{r}
commuting_by_zip <- read.csv('Cleaned Data/LACommutingDataByZIP.csv')

commuting_by_zip_pivoted <- commuting_by_zip %>%
  mutate(LabelGroup = case_when(
    LabelGroup3 == 'DROVE ALONE' ~ 'Driving',
    LabelGroup3 == 'CARPOOLED' ~ 'Carpooling',
    LabelGroup2 %in% c('PUBLIC TRANSPORTATION (EXCLUDING TAXICAB)',
                       'PUBLIC TRANSPORTATION') ~ 'PublicTransit',
    LabelGroup2 == 'WALKED' ~ 'Walking',
    LabelGroup2 %in% c('WORKED AT HOME', 'WORKED FROM HOME') ~ 'WorkingFromHome',
    LabelGroup2 %in% c('BICYCLE', 'MOTORCYCLE', 'TAXICAB', 
                       'TAXI OR RIDE-HAILING SERVICES', 'OTHER MEANS') ~ 'TaxiBikeAndOther'
  )) %>% 
  filter(!is.na(LabelGroup)) %>%
  select(Year, ZIP, LabelGroup, Population) %>%
  group_by(Year, ZIP, LabelGroup) %>% summarise(Population = sum(Population)) %>% 
  pivot_wider(names_from=LabelGroup, values_from=Population)
```

## Observing the census and commuter data
```{r}
colnames(census_by_zip_pivoted)

# sex
print(
  census_by_zip %>%
    filter(LabelGroup1 == 'SEX AND AGE', LabelGroup2 %in% c('MALE', 'FEMALE')) %>%
    group_by(Sex = LabelGroup2, Year) %>%
    summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=Sex)) +
    geom_line() +
    ylim(0, NA))

# age

# race/ethnicity
```

## Comparing the census and commuter data to collisions data
```{r}
set.seed(55)

zip_codes_all_metrics <- collisions_with_labeled_intersections %>%
  rename(Year = OccurredYear) %>%
  left_join(
    all_coordinates %>%
      filter(CoordinateType == 'Collision') %>%
      rowwise() %>%
      mutate(ZIPCode = ifelse(Backfilled != Frontfilled, 
                              sample(c(Backfilled, Frontfilled), 1), Backfilled)) %>%
      select(Latitude, Longitude, ZIPCode), by=c('Latitude', 'Longitude')) %>%
  left_join(census_by_zip_pivoted %>% rename(ZIPCode = ZIP), by=c('ZIPCode', 'Year')) %>%
  left_join(commuting_by_zip_pivoted %>% rename(ZIPCode = ZIP), by=c('ZIPCode', 'Year'))
```




