---
title: "Traffic Collisions Location Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
library(Morpho)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Census data
```{r}
parse_census_data <- function(year) {
  
  census_overall <- read.csv(paste('Data/Census Data/demographics_',
                                 year, '.csv', sep='')) %>%
      mutate(Year = local(year),
             Demographic = str_trim(`Label..Grouping.`),
             Demographic = case_when(
               Demographic %in% c('White Alone') ~ 'White',
               Demographic %in% c('Black or African American alone') ~ 'Black',
               Demographic %in% c('Asian alone') ~ 'Asian',
               Demographic %in% c('Some Other Race alone') ~ 'Other',
               Demographic %in% c('Two or More Races') ~ 'Two or More Races',
               Demographic %in% c('Hispanic or Latino (of any race)') ~ 'Hispanic',
               TRUE ~ Demographic),
             DemographicType = case_when(
               Demographic %in% c('Male', 'Female') ~ 'Sex',
               Demographic %in% c('Under 5 years', '5 to 9 years', '10 to 14 years', 
                                  '15 to 19 years', '20 to 24 years', '25 to 34 years', 
                                  '35 to 44 years', '45 to 54 years', '55 to 59 years', 
                                  '60 to 64 years', '65 to 74 years', '75 to 84 years', 
                                  '85 years and over') ~ 'Age',
               Demographic %in% c('White', 'Black', 'Asian', 'Other', 
                                  'Two or More Races', 'Hispanic') ~ 'RaceEthnicity',
               TRUE ~ NA),
             Population = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ',')),
             TotalPopulation = ifelse(Demographic %in% c('Total population'), Population, NA)) %>%
      group_by(Year) %>% mutate(TotalPopulation = max(TotalPopulation, na.rm=TRUE)) %>% ungroup() %>%
      filter(!is.na(DemographicType)) %>%
      select(Year, DemographicType, Demographic, Population, TotalPopulation)

  return(census_overall)
}

census_data <- data.frame()

for (year in c(2013:2024)) {census_data <- rbind(census_data, parse_census_data(year))}
```

## Age trends over time
```{r}
print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    ggplot(aes(x=Year, y=Population, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group')
)

print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    mutate(Proportion = Population/Total) %>%
    ggplot(aes(x=Year, y=Proportion, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group') +
    scale_y_continuous(labels=scales::percent)
)
```

## Breakdown of collisions by race/ethnicity
```{r, fig.height=4, fig.width=6}
for (race in c('Asian', 'Black American', 'Hispanic/Latino', 'White')) {
  print(
    collisions_with_labeled_intersections %>%
      filter(Latitude != 0, Longitude != 0, VictimEthnicity == local(race)) %>%
      group_by(Latitude = round(Latitude, 3), Longitude = round(Longitude, 3), VictimEthnicity) %>%
      summarise(Collisions = n()) %>%
      arrange(Collisions) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point() +
      xlim(-118.7, -118.1) +
      ylim(33.6, 34.4) +
      facet_wrap(~VictimEthnicity)
  )
}
```

## Collision rates over time
```{r}
print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    mutate(CollisionRate = Collisions/TotalPopulation) %>%
    ggplot(aes(x=Year, y=CollisionRate)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA))
)

print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    left_join(commuting_data_overall, by='Year') %>%
    mutate(CollisionRate = Collisions/TotalPopulation,
           CarCommutersByPopulation = (Driving + Carpooling)/TotalPopulation
           ) %>%
    ggplot(aes(x=CarCommutersByPopulation, y=CollisionRate, color=as.factor(Year))) +
    geom_point() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA)) +
    scale_x_continuous(labels=scales::percent) +
    labs(x='Number of Car Commuters per Population', y='Collisions per Population', color='Year')
)
```

# Commutes over Time
## Loading and cleaning data
```{r}
pivot_commuting_overall <- function(year) {
  commuting_overall <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                               year, '.csv', sep='')) %>%
    mutate(Year = local(year),
           Labels = str_trim(`Label..Grouping.`),
           Labels = case_when(
             Labels == 'Taxi or ride-hailing services' ~ 'Taxicab', 
             Labels == 'Public transportation (excluding taxicab):' ~ 'Public transportation:',
             Labels == 'Worked at home' ~ 'Worked from home',
             TRUE ~ Labels),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, Labels, Values) %>%
    filter(Labels %in% c('Total:', 'Drove alone', 'Carpooled:', 'Public transportation:', 
                         'Walked', 'Taxicab', 'Motorcycle', 'Bicycle', 'Other means', 'Worked from home')) %>%
    pivot_wider(names_from=Labels, values_from=Values) %>%
    mutate(TaxiBikeAndOther = Taxicab + Motorcycle + Bicycle + `Other means`) %>% 
    select(Year, `Total:`, `Drove alone`, `Carpooled:`, `Public transportation:`, 
           Walked, TaxiBikeAndOther, `Worked from home`)
  
  colnames(commuting_overall) <- c('Year', 'TotalWorkers', 'Driving', 'Carpooling', 'PublicTransit', 
                                   'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(commuting_overall)
}


pivot_commuting_data_by_ethn <- function(year, file_subname, race_ethnicity_label) {
  
  ethn_df <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                            file_subname, '_', year, '.csv', sep='')) %>% 
    mutate(RaceEthnicity = local(race_ethnicity_label), Year = local(year),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, RaceEthnicity, `Label..Grouping.`, Values) %>%
    pivot_wider(names_from=`Label..Grouping.`, values_from=Values)
  
  colnames(ethn_df) <- c('Year', 'RaceEthnicity', 'TotalWorkers', 'Driving', 'Carpooling',
                         'PublicTransit', 'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(ethn_df)
}

parse_commute_data <- function(years) {
  overall <- data.frame()
  by_ethn <- data.frame()
  by_lang <- data.frame()
  
  for (year in years) {
    overall <- rbind(overall, pivot_commuting_overall(year))
    by_ethn <- rbind(by_ethn, pivot_commuting_data_by_ethn(year, 'asian', 'Asian') %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'black', 'Black')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'hispanic', 'Hispanic')) %>%
                       # rbind(pivot_commuting_data_by_ethn(year, 'white', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'white_nonhispanic', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'multiple_races', 'Two or More Races')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'other', 'Other')))
    # by_lang <- rbind(by_lang, read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_by_language_', year, '.csv', sep='')))
  }
  
  return(list(overall, by_ethn, by_lang))
}

commuting_data <- parse_commute_data(c(2013:2024))
commuting_data_overall <- commuting_data[1] %>% as.data.frame()
commuting_data_by_ethnicity <- commuting_data[2] %>% as.data.frame()
```

## Commuting trends over time
```{r}
# Proportion of each mode of transportation over time
print(
  commuting_data_overall %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)

print(
  commuting_data_by_ethnicity %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    facet_wrap(~RaceEthnicity) +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)
```

## Commuting shifts after COVID
```{r}
commuting_by_covid_period <- commuting_data_by_ethnicity %>%
  filter(Year >= 2016, Year != 2020) %>%
  mutate(Period = ifelse(Year < 2020, 'Pre-COVID', 'Post-COVID') %>%
           factor(levels = c('Pre-COVID', 'Post-COVID'))) %>%
  group_by(Period, RaceEthnicity) %>%
  summarise(
    AvgWorkersPerYear = mean(TotalWorkers) %>% round(),
    AvgCommutingByCar = mean(Driving + Carpooling) %>% round(),
    AvgCommutingByPublicTransit = mean(PublicTransit) %>% round(),
    AvgCommutingByWalking = mean(Walking) %>% round(),
    AvgCommutingByTaxiBikeAndOther = mean(TaxiBikeAndOther) %>% round(),
    AvgWorkFromHome = mean(WorkingFromHome) %>% round(),
    
    CommutingByCarRate = round(AvgCommutingByCar/AvgWorkersPerYear, 2),
    CommutingByPublicTransitRate = round(AvgCommutingByPublicTransit/AvgWorkersPerYear, 2),
    CommutingByWalkingRate = round(AvgCommutingByWalking/AvgWorkersPerYear, 2),
    CommutingByTaxiBikeAndOtherRate = round(AvgCommutingByTaxiBikeAndOther/AvgWorkersPerYear, 2),
    WFHRate = round(AvgWorkFromHome/AvgWorkersPerYear, 2),
  )

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=CommutingByCarRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Commuting by Car Rate', fill='') +
    ggtitle('Proportion of Workers Commuting by Car, Pre- and Post-COVID')
)

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=WFHRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Working from Home Rate', fill='') +
    ggtitle('Proportion of Workers Working from Home, Pre- and Post-COVID')
)
```

## Comparing number of collision victims by ethnicity, compared to WFH rate
Too correlated w/ decrease in collisions after COVID
```{r}
collisions_by_ethnicity <- collisions_with_labeled_intersections %>%
  mutate(
    VictimEthnicity = case_when(
      VictimDescent == 'H' ~ 'Hispanic',
      VictimDescent %in% c('A', 'C', 'D', 'F', 'G', 'J', 'K', 'L', 'V', 'Z') ~ 'Asian',
      VictimDescent %in% c('P', 'S', 'U') ~ 'Pacific Islander',
      VictimDescent == 'B' ~ 'Black',
      VictimDescent == 'W' ~ 'White',
      VictimDescent %in% c('O', 'X', '', '-') ~ 'Other',
      VictimDescent == 'I' ~ 'Native American/Alaskan',
      TRUE ~ '?'    )
  ) %>%
  group_by(OccurredYear, VictimEthnicity) %>%
  summarise(Collisions = n()) %>%
  right_join(commuting_data_by_ethnicity, 
             by=c('VictimEthnicity'='RaceEthnicity', 'OccurredYear'='Year')) %>%
  right_join(census_data %>% filter(DemographicType == 'RaceEthnicity'),
             by=c('VictimEthnicity'='Demographic', 'OccurredYear'='Year')) %>%
  mutate(
    Car = Driving + Carpooling,
    DrivingProportion = Car/TotalWorkers,
    WalkingProportion = Walking/TotalWorkers,
    PublicTransitProportion = PublicTransit/TotalWorkers,
    WFHProportion = WorkingFromHome/TotalWorkers,
    CollisionRate = Collisions/Population)
```

```{r}
print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=DrivingProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Driving to Work vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)

print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=WFHProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Working from Home vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)
```

## Distributions of collision time over the years
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(!is.na(OccurredTime), OccurredYear >= 2016, 
           OccurredYear < 2025, OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID'),) %>%
    group_by(Period) %>% mutate(Total = n()) %>% ungroup() %>%
    group_by(OccurredTime = round(OccurredTime, -2), Period) %>%
    summarise(Distribution = n()/first(Total)) %>%
    ggplot(aes(x=OccurredTime, y=Distribution, color=Period)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Occurrence Time, Pre- and Post-COVID')
)
```

# Parsing ZIP Codes
```{r}
zip_code_boundaries <- read_json('Data/ZIP Codes/Zipcodes.geojson')

# zip_codes_df <- data.frame()

# for (i in c(1:length(zip_code_boundaries$features))) {
#   row <- zip_code_boundaries$features[[i]]
# 
#   object_id <- row$properties$OBJECTID
#   zip_code <- row$properties$ZIPCODE
#   shape_type <- row$geometry$type
# 
#   if (shape_type == 'Polygon') {
#     coordinates <- row$geometry$coordinates[[1]]
#   } else {
#     coordinates <- row$geometry$coordinates[[1]][[1]]
#   }
#   for (coord_set in coordinates) {
#     zip_codes_df <- rbind(zip_codes_df,
#                           list('ObjectID'=object_id, 'ZIPCode'=zip_code,
#                                'Longitude'=coord_set[[1]],
#                                'Latitude'=coord_set[[2]]))
#   }
# }
# 
# zip_codes_df <- zip_codes_df %>% distinct()
# zip_codes_df %>% group_by(ObjectID) %>% summarise(n())

# zip_codes_df %>% write.csv('Cleaned Data/LAZipCodes.csv', row.names=FALSE)
zip_codes_df <- read.csv('Cleaned Data/LAZipCodes.csv')
```

```{r, fig.height=6, fig.width=4}
zip_codes_centralized <- zip_codes_df %>%
  group_by(ZIPCode) %>% 
  mutate(CenterLong = mean(Longitude, na.rm=TRUE), 
         CenterLat = mean(Latitude, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(
    AdjustedLongitude = (Longitude - CenterLong)*0.95 + CenterLong,
    AdjustedLatitude = (Latitude - CenterLat)*0.95 + CenterLat
  )

print(
  zip_codes_centralized %>%
    ggplot(aes(x=AdjustedLongitude, y=AdjustedLatitude, color=(ZIPCode))) +
    geom_point(size=0.1) + labs(x='Longitude', y='Latitude', color='') +
    theme(legend.position='none')
)
```

## Labeling each collision coordinate location by ZIP code
```{r}
collision_coordinates <- collisions_with_labeled_intersections %>%
  filter(Latitude != 0, Longitude != 0) %>%
  group_by(Latitude, Longitude) %>% summarise(Collisions = n())

identify_zip_code <- function(collision_latitude, collision_longitude) {
  possible_zip_codes <- zip_codes_centralized %>%
    mutate(CollisionLatitude = (collision_latitude),
           CollisionLongitude = (collision_longitude)) %>%
    filter(abs(Latitude - CollisionLatitude) < 0.1,
           abs(Longitude - CollisionLongitude) < 0.1) %>%
    mutate(
      NorthSouthOfCenter = case_when(
        Latitude > CenterLat ~ 'North',
        Latitude < CenterLat ~ 'South',
        TRUE ~ 'Neither'),
      EastWestOfCenter = case_when(
        Longitude > CenterLong ~ 'East',
        Longitude < CenterLong ~ 'West',
        TRUE ~ 'Neither'
      ))
           
  possible_zip_codes <- possible_zip_codes %>%
    mutate(
      CollisionLatitudeBetween = case_when(
        NorthSouthOfCenter == 'North' & 
          CollisionLatitude > CenterLat & 
          CollisionLatitude < Latitude ~ 1,
        NorthSouthOfCenter == 'South' &
          CollisionLatitude < CenterLat &
          CollisionLatitude > Latitude ~ 1,
        NorthSouthOfCenter == 'Neither' &
          CollisionLatitude == CenterLat ~ 1,
        TRUE ~ 0
      )
    ) %>%
    filter(CollisionLatitudeBetween == 1) %>%
    mutate(
      CollisionLongitudeBetween = case_when(
        EastWestOfCenter == 'East' & 
          CollisionLongitude > CenterLong & 
          CollisionLongitude < Longitude ~ 1,
        EastWestOfCenter == 'West' &
          CollisionLongitude < CenterLong &
          CollisionLongitude > Longitude ~ 1,
        EastWestOfCenter == 'Neither' &
          CollisionLongitude == CenterLong ~ 1,
        TRUE ~ 0
      )
    ) %>%
    filter(CollisionLongitudeBetween == 1) %>%
    group_by(ZIPCode) %>% summarise(Count = n()) %>% arrange(desc(Count))
  
  return(possible_zip_codes[1, 'ZIPCode'][[1]])
}

collision_coordinates <- collision_coordinates %>%
  rowwise() %>% mutate(ZIPCode = identify_zip_code(Latitude, Longitude))

collision_coordinates %>% write.csv('Cleaned Data/CollisionCoordinatesByZIPCode.csv', row.names=FALSE)
```

### Validating this labeling method
Works better on convex shapes, but there are a lot of oddly shaped ZIP codes this method doesn't quite work for
```{r, fig.height=6, fig.width=4}
collision_coordinates %>% filter(is.na(ZIPCode))
collision_coordinates %>% group_by(ZIPCode) %>% summarise(Count=n()) %>% arrange(desc(Count))

collision_zip_codes <- collision_coordinates %>% ungroup() %>% select(ZIPCode) %>% distinct() %>% drop_na()

all_zip_codes <- collision_coordinates %>% select(ZIPCode, Longitude, Latitude) %>% 
  ungroup() %>% mutate(Type = 'collision') %>%
  rbind(zip_codes_centralized %>% select(ZIPCode, Longitude, Latitude) %>% 
          mutate(Type = 'Boundary'))

for (zip_code in collision_zip_codes$ZIPCode) {
  print(
    all_zip_codes %>%
      filter(ZIPCode == local(zip_code)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Type)) +
      geom_point(size=1) + labs(x='Longitude', y='Latitude', color='') +
      theme(legend.position='none') 
  )
}
```


## Interpolating the ZIP code boundary coordinates
```{r}
zip_codes_indexed <- zip_codes_centralized %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  group_by(ZIPCode) %>% 
  mutate(Index =  row_number()) %>% 
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2))
```

```{r}
print(
  zip_codes_indexed %>%
    ggplot(aes(x=DistanceToNext)) +
    geom_histogram(binwidth = 0.001)
)

zip_codes_indexed %>% arrange(desc(DistanceToNext))
zip_codes_indexed %>% group_by(DistanceToNext = round(DistanceToNext, 4)) %>%
  summarise(Count = n()) %>% arrange(desc(DistanceToNext))
```

```{r}

interpolate_coordinates <- function(distance_to_next, boundary_latitude, boundary_longitude, 
                                    next_boundary_latitude, next_boundary_longitude, interpol_distance) {
    num_steps <- floor(distance_to_next/interpol_distance)
    latitude_diff <- next_boundary_latitude - boundary_latitude
    longitude_diff <- next_boundary_longitude - boundary_longitude
    
    new_latitudes <- ''
    new_longitudes <- ''
    
    for (i in c(1:num_steps)) {
      new_latitudes <- paste(new_latitudes, ', ', boundary_latitude + latitude_diff*i/(1+num_steps))
      new_longitudes <- paste(new_longitudes, ', ', boundary_longitude + longitude_diff*i/(1+num_steps))
    }
    
    return(paste(new_latitudes, new_longitudes, sep=';'))
  }

start_time <- Sys.time()
interpolation_distance <- 0.00005

zip_codes_interpolated <- zip_codes_indexed %>%
  rowwise() %>%
  mutate(
    InterpolatedCoordinates = ifelse(
      DistanceToNext > local(interpolation_distance) & (DistanceToNext < 0.0095 | ZIPCode != '90803'), # 90803 has some non-contiguous areas 
      interpolate_coordinates(DistanceToNext, Latitude, Longitude,
                              NextLatitude, NextLongitude, local(interpolation_distance)), NA)
  ) %>%
  separate(InterpolatedCoordinates, sep=';', 
           into=c('InterpolatedLatitude', 'InterpolatedLongitude')) %>%
  separate_longer_delim(c(InterpolatedLatitude, InterpolatedLongitude), delim=', ') %>%
  mutate(
    InterpolatedLongitude = as.numeric(InterpolatedLongitude),
    InterpolatedLatitude = as.numeric(InterpolatedLatitude),
    Longitude = ifelse(is.na(InterpolatedLongitude), Longitude, InterpolatedLongitude),
    Latitude = ifelse(is.na(InterpolatedLatitude), Latitude, InterpolatedLatitude),
  ) %>% select(-InterpolatedLatitude, -InterpolatedLongitude)

zip_codes_reindexed <- zip_codes_interpolated %>%
  select(ObjectID, ZIPCode, Longitude, Latitude) %>%
  mutate(Latitude = round(Latitude, 4), Longitude = round(Longitude, 4)) %>%
  distinct() %>%
  group_by(ZIPCode) %>%
  mutate(Index = row_number()) %>%
  ungroup() %>%
  (function(df) df %>% left_join(df %>% group_by(ZIPCode) %>% mutate(LastIndex = max(Index)) %>%
                                   mutate(Index = ifelse(Index == 1, LastIndex, Index - 1), 
                                          NextLatitude = Latitude, NextLongitude = Longitude) %>% 
                                   select(ZIPCode, Index, NextLatitude, NextLongitude), by=c('ZIPCode', 'Index'))) %>%
  mutate(DistanceToNext = sqrt((NextLatitude - Latitude)**2  + (NextLongitude - Longitude)**2 ))

print(Sys.time() - start_time)

zip_codes_reindexed %>% filter(DistanceToNext > 0.0001) # 0.0001414214
zip_codes_reindexed %>% mutate(LatitudeDiff = NextLatitude - Latitude) %>% filter(LatitudeDiff > 0.0001)
zip_codes_reindexed %>% mutate(LongitudeDiff = NextLongitude - Longitude) %>% filter(LongitudeDiff > 0.0001)

# zip_codes_reindexed %>% write.csv('Cleaned Data/LAZipCodesInterpolated.csv', row.names=FALSE)
# zip_codes_reindexed <- read.csv('Cleaned Data/LAZipCodesInterpolated.csv')
```

## Second attempt at labeling each collision coordinate location by ZIP code
```{r}
identify_zip_code_by_axis <- function(collision_latitude, collision_longitude) {
  rounded_latitude <- round(collision_latitude, 4)
  rounded_longitude <- round(collision_longitude, 4)
  
  latitude_rounding_offset <- collision_latitude - rounded_latitude
  longitude_rounding_offset <- collision_longitude - rounded_longitude
  
  possible_zip_codes <- zip_codes_reindexed %>%
    mutate(RoundedLatitude = round(Latitude - latitude_rounding_offset, 4) ,
           RoundedLongitude = round(Longitude - latitude_rounding_offset, 4)) %>%
    filter(RoundedLatitude == local(rounded_latitude) |
           RoundedLongitude == local(rounded_longitude))
  
  likely_zip_code <- possible_zip_codes %>% 
    filter(RoundedLatitude == local(rounded_latitude),
           Longitude < collision_longitude) %>% select(ZIPCode) %>% distinct() %>%
    rbind(possible_zip_codes %>%
            filter(RoundedLatitude == local(rounded_latitude),
                   Longitude > collision_longitude) %>% 
            select(ZIPCode) %>% distinct()) %>%
    rbind(possible_zip_codes %>% 
            filter(RoundedLongitude == local(rounded_longitude),
                   Latitude > collision_latitude) %>% 
            select(ZIPCode) %>% distinct()) %>%
    rbind(possible_zip_codes %>% 
            filter(RoundedLongitude == local(rounded_longitude), 
                   Latitude < collision_latitude) %>% 
            select(ZIPCode) %>% distinct()) %>%
    group_by(ZIPCode) %>% summarise(Count = n()) %>% 
    filter(Count == 4) %>%
    arrange(desc(Count))
  
  return(likely_zip_code[1, 'ZIPCode'][[1]])
} 

collision_coordinates <- collision_coordinates %>%
  rowwise() %>% mutate(ZIPCode = identify_zip_code_by_axis(Latitude, Longitude))

collision_coordinates %>% write.csv('Cleaned Data/CollisionCoordinatesByZIPCode.csv', row.names=FALSE)
collision_coordinates <- read.csv('Cleaned Data/CollisionCoordinatesByZIPCode.csv')
```

```{r}
collision_coordinates %>% filter(is.na(ZIPCode))
collision_coordinates %>% group_by(ZIPCode) %>% summarise(Count=n()) %>% arrange(desc(Count))

collision_zip_codes <- collision_coordinates %>% ungroup() %>% select(ZIPCode) %>% distinct() %>% drop_na()

all_zip_codes <- collision_coordinates %>% select(ZIPCode, Longitude, Latitude) %>% 
  ungroup() %>% mutate(Type = 'collision') %>%
  rbind(zip_codes_reindexed %>% select(ZIPCode, Longitude, Latitude) %>% 
          mutate(Type = 'Boundary'))

for (zip_code in collision_zip_codes$ZIPCode) {
# for (zip_code in c('90732', '90275')) {
  print(
    all_zip_codes %>%
      filter(ZIPCode == local(zip_code)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Type)) +
      geom_point(size=1) + labs(x='Longitude', y='Latitude', color='') +
      theme(legend.position='none') 
  )
}
``` 

## Attempting another method to label the collision coordinates by ZIP code
```{r}
# find the angle of each boundary point walk, then find the differences between each angle
# if the angles tend to increase, the walk is counterclockwise
zip_codes_by_rotation <- zip_codes_reindexed %>% # filter(ZIPCode == '90001') %>%
  mutate(
    LatitudeDiff = NextLatitude - Latitude,
    LongitudeDiff = NextLongitude - Longitude,
    Slope = (LatitudeDiff/LongitudeDiff)*DistanceToNext
  ) %>%
  group_by(ZIPCode) %>% 
  summarise(
    TotalSlope = sum(Slope, na.rm=TRUE),
    TotalLatitudeChange = sum(LatitudeDiff, na.rm=TRUE),
    TotalLongitudeChange = sum(LongitudeDiff, na.rm=TRUE),
  )

test_df <- data.frame(c('Latitude' = c(1, 3, 5, 4, 6),
                        'Longitude' = c(1, -2, -1, 3, )))
```

```{r, fig.height=6, fig.width=4}

identify_zip_code_by_nearest_boundaries <- function(collision_latitude, collision_longitude) {
  
  possible_zip_codes <- zip_codes_recentralized %>%
    filter(abs(AdjustedLatitude - collision_latitude) < 0.001,
           abs(AdjustedLongitude - collision_longitude) < 0.001)
  
  # likely_zip_code <- possible_zip_codes %>%
  #   mutate(LatitudeDiff = collision_latitude - AdjustedLatitude,
  #          LongitudeDiff = collision_longitude - AdjustedLongitude,
  #          Distance = sqrt(LatitudeDiff**2 + LongitudeDiff**2)) %>%
  #   # arrange(Distance) %>% group_by(ZIPCode) %>% mutate(Rank = row_number()) %>% ungroup() %>% filter(Rank <= 5) %>% 
  #   group_by(ZIPCode) %>% 
  #   summarise(AvgDistance = mean(Distance)) %>%
  #   arrange((AvgDistance))
  
  # return(likely_zip_code[1, 'ZIPCode'][[1]])
  return(nrow(possible_zip_codes))
}

boundary_adjustment <- 0.00001

zip_codes_recentralized <- zip_codes_reindexed %>%
  group_by(ZIPCode) %>% 
  mutate(MaxLatitude = max(Latitude), MinLatitude = min(Latitude),
         MaxLongitude = max(Longitude), MinLongitude = min(Longitude),
         CenterLat = (MaxLatitude + MinLatitude)/2,
         CenterLong = (MaxLongitude + MinLongitude)/2) %>%
  ungroup() %>%
  mutate(
    # AdjustedLatitude = (Latitude - CenterLat)*local(boundary_adjustment) + CenterLat,
    # AdjustedLongitude = (Longitude - CenterLong)*local(boundary_adjustment) + CenterLong,
    
    AdjustedLatitude = ifelse(Latitude > CenterLat, Latitude - local(boundary_adjustment), 
                              Latitude + local(boundary_adjustment)),
    AdjustedLongitude = ifelse(Longitude > CenterLong, Longitude - local(boundary_adjustment), 
                              Longitude + local(boundary_adjustment)),
  )


print(
  zip_codes_recentralized %>% 
    ggplot(aes(x=AdjustedLongitude, y=AdjustedLatitude, color=ZIPCode)) +
    geom_point(size=0.1) +
    theme(legend.position='none')
)

collision_coordinates_by_nearest_boundaries <- collision_coordinates %>%
  rowwise() %>% mutate(ZIPCode = identify_zip_code_by_nearest_boundaries(Latitude, Longitude))

collision_coordinates_by_nearest_boundaries %>% arrange(ZIPCode)
```

```{r}
collision_coordinates_by_nearest_boundaries %>% filter(is.na(ZIPCode))
collision_coordinates_by_nearest_boundaries %>% group_by(ZIPCode) %>% summarise(Count=n()) %>% arrange(desc(Count))

collision_zip_codes_by_nearest_boundaries <- collision_coordinates_by_nearest_boundaries %>% 
  ungroup() %>% select(ZIPCode) %>% distinct() %>% drop_na()

all_zip_codes_by_nearest_boundary <- collision_zip_codes_by_nearest_boundaries %>% 
  select(ZIPCode, Longitude, Latitude) %>% 
  ungroup() %>% mutate(Type = 'collision') %>%
  rbind(zip_codes_reindexed %>% select(ZIPCode, Longitude, Latitude) %>% 
          mutate(Type = 'Boundary'))

for (zip_code in (all_zip_codes_by_nearest_boundary %>% select(ZIPCode) %>% distinct(ZIPCode))$ZIPCode) {
# for (zip_code in c('90732', '90275')) {
  print(
    all_zip_codes_by_nearest_boundary %>%
      filter(ZIPCode == local(zip_code)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Type)) +
      geom_point(size=1) + labs(x='Longitude', y='Latitude', color='') +
      theme(legend.position='none') 
  )
}
```

## Labeling the collision coordinates ZIP codes using a Latitude-wise method
```{r}
all_coordinates <- zip_codes_reindexed %>% mutate(CoordinateType = 'Boundary') %>%
  mutate(Latitude = round(Latitude, 4), Longitude = round(Longitude, 4),
         LateralDirection = case_when(
           NextLatitude > Latitude ~ 1,
           NextLatitude < Latitude ~ -1,
           TRUE ~ 0),
         Longitude = case_when(
           LateralDirection == 1 ~ Longitude - 0.00001,
           LateralDirection == -1 ~ Longitude + 0.00001,
           TRUE ~ Longitude)) %>%
  select(Latitude, Longitude, ZIPCode, CoordinateType) %>%
  rbind(collisions_with_labeled_intersections %>% 
          filter(Latitude != 0, Longitude != 0) %>% 
          mutate(ZIPCode = NA, CoordinateType = 'Collision') %>% 
          select(Latitude, Longitude, ZIPCode, CoordinateType) %>% distinct()) %>%
  arrange(Latitude, Longitude, ZIPCode) %>%
  mutate(IsCollisionPoint = ifelse(CoordinateType == 'Collision', 1, 0)) %>%
  mutate(Backfilled = ZIPCode, ForwardFilled = ZIPCode) %>%
  fill(Backfilled, .direction='down') %>%
  fill(ForwardFilled, .direction='up') 
  

# nudge each pair of boundary coordinates from the same ZIP code closer to each other
all_coordinates %>% group_by(Latitude) %>% mutate(FirstLongitude = min(Longitude)) %>% ungroup() %>%
  filter(Longitude == FirstLongitude) %>% group_by(Latitude, Longitude) %>% summarise(NumZips = n_distinct(ZIPCode)) %>% arrange(desc(NumZips))
```

```{r}
all_coordinates %>% mutate(Backfilled = ZIPCode, ForwardFilled = ZIPCode) %>%
  fill(Backfilled, .direction='down') %>%
  fill(ForwardFilled, .direction='up') %>%
  filter(Backfilled != ForwardFilled)
```

```{r}
relevant_zip_codes <- (all_coordinates %>% group_by(ForwardFilled) %>% 
  summarise(NumCollisionPoints = sum(IsCollisionPoint)) %>%
  filter(NumCollisionPoints > 0) %>% ungroup() %>%
  select(ForwardFilled))$ForwardFilled

for (zip_code in relevant_zip_codes) {
  print(
    all_coordinates %>% filter(ForwardFilled == local(zip_code)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=CoordinateType)) +
      geom_point(size=0.2) +
      ggtitle(local(zip_code))
  )
}

```
```{r}
all_coordinates %>% filter(ForwardFilled == '90007') 
all_coordinates %>% filter(Latitude == 33.9581) %>% arrange(Longitude)
```
