---
title: "Traffic Collisions Location Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Census data
```{r}
parse_census_data <- function(year) {
  
  census_overall <- read.csv(paste('Data/Census Data/demographics_',
                                 year, '.csv', sep='')) %>%
      mutate(Year = local(year),
             Demographic = str_trim(`Label..Grouping.`),
             Demographic = case_when(
               Demographic %in% c('White Alone') ~ 'White',
               Demographic %in% c('Black or African American alone') ~ 'Black',
               Demographic %in% c('Asian alone') ~ 'Asian',
               Demographic %in% c('Some Other Race alone') ~ 'Other',
               Demographic %in% c('Two or More Races') ~ 'Two or More Races',
               Demographic %in% c('Hispanic or Latino (of any race)') ~ 'Hispanic',
               TRUE ~ Demographic),
             DemographicType = case_when(
               Demographic %in% c('Male', 'Female') ~ 'Sex',
               Demographic %in% c('Under 5 years', '5 to 9 years', '10 to 14 years', 
                                  '15 to 19 years', '20 to 24 years', '25 to 34 years', 
                                  '35 to 44 years', '45 to 54 years', '55 to 59 years', 
                                  '60 to 64 years', '65 to 74 years', '75 to 84 years', 
                                  '85 years and over') ~ 'Age',
               Demographic %in% c('White', 'Black', 'Asian', 'Other', 
                                  'Two or More Races', 'Hispanic') ~ 'RaceEthnicity',
               TRUE ~ NA),
             Population = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ',')),
             TotalPopulation = ifelse(Demographic %in% c('Total population'), Population, NA)) %>%
      group_by(Year) %>% mutate(TotalPopulation = max(TotalPopulation, na.rm=TRUE)) %>% ungroup() %>%
      filter(!is.na(DemographicType)) %>%
      select(Year, DemographicType, Demographic, Population, TotalPopulation)

  return(census_overall)
}

census_data <- data.frame()

for (year in c(2013:2024)) {census_data <- rbind(census_data, parse_census_data(year))}
```

## Age trends over time
```{r}
print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    ggplot(aes(x=Year, y=Population, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group')
)

print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    mutate(Proportion = Population/Total) %>%
    ggplot(aes(x=Year, y=Proportion, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group') +
    scale_y_continuous(labels=scales::percent)
)
```

## Breakdown of collisions by race/ethnicity
```{r, fig.height=4, fig.width=6}
for (race in c('Asian', 'Black American', 'Hispanic/Latino', 'White')) {
  print(
    collisions_with_labeled_intersections %>%
      filter(Latitude != 0, Longitude != 0, VictimEthnicity == local(race)) %>%
      group_by(Latitude = round(Latitude, 3), Longitude = round(Longitude, 3), VictimEthnicity) %>%
      summarise(Collisions = n()) %>%
      arrange(Collisions) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point() +
      xlim(-118.7, -118.1) +
      ylim(33.6, 34.4) +
      facet_wrap(~VictimEthnicity)
  )
}
```

## Collision rates over time
```{r}
print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    mutate(CollisionRate = Collisions/TotalPopulation) %>%
    ggplot(aes(x=Year, y=CollisionRate)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA))
)

print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_with_labeled_intersections %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    left_join(commuting_data_overall, by='Year') %>%
    mutate(CollisionRate = Collisions/TotalPopulation,
           CarCommutersByPopulation = (Driving + Carpooling)/TotalPopulation
           ) %>%
    ggplot(aes(x=CarCommutersByPopulation, y=CollisionRate, color=as.factor(Year))) +
    geom_point() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA)) +
    scale_x_continuous(labels=scales::percent) +
    labs(x='Number of Car Commuters per Population', y='Collisions per Population', color='Year')
)
```

# Commutes over Time
## Loading and cleaning data
```{r}
pivot_commuting_overall <- function(year) {
  commuting_overall <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                               year, '.csv', sep='')) %>%
    mutate(Year = local(year),
           Labels = str_trim(`Label..Grouping.`),
           Labels = case_when(
             Labels == 'Taxi or ride-hailing services' ~ 'Taxicab', 
             Labels == 'Public transportation (excluding taxicab):' ~ 'Public transportation:',
             Labels == 'Worked at home' ~ 'Worked from home',
             TRUE ~ Labels),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, Labels, Values) %>%
    filter(Labels %in% c('Total:', 'Drove alone', 'Carpooled:', 'Public transportation:', 
                         'Walked', 'Taxicab', 'Motorcycle', 'Bicycle', 'Other means', 'Worked from home')) %>%
    pivot_wider(names_from=Labels, values_from=Values) %>%
    mutate(TaxiBikeAndOther = Taxicab + Motorcycle + Bicycle + `Other means`) %>% 
    select(Year, `Total:`, `Drove alone`, `Carpooled:`, `Public transportation:`, 
           Walked, TaxiBikeAndOther, `Worked from home`)
  
  colnames(commuting_overall) <- c('Year', 'TotalWorkers', 'Driving', 'Carpooling', 'PublicTransit', 
                                   'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(commuting_overall)
}


pivot_commuting_data_by_ethn <- function(year, file_subname, race_ethnicity_label) {
  
  ethn_df <- read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_', 
                            file_subname, '_', year, '.csv', sep='')) %>% 
    mutate(RaceEthnicity = local(race_ethnicity_label), Year = local(year),
           Values = as.numeric(str_remove_all(`Los.Angeles.city..California..Estimate`, ','))) %>%
    select(Year, RaceEthnicity, `Label..Grouping.`, Values) %>%
    pivot_wider(names_from=`Label..Grouping.`, values_from=Values)
  
  colnames(ethn_df) <- c('Year', 'RaceEthnicity', 'TotalWorkers', 'Driving', 'Carpooling',
                         'PublicTransit', 'Walking', 'TaxiBikeAndOther', 'WorkingFromHome')
  
  return(ethn_df)
}

parse_commute_data <- function(years) {
  overall <- data.frame()
  by_ethn <- data.frame()
  by_lang <- data.frame()
  
  for (year in years) {
    overall <- rbind(overall, pivot_commuting_overall(year))
    by_ethn <- rbind(by_ethn, pivot_commuting_data_by_ethn(year, 'asian', 'Asian') %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'black', 'Black')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'hispanic', 'Hispanic')) %>%
                       # rbind(pivot_commuting_data_by_ethn(year, 'white', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'white_nonhispanic', 'White')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'multiple_races', 'Two or More Races')) %>%
                       rbind(pivot_commuting_data_by_ethn(year, 'other', 'Other')))
    # by_lang <- rbind(by_lang, read.csv(paste('Data/Commuting Data/', year, '/means_of_commuting_by_language_', year, '.csv', sep='')))
  }
  
  return(list(overall, by_ethn, by_lang))
}

commuting_data <- parse_commute_data(c(2013:2024))
commuting_data_overall <- commuting_data[1] %>% as.data.frame()
commuting_data_by_ethnicity <- commuting_data[2] %>% as.data.frame()
```

## Commuting trends over time
```{r}
# Proportion of each mode of transportation over time
print(
  commuting_data_overall %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)

print(
  commuting_data_by_ethnicity %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    facet_wrap(~RaceEthnicity) +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)
```

## Commuting shifts after COVID
```{r}
commuting_by_covid_period <- commuting_data_by_ethnicity %>%
  filter(Year >= 2016, Year != 2020) %>%
  mutate(Period = ifelse(Year < 2020, 'Pre-COVID', 'Post-COVID') %>%
           factor(levels = c('Pre-COVID', 'Post-COVID'))) %>%
  group_by(Period, RaceEthnicity) %>%
  summarise(
    AvgWorkersPerYear = mean(TotalWorkers) %>% round(),
    AvgCommutingByCar = mean(Driving + Carpooling) %>% round(),
    AvgCommutingByPublicTransit = mean(PublicTransit) %>% round(),
    AvgCommutingByWalking = mean(Walking) %>% round(),
    AvgCommutingByTaxiBikeAndOther = mean(TaxiBikeAndOther) %>% round(),
    AvgWorkFromHome = mean(WorkingFromHome) %>% round(),
    
    CommutingByCarRate = round(AvgCommutingByCar/AvgWorkersPerYear, 2),
    CommutingByPublicTransitRate = round(AvgCommutingByPublicTransit/AvgWorkersPerYear, 2),
    CommutingByWalkingRate = round(AvgCommutingByWalking/AvgWorkersPerYear, 2),
    CommutingByTaxiBikeAndOtherRate = round(AvgCommutingByTaxiBikeAndOther/AvgWorkersPerYear, 2),
    WFHRate = round(AvgWorkFromHome/AvgWorkersPerYear, 2),
  )

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=CommutingByCarRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Commuting by Car Rate', fill='') +
    ggtitle('Proportion of Workers Commuting by Car, Pre- and Post-COVID')
)

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=WFHRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Working from Home Rate', fill='') +
    ggtitle('Proportion of Workers Working from Home, Pre- and Post-COVID')
)
```

## Comparing number of collision victims by ethnicity, compared to WFH rate
Too correlated w/ decrease in collisions after COVID
```{r}
collisions_by_ethnicity <- collisions_with_labeled_intersections %>%
  mutate(
    VictimEthnicity = case_when(
      VictimDescent == 'H' ~ 'Hispanic',
      VictimDescent %in% c('A', 'C', 'D', 'F', 'G', 'J', 'K', 'L', 'V', 'Z') ~ 'Asian',
      VictimDescent %in% c('P', 'S', 'U') ~ 'Pacific Islander',
      VictimDescent == 'B' ~ 'Black',
      VictimDescent == 'W' ~ 'White',
      VictimDescent %in% c('O', 'X', '', '-') ~ 'Other',
      VictimDescent == 'I' ~ 'Native American/Alaskan',
      TRUE ~ '?'    )
  ) %>%
  group_by(OccurredYear, VictimEthnicity) %>%
  summarise(Collisions = n()) %>%
  right_join(commuting_data_by_ethnicity, 
             by=c('VictimEthnicity'='RaceEthnicity', 'OccurredYear'='Year')) %>%
  right_join(census_data %>% filter(DemographicType == 'RaceEthnicity'),
             by=c('VictimEthnicity'='Demographic', 'OccurredYear'='Year')) %>%
  mutate(
    Car = Driving + Carpooling,
    DrivingProportion = Car/TotalWorkers,
    WalkingProportion = Walking/TotalWorkers,
    PublicTransitProportion = PublicTransit/TotalWorkers,
    WFHProportion = WorkingFromHome/TotalWorkers,
    CollisionRate = Collisions/Population)
```

```{r}
print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=DrivingProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Driving to Work vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)

print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=WFHProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Working from Home vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)
```

## Distributions of collision time over the years
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(!is.na(OccurredTime), OccurredYear >= 2016, 
           OccurredYear < 2025, OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID'),) %>%
    group_by(Period) %>% mutate(Total = n()) %>% ungroup() %>%
    group_by(OccurredTime = round(OccurredTime, -2), Period) %>%
    summarise(Distribution = n()/first(Total)) %>%
    ggplot(aes(x=OccurredTime, y=Distribution, color=Period)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Occurrence Time, Pre- and Post-COVID')
)
```

# Parsing ZIP Codes
```{r}
zip_code_boundaries <- read_json('Data/ZIP Codes/Zipcodes.geojson')

# zip_codes_df <- read.csv('Cleaned Data/LAZipCodes.csv') %>%
#   filter(ObjectID < 285)
# 
# for (i in range(280:length(zip_code_boundaries$features))) {
#   row <- zip_code_boundaries$features[[i]]
#   
#   object_id <- row$properties$OBJECTID
#   zip_code <- row$properties$ZIPCODE
#   shape_type <- row$geometry$type
#   
#   if (shape_type == 'Polygon') {
#     coordinates <- row$geometry$coordinates[[1]]
#   } else {
#     coordinates <- row$geometry$coordinates[[1]][[1]]
#   }
#   for (coord_set in coordinates) {
#     zip_codes_df <- rbind(zip_codes_df,
#                           list('ObjectID'=object_id, 'ZIPCode'=zip_code,
#                                'Longitude'=coord_set[[1]],
#                                'Latitude'=coord_set[[2]]))
#   }
# }
# 
# zip_codes_df <- zip_codes_df %>% distinct() 
# zip_codes_df %>% write.csv('Cleaned Data/LAZipCodes.csv', row.names=FALSE)
```

```{r}

```
