---
title: "COVID Cases Forecasting"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
```

# Breakdown of Fields
## Categorical Fields
Histograms of OccurredTime, AreaKey/AreaName, ReportingDistrict, CrimeCode/CrimeCodeDescription, VictimAge, VictimSex, VictimDescent, PremiseCode/PremiseDescription, Address, CrossStreet; need to make sure I'm correctly tracking what is or isn't a collision
```{r, fig.height=5, fig.width=7}
for (col in c('OccurredTime', 'AreaKey', 'VictimAge')) {
  print(
    collisions_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram() +
      ggtitle(paste0(col, ' Distribution'))
  )
}

for (col in c('AreaName', 'VictimSex', 'VictimDescent')) {
  print(
    collisions_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram(stat='count') +
      theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust=1)) +
      ggtitle(paste0(col, ' Distribution'))
  )
}

# CrimeCode and CrimeCodeDescription are all the same - Traffic Collision

# No clear concentration of collisions
print(collisions_cleaned %>% group_by(ReportingDistrict) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

# Most common by far is STREET (101)
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

# Each PremiseCode only has one PremiseDescription
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% ungroup() %>% group_by(PremiseCode) %>% 
       summarise(Count=n()) %>% arrange(desc(Count)))

# PremiseDecription for 418 is empty, which comes up once; should replace with NA or vice versa
# Otherwise, each PremiseDescription only has one PremiseCode
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% ungroup() %>% group_by(PremiseDescription) %>% 
       summarise(Count=n()) %>% arrange(desc(Count)))
print(collisions_cleaned %>% group_by(PremiseCode, PremiseDescription) %>%
     summarise(Count = n()) %>% filter(PremiseDescription == ''))

# No clear concentration of collisions
print(collisions_cleaned %>% group_by(Address, CrossStreet) %>%
     summarise(Count = n()) %>% arrange(desc(Count)))

```

## Time series graphs
```{r}
print(collisions_by_reported %>%ggplot(aes(x=ReportedDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Reported Date'))
print(collisions_by_occurred %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Occurred Date'))

print(collisions_by_reported_by_week %>%ggplot(aes(x=ReportedDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Reported Date'))
print(collisions_by_occurred_by_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

### Looking at it by week and by area
```{r}
print(collisions_by_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(collisions_by_area_and_week %>% ggplot(aes(x=OccurredDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```

### Mean and median occurred time over time
```{r}
print(occurred_time_stats_over_time %>% ggplot(aes(x=OccurredDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=OccurredDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```

### Occurred time distribution
```{r}
print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

```{r}
collisions_by_coordinates_and_year %>% 
  filter(OccurredYear == local(year),
         Latitude <= local(34.4),
         Latitude >= local(33.6),
         Longitude <= local(-118.125),
         Longitude >= local(-118.75),
         )
```

## Scatterplot of Latitude, Longitude, 
```{r, fig.width=6, fig.height=3.5}
for (year in c(2011:2025)) {
  print(
    collisions_by_coordinates_and_year %>%
      arrange(TotalCollisions) %>% 
      filter(OccurredYear == local(year),
             Latitude <= local(34.4),
             Latitude >= local(33.6),
             Longitude <= local(-118.125),
             Longitude >= local(-118.75),
             ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=TotalCollisions)) +
      geom_point(size=2) +
      ggtitle(paste0('Total Collisions by Coordinates in ', year, sep=''))
  )
}
```

```{r, fig.width=6, fig.height=3.5}
for (year in c(2011:2025)) {
  print(
    collisions_by_coordinates_and_year %>%
      arrange(AvgCollisionsPerLocation) %>%
      filter(OccurredYear == local(year),
             Latitude <= local(34.4),
             Latitude >= local(33.6),
             Longitude <= local(-118.125),
             Longitude >= local(-118.75),
             ) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=AvgCollisionsPerLocation)) +
      geom_point(size=2) +
      ggtitle(paste0('Avg Collisions Per Location by Coordinates in ', year, sep=''))
  )
}
```

# Forecasting Overall Collision Counts
## Pre-COVID
```{r}
start_time <- c(year(min(collisions_over_time_pre_covid$OccurredDate)), 
                month(min(collisions_over_time_pre_covid$OccurredDate)))

collisions_pre_covid_ts <- ts(collisions_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(collisions_pre_covid_ts, round(length(collisions_pre_covid_ts) * 0.8))
h <- length(collisions_pre_covid_ts) - length(train)
test <- tail(collisions_pre_covid_ts, h)
```

```{r}
collisions_pre_covid_auto_model <- auto.arima(train)
collisions_pre_covid_auto_model
```

### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

### AR
```{r}
pacf(train_dff1) # 3-4
```
### MA
```{r}
acf(train_dff1) #2
```
### ARIMA
```{r}
collisions_pre_covid_model <- Arima(train, order=c(3, 1, 2))
collisions_pre_covid_model
```

### Comparing AIC
```{r}
exp((collisions_pre_covid_auto_model$aic - collisions_pre_covid_model$aic)/2)
```
### Forecast
```{r}
forecast_1 <- forecast(collisions_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(collisions_pre_covid_model, h=length(test))

autoplot(collisions_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

## Post-COVID
```{r}
start_time <- c(year(min(collisions_over_time_post_covid$OccurredDate)), 
                month(min(collisions_over_time_post_covid$OccurredDate)))

collisions_post_covid_ts <- ts(collisions_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(collisions_post_covid_ts, round(length(collisions_post_covid_ts) * 0.8))
h <- length(collisions_post_covid_ts) - length(train)
test <- tail(collisions_post_covid_ts, h)
```

```{r}
collisions_post_covid_auto_model <- auto.arima(train)
collisions_post_covid_auto_model
```

### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

### AR
```{r}
pacf(train) # 1-2
```
### MA
```{r}
acf(train) #3
```
### ARIMA
```{r}
collisions_post_covid_model <- Arima(train, order=c(1, 0, 3))
collisions_post_covid_model
```

### Comparing AIC
```{r}
exp((collisions_post_covid_auto_model$aic - collisions_post_covid_model$aic)/2)
```
### Forecast
```{r}
forecast_1 <- forecast(collisions_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(collisions_post_covid_model, h=length(test))

autoplot(collisions_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

# Identifying Collision Hot Spots
## Check the spread of collision statistics
```{r}
print(
  collisions_by_location_2024 %>%
    ggplot(aes(x=NumCollisionRecords)) +
    geom_histogram()
)
```

```{r}
print(
  collisions_by_location_2024 %>% 
    ggplot(aes(x=RangeOfCollisions, y=StdCollisions, size=NumCollisionRecords)) +
    geom_point()
)
```

```{r}
collisions_by_location_2024 %>% arrange(desc(RangeOfCollisions))
```

```{r}
collisions_by_location_2024 %>% arrange(desc(StdCollisions))
```

```{r}
collisions_cleaned %>%
  mutate(OccurredYear = year(OccurredDate),
         RoundedLat = round(Latitude*local(rounding_factor))/local(rounding_factor),
         RoundedLon = round(Longitude*local(rounding_factor))/local(rounding_factor),
         ) %>%
  filter(
    OccurredYear == 2024,
    (RoundedLat == 33.99 & RoundedLon == -118.28) |
    (RoundedLat == 34.22 & RoundedLon == -118.45) |
    (RoundedLat == 34.03 & RoundedLon == -118.52) |
    (RoundedLat == 34.04 & RoundedLon == -118.56) 
  ) %>%
  group_by(Address, CrossStreet, Latitude, Longitude, RoundedLat, RoundedLon) %>% 
  summarise(Collisions = n()) %>% arrange(desc(RoundedLat), desc(RoundedLon))
```
