---
title: "Traffic Collisions Exploratory Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
library(sgo)
```

# LADOT Data - insufficient
```{r}
ladot_traffic_df <- read.csv('Data/LADOT_Traffic_Counts_Summary.csv')

ladot_traffic_cleaned <- ladot_traffic_df %>% rename(
  PrimaryStreet = `Primary.Street`,
  Direction = Dir,
  CrossStreet = `Cross.Street`,
  District = Dist,
  CountDate = `Count.Date`,
  DayOfWeek = Day,
  WestBoundTraffic = `W.B`,
  EastBoundTraffic = `E.B`,
  NorthBoundTraffic = `N.B`,
  SouthBoundTraffic = `S.B`,
  TotalTraffic = Total
) %>% mutate(
  PrimaryStreet = gsub(' st ', 'st ', PrimaryStreet),
  PrimaryStreet = gsub(' nd ', 'nd ', PrimaryStreet),
  PrimaryStreet = gsub(' rd ', 'rd ', PrimaryStreet),
  PrimaryStreet = gsub(' th ', 'th ', PrimaryStreet),
  CountDate = ifelse(CountDate == 'January 10 0411', 'January 10 2011', CountDate),
  CountDate = as.Date(CountDate, format='%B %d %Y')
) %>% mutate( # these are data transformations I added later, after checking the data a bit more
  # Direction = 
  Type = str_trim(Type),
  Type = case_when(
    Type == "AUTO'" ~ 'AUTO',
    Type %in% c('MIO', 'AMAN') ~ 'MISC',
    TRUE ~ Type
  ),
  District = str_trim(District),
  
)
ladot_traffic_cleaned$CountDate %>% unique()
ladot_traffic_cleaned %>% arrange((CountDate))
ladot_traffic_cleaned %>% arrange((PrimaryStreet), (CrossStreet))
ladot_traffic_cleaned %>% filter(
  PrimaryStreet == '1st ST',
  CrossStreet == 'MC CADDEN PL'
) %>% group_by(District) %>% summarise(n())
```

```{r}
for (district in ladot_traffic_cleaned$District %>% unique()) {
  print(ladot_traffic_cleaned %>%
    filter(District == local(district)) %>%
    group_by(District, PrimaryStreet, CrossStreet) %>%
    summarise(n()))
}
```

```{r}
ladot_traffic_cleaned
```

## Time series graphs
```{r}
traffic_by_day <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  mutate(StreetCombo = paste0(PrimaryStreet, ' X ', CrossStreet, sep='')) %>%
  group_by(CountDate) %>% 
  summarise(
    TotalTraffic= sum(TotalTraffic, na.rm=TRUE),
    NumUniqueStreets = n_distinct(StreetCombo)
  )

traffic_by_week <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(TotalTraffic= sum(TotalTraffic, na.rm=TRUE))

print(traffic_by_day %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))
print(traffic_by_day %>%ggplot(aes(x=CountDate, y=NumUniqueStreets)) +
    geom_line() + ggtitle('Number of Streets Tracked by Date'))

print(traffic_by_week %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))

traffic_cleaned %>% filter(CountDate == as.Date('2011-01-05'))
```

### Looking at it by week and by area
```{r}
traffic_by_week <- traffic_cleaned %>% 
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

traffic_by_area_and_week <- traffic_cleaned %>% 
  group_by(AreaName, CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(traffic_by_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(traffic_by_area_and_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```
### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- traffic_cleaned %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```
### Occurred time distribution
```{r}
occurred_time_dist_over_time <- traffic_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(CountDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())

print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

## Scatterplot of Latitude, Longitude, 
```{r}
traffic_by_coordinates_and_year <- traffic_cleaned %>% 
  group_by(OccurredYear = year(CountDate), 
           Latitude = floor(Latitude*100)/100, 
           Longitude = floor(Longitude*100)/100) %>%
  summarise(Collisions = n()) %>%
  filter(Latitude != 0 | Longitude != 0) %>%
  arrange(Collisions)

for (year in c(2011:2025)) {
  print(
    traffic_by_coordinates_and_year %>% filter(OccurredYear == local(year)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point(size=4) +
      ggtitle(paste0('Collisions by Coordinates in ', year, sep=''))
  )
}
```

## Forecasting Overall Collision Counts
### Pre-COVID
```{r}
traffic_over_time_pre_covid <- traffic_cleaned %>% 
  filter(CountDate < as.Date('2020-03-15')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_pre_covid$CountDate)), 
                month(min(traffic_over_time_pre_covid$CountDate)))

traffic_pre_covid_ts <- ts(traffic_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_pre_covid_ts, round(length(traffic_pre_covid_ts) * 0.8))
h <- length(traffic_pre_covid_ts) - length(train)
test <- tail(traffic_pre_covid_ts, h)
```


```{r}
traffic_pre_covid_auto_model <- auto.arima(train)
traffic_pre_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train_dff1) # 3-4
```
#### MA
```{r}
acf(train_dff1) #2
```
#### ARIMA
```{r}
traffic_pre_covid_model <- Arima(train, order=c(3, 1, 2))
traffic_pre_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_pre_covid_auto_model$aic - traffic_pre_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_pre_covid_model, h=length(test))

autoplot(traffic_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

### Post-COVID
```{r}
traffic_over_time_post_covid <- traffic_cleaned %>% 
  filter(CountDate >= as.Date('2021-01-03')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_post_covid$CountDate)), 
                month(min(traffic_over_time_post_covid$CountDate)))

traffic_post_covid_ts <- ts(traffic_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_post_covid_ts, round(length(traffic_post_covid_ts) * 0.8))
h <- length(traffic_post_covid_ts) - length(train)
test <- tail(traffic_post_covid_ts, h)
```


```{r}
traffic_post_covid_auto_model <- auto.arima(train)
traffic_post_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train) # 1-2
```
#### MA
```{r}
acf(train) #3
```
#### ARIMA
```{r}
traffic_post_covid_model <- Arima(train, order=c(1, 0, 3))
traffic_post_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_post_covid_auto_model$aic - traffic_post_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_post_covid_model, h=length(test))

autoplot(traffic_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

## Time series graphs
```{r}
traffic_by_day <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  mutate(StreetCombo = paste0(PrimaryStreet, ' X ', CrossStreet, sep='')) %>%
  group_by(CountDate) %>% 
  summarise(
    TotalTraffic= sum(TotalTraffic, na.rm=TRUE),
    NumUniqueStreets = n_distinct(StreetCombo)
  )

traffic_by_week <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(TotalTraffic= sum(TotalTraffic, na.rm=TRUE))

print(traffic_by_day %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))
print(traffic_by_day %>%ggplot(aes(x=CountDate, y=NumUniqueStreets)) +
    geom_line() + ggtitle('Number of Streets Tracked by Date'))

print(traffic_by_week %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))

traffic_cleaned %>% filter(CountDate == as.Date('2011-01-05'))
```

### Looking at it by week and by area
```{r}
traffic_by_week <- traffic_cleaned %>% 
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

traffic_by_area_and_week <- traffic_cleaned %>% 
  group_by(AreaName, CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(traffic_by_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(traffic_by_area_and_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```
### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- traffic_cleaned %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```
### Occurred time distribution
```{r}
occurred_time_dist_over_time <- traffic_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(CountDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())

print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

## Scatterplot of Latitude, Longitude, 
```{r}
traffic_by_coordinates_and_year <- traffic_cleaned %>% 
  group_by(OccurredYear = year(CountDate), 
           Latitude = floor(Latitude*100)/100, 
           Longitude = floor(Longitude*100)/100) %>%
  summarise(Collisions = n()) %>%
  filter(Latitude != 0 | Longitude != 0) %>%
  arrange(Collisions)

for (year in c(2011:2025)) {
  print(
    traffic_by_coordinates_and_year %>% filter(OccurredYear == local(year)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point(size=4) +
      ggtitle(paste0('Collisions by Coordinates in ', year, sep=''))
  )
}
```

## Forecasting Overall Collision Counts
### Pre-COVID
```{r}
traffic_over_time_pre_covid <- traffic_cleaned %>% 
  filter(CountDate < as.Date('2020-03-15')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_pre_covid$CountDate)), 
                month(min(traffic_over_time_pre_covid$CountDate)))

traffic_pre_covid_ts <- ts(traffic_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_pre_covid_ts, round(length(traffic_pre_covid_ts) * 0.8))
h <- length(traffic_pre_covid_ts) - length(train)
test <- tail(traffic_pre_covid_ts, h)
```


```{r}
traffic_pre_covid_auto_model <- auto.arima(train)
traffic_pre_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train_dff1) # 3-4
```
#### MA
```{r}
acf(train_dff1) #2
```
#### ARIMA
```{r}
traffic_pre_covid_model <- Arima(train, order=c(3, 1, 2))
traffic_pre_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_pre_covid_auto_model$aic - traffic_pre_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_pre_covid_model, h=length(test))

autoplot(traffic_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

### Post-COVID
```{r}
traffic_over_time_post_covid <- traffic_cleaned %>% 
  filter(CountDate >= as.Date('2021-01-03')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_post_covid$CountDate)), 
                month(min(traffic_over_time_post_covid$CountDate)))

traffic_post_covid_ts <- ts(traffic_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_post_covid_ts, round(length(traffic_post_covid_ts) * 0.8))
h <- length(traffic_post_covid_ts) - length(train)
test <- tail(traffic_post_covid_ts, h)
```


```{r}
traffic_post_covid_auto_model <- auto.arima(train)
traffic_post_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train) # 1-2
```
#### MA
```{r}
acf(train) #3
```
#### ARIMA
```{r}
traffic_post_covid_model <- Arima(train, order=c(1, 0, 3))
traffic_post_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_post_covid_auto_model$aic - traffic_post_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_post_covid_model, h=length(test))

autoplot(traffic_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```


# Cleaned Traffic Data
## Load Data
```{r}
traffic_cleaned <- read.csv('Cleaned Data/LATrafficOverTime.csv') %>%
  mutate(
    TotalPeakHour = BackPeakHour + AheadPeakHour,
    TotalPeakMonth = BackPeakMonth + AheadPeakMonth,
    TotalAnnualAvgDaily = BackAnnualAvgDaily + AheadAnnualAvgDaily,
    
    RouteSuffix = ifelse(is.na(RouteSuffix), '', RouteSuffix),
    PMPrefix = ifelse(is.na(PMPrefix), '', PMPrefix),
    AlignCode = ifelse(is.na(AlignCode), '', AlignCode)
  ) %>%
  select(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode, Year, Description,
         TotalPeakHour, TotalPeakMonth, TotalAnnualAvgDaily) %>%
  as.data.frame()

# Check for duplicates
traffic_cleaned %>% 
  distinct() %>%
  drop_na() %>%
  group_by(District, Route, RouteSuffix, County, PMPrefix, 
           PostMile, AlignCode, Year, Description) %>%
  summarise(Count=n()) %>% arrange(desc(Count))

traffic_cleaned %>% 
  distinct() %>%
  drop_na() %>%
  group_by(District, Route, RouteSuffix, County, PMPrefix, 
           PostMile, AlignCode, Year, Description) %>%
  summarise(Count=n()) %>% arrange(desc(Count))
```

# Breakdown of Fields
## Categorical Fields
```{r, fig.height=5, fig.width=7}
for (col in c('TotalPeakHour', 'TotalPeakMonth', 'TotalAnnualAvgDaily')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram() +
      ggtitle(paste0(col, ' Distribution'))
  )
}

for (col in c('District', 'Route', 'County')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      group_by(FeatureName = local(col), TargetFeature, Year) %>%
      summarise(TotalAnnualAvgDaily = sum(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      group_by(FeatureName, TargetFeature) %>%
      summarise(AvgDailTraffic = mean(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      ggplot(aes(x=TargetFeature, y=AvgDailTraffic)) +
      geom_bar(stat='identity') +
      theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust=1)) +
      ggtitle(paste0(col, ' Distribution'))
  )
}
```

## Over Time
```{r}
for (col in c('District', 'County')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      group_by(FeatureName = local(col), TargetFeature, Year) %>%
      summarise(TotalAnnualAvgDaily = sum(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      ggplot(aes(x=Year, y=TotalAnnualAvgDaily, color=as.factor(TargetFeature))) +
      geom_line() + ggtitle(paste0('Traffic by ', col, ' Over Time'))
  )
}
```
# Try to Identify the Locations
```{r}
la_county_traffic <- traffic_cleaned %>% filter(County == 'LA')

la_county_traffic %>% select(District) %>% distinct()
la_county_traffic %>% select(Route) %>% distinct()
la_county_traffic %>% select(District, Route, PostMile) %>% distinct()
```

## Clean up the PostMile coordinate data
```{r}
postmile_coordinates <- read.csv('Data/State_Highway_Network_Postmiles_Tenth.csv') %>%
  mutate(PostMile = round(PM, 3))

postmile_coordinates %>% 
  group_by(District, Route, County, PostMile) %>% 
  summarise(Count = n(), X_Coords = paste0(X, collapse=','), 
            Y_Coords = paste0(Y, collapse=',')) %>%
  arrange(desc(Count))

postmile_coordinates %>% select(RteSuffix) %>% distinct()
postmile_coordinates %>% select(PMPrefix) %>% distinct()
postmile_coordinates %>% select(AlignCode) %>% distinct()
```

```{r}
postmile_coordinates %>% filter(District == 5, Route == 135, County == 'SB', PostMile == 9.095) # semi-duplicate seems to be the Prefixes, try removing
traffic_cleaned
```

```{r}

postmile_coordinates_cleaned <- data.frame()

for (i in 1:nrow(postmile_coordinates)) {
  coordinate_data <- postmile_coordinates[i, colnames(postmile_coordinates)]
  x <- coordinate_data['X'] %>% as.numeric()
  y <- coordinate_data['Y'] %>% as.numeric()
  lonlat <- sgo_en_wgs84(sgo_points(list(x, y), epsg=3857), to=4326)
  postmile_coordinates_cleaned <- rbind(postmile_coordinates_cleaned,
    cbind(coordinate_data[c('District', 'Route', 'County', 'PostMile')], 
          Longitude=lonlat['x'], Latitude=lonlat['y']))
}

postmile_coordinates_cleaned_2 <- postmile_coordinates_cleaned %>% 
  mutate(PostMile = round(PostMile, 3), x = round(x, 4), y = round(y, 5))
# postmile_coordinates_cleaned_2 %>% write.csv('Cleaned Data/LAPostMileCoordinates.csv', row.names=FALSE)
```

# Re-doing the PostMile Identification w/ Route Suffix, PostMile Prefix, and AlignCode
```{r}
postmile_coordinates2 <- postmile_coordinates %>% select(-AlignCode) %>% 
  mutate(PostMile = round(PM, 3), X = round(X), Y = round(Y)) %>%
  rename(RouteSuffix = RteSuffix, AlignCode = PMSuffix) %>%
  select(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode, X, Y)

postmile_coordinates_no_prefix_suffix <- postmile_coordinates2 %>%
  distinct() %>% group_by(District, Route, County, PostMile) %>% 
  mutate(Count=n()) %>% ungroup() %>% filter(Count == 1) %>% select(-Count, -X, -Y) %>%
  left_join(postmile_coordinates_cleaned_2 %>% distinct() %>%
              group_by(District, Route, County, PostMile) %>%
              summarise(x = min(x), y = min(y)), by=c('District', 'Route', 'County', 'PostMile'))

postmile_coordinates_with_prefix_suffix <- postmile_coordinates2 %>%
  distinct() %>% group_by(District, Route, County, PostMile) %>% 
  mutate(Count=n()) %>% ungroup() %>% filter(Count > 1)

postmile_coordinates_with_prefix_suffix_cleaned <- data.frame()

for (i in 1:nrow(postmile_coordinates_with_prefix_suffix)) {
  coordinate_data <- postmile_coordinates_with_prefix_suffix[i, colnames(postmile_coordinates_with_prefix_suffix)]
  x <- coordinate_data['X'] %>% as.numeric()
  y <- coordinate_data['Y'] %>% as.numeric()
  lonlat <- sgo_en_wgs84(sgo_points(list(x, y), epsg=3857), to=4326)
  postmile_coordinates_with_prefix_suffix_cleaned <- rbind(postmile_coordinates_with_prefix_suffix_cleaned,
    cbind(coordinate_data[c('District', 'Route', 'RouteSuffix', 'County', 
                            'PMPrefix', 'PostMile', 'AlignCode')], 
          Longitude=lonlat['x'], Latitude=lonlat['y']))
}

postmile_coordinates_full_cleaned_v2 <- postmile_coordinates_no_prefix_suffix %>% 
  rbind(postmile_coordinates_with_prefix_suffix_cleaned) 

postmile_coordinates_full_cleaned_v2 %>% write.csv('Cleaned Data/LAPostMileCoordinates_v2.csv', row.names=FALSE)
```

# Re-doing it again, because many areas have duplicate listings that seem to be updating the coordinates
```{r}
postmile_coordinates_full_cleaned_v2 %>%
  group_by(x, y) %>% mutate(Count = n()) %>% ungroup() %>%
  filter(Count > 1) %>% arrange(x, y)

postmile_coordinates_full_cleaned_v2 %>%
  distinct() %>%
  group_by(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode) %>% 
  mutate(Count = n()) %>% ungroup() %>%
  filter(Count > 1) %>% arrange(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode) %>%
  inner_join(traffic_cleaned %>% 
               group_by(District, Route, RouteSuffix, County, 
                        PMPrefix, PostMile, AlignCode) %>% 
               summarise(n()), by=c(
    'District', 'Route', 'RouteSuffix', 'County', 'PMPrefix', 'PostMile', 'AlignCode'))

updated_postmile_coordinates_df <- postmile_coordinates %>% 
  group_by(District, Route, RteSuffix, County, PMPrefix, PostMile, PMSuffix) %>%
  mutate(LatestObjectID = max(OBJECTID)) %>% ungroup() %>% filter(OBJECTID == LatestObjectID)

duplicated_postmile_coordinates <- postmile_coordinates_full_cleaned_v2 %>%
  distinct() %>% group_by(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode) %>% 
  summarise(Count = n()) %>% ungroup() %>% filter(Count > 1) %>% select(-Count) %>% 
  inner_join(updated_postmile_coordinates_df %>% select(-AlignCode) %>% 
               mutate(PostMile = round(PM, 3), X = round(X), Y = round(Y)) %>%
               rename(RouteSuffix = RteSuffix, AlignCode = PMSuffix) %>%
               select(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode, X, Y),
             by=c('District', 'Route', 'RouteSuffix', 'County', 'PMPrefix', 'PostMile', 'AlignCode'))

duplicated_postmile_coordinates_cleaned <- data.frame()

for (i in 1:nrow(duplicated_postmile_coordinates)) {
  coordinate_data <- duplicated_postmile_coordinates[i, colnames(duplicated_postmile_coordinates)]
  x <- coordinate_data['X'] %>% as.numeric()
  y <- coordinate_data['Y'] %>% as.numeric()
  lonlat <- sgo_en_wgs84(sgo_points(list(x, y), epsg=3857), to=4326)
  duplicated_postmile_coordinates_cleaned <- rbind(duplicated_postmile_coordinates_cleaned,
    cbind(coordinate_data[c('District', 'Route', 'RouteSuffix', 'County', 
                            'PMPrefix', 'PostMile', 'AlignCode')], 
          Longitude=lonlat['x'], Latitude=lonlat['y']))
}

postmile_coordinates_full_cleaned_v3 <- rbind(duplicated_postmile_coordinates_cleaned,
      postmile_coordinates_full_cleaned_v2 %>% distinct() %>% 
        group_by(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode) %>% 
        mutate(Count = n()) %>% ungroup() %>% filter(Count == 1) %>% select(-Count))


postmile_coordinates_full_cleaned_v3 %>% write.csv('Cleaned Data/LAPostMileCoordinates_v3.csv', row.names=FALSE)
postmile_coordinates_full_cleaned_v3 <- read.csv('Cleaned Data/LAPostMileCoordinates_v3.csv')
```

## Checking again for duplicates
```{r}
# no duplicate rows
postmile_coordinates_full_cleaned_v3 %>% 
  group_by(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode) %>%
  summarise(Count = n()) %>% arrange(desc(Count))

# there are still many coordinates that show up multiple times
postmile_coordinates_full_cleaned_v3 %>% 
  group_by(x, y) %>% summarise(Count = n()) %>% arrange(desc(Count))

# these coordinate errors won't interact with our traffic data negatively
postmile_coordinates_full_cleaned_v3 %>% 
  group_by(x, y) %>% mutate(Count = n()) %>% 
  ungroup() %>% filter(Count > 1) %>%
  inner_join(traffic_cleaned %>% 
               select(District, Route, RouteSuffix, County, 
                      PMPrefix, PostMile, AlignCode) %>% distinct(), 
             by=c('District', 'Route', 'RouteSuffix', 'County', 
                  'PMPrefix', 'PostMile', 'AlignCode')) %>%
  group_by(x, y) %>% summarise(Count = n()) %>% arrange(desc(Count))
```

## Checking the data is comprehensive
```{r}
# lots of these are empty; seems the PostMile numbers don't line up
traffic_cleaned %>% 
  left_join(postmile_coordinates_full_cleaned_v3, 
            by=c('District', 'Route', 'RouteSuffix', 'County', 
                 'PMPrefix', 'PostMile', 'AlignCode')) %>%
  filter(is.na(x), is.na(y))

# check the PostMile numbers to see how we can round them to kinda line up
postmile_values <- traffic_cleaned %>% select(PostMile) %>% 
  distinct() %>% mutate(Data='Traffic') %>%
  rbind(postmile_coordinates_full_cleaned_v3 %>% select(PostMile) %>% 
          distinct() %>% mutate(Data='Coordinates')) %>%
  group_by(PostMile) %>% summarise(Count = n(), Data = first(Data)) %>%
  arrange(PostMile)

# WOW that's a big gap; gotta see which one's are missing and whether I can get this data somehow
postmile_values %>% filter(Count == 1, Data == 'Traffic')
```

## Tracking the missing traffic data coordinates
```{r}
traffic_cleaned %>% 
  mutate(RoundedPostMile = round(PostMile, 1)) %>%
  left_join(postmile_coordinates_full_cleaned_v3 %>% 
              mutate(RoundedPostMile = round(PostMile, 1)) %>%
              group_by(Route, RouteSuffix, County, 
                       PMPrefix, RoundedPostMile, AlignCode) %>%
              summarise(x = mean(x), y = mean(y)), 
            by=c('Route', 'RouteSuffix', 'County', 
                 'PMPrefix', 'RoundedPostMile', 'AlignCode')) %>%
  filter(is.na(x), is.na(y))
```

```{r}
postmile_coordinates_full_cleaned_v3 %>% filter(round(y, 3) == 33.532)
traffic_cleaned %>% filter(PostMile == 9.006)
postmile_coordinates_full_cleaned_v3 %>% filter(Route == 58, County == 'SBD')
```

# Alternative Coordinate Data
```{r}
highway_network_properties <- sf::read_sf('Data/State_Highway_Network_Postmiles_Tenth.geojson')
highway_network_geometry <- read_json('Data/State_Highway_Network_Postmiles_Tenth.geojson')
highway_network_geometry_features <- highway_network_geometry$features
highway_network_coordinates <- data.frame(Longitude=c(), Latitude=c())

highway_network_properties[1, 'geometry']['Dimension']

for (geom in highway_network_geometry_features) {
  highway_network_coordinates <- rbind(highway_network_coordinates, 
    cbind(Longitude=geom$geometry$coordinates[[1]], 
          Latitude=geom$geometry$coordinates[[2]]))
}

traffic_cleaned %>% 
  mutate(PM = round(PostMile, 1)) %>%
  left_join(
    highway_network_properties %>% select(-AlignCode) %>%
      rename(RouteSuffix = RteSuffix, AlignCode = PMSuffix),
    by=c('Route', 'County', 'RouteSuffix', 'PM', 'PMPrefix', 'AlignCode')) %>%
  filter(is.na(OBJECTID))

highway_network_coordinates[1, 'geometry']
coordinates_df <- data.frame()
new_coordinates_row <- list(Latitude=0, Longitude=0)

for (coords in length(coordinates)) {
    if (coords %% 2 == 0) {
      new_coordinates_row['Latitude'] = coordinates[coords]
      coordinates_df <- rbind(coordinates_df, new_coordinates_row)
    } else {
      new_coordinates_row['Longitude'] = coordinates[coords]
    }
}
```

```{r}
highway_network_properties %>% filter(Route == 1, County == 'ORA',
                                      PM == 0.78)
```

# Combining PostMile Location Data w/ Traffic Volume
## Mapping the PostMile markers against traffic collisions data
```{r}
traffic_by_postmile <- traffic_cleaned %>% 
  group_by(District, Route, RouteSuffix, County, PMPrefix, 
           PostMile = round(PostMile, 1), AlignCode, Year, Description) %>%
  summarise(TotalPeakHour = mean(TotalPeakHour, na.rm=TRUE),
            TotalPeakMonth = mean(TotalPeakMonth, na.rm=TRUE),
            TotalAnnualAvgDaily = mean(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
  left_join(postmile_coordinates_full_cleaned_v3 %>%
              group_by(District, Route, RouteSuffix, County, PMPrefix, 
                       PostMile = round(PostMile, 1), AlignCode) %>%
              summarise(Longitude = mean(x, na.rm=TRUE),
                        Latitude = mean(y, na.rm=TRUE)), 
            by=c('District', 'Route', 'RouteSuffix', 'County', 
                 'PMPrefix', 'PostMile', 'AlignCode'))
traffic_by_postmile
```

```{r}
traffic_by_postmile %>% write.csv('Cleaned Data/LATrafficOverTimeByPostmile.csv', row.names=FALSE)
```


