---
title: "Traffic Collisions Location Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
library(Morpho)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Census Data
## Age trends over time
```{r}
print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    ggplot(aes(x=Year, y=Population, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group')
)

print(
  census_data %>%
    filter(DemographicType == 'Age') %>%
    mutate(Proportion = Population/TotalPopulation) %>%
    ggplot(aes(x=Year, y=Proportion, color=Demographic)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Age Group') +
    scale_y_continuous(labels=scales::percent)
)
```

## Breakdown of collisions by race/ethnicity
```{r, fig.height=4, fig.width=6}
for (race in c('Asian', 'Black', 'Hispanic', 'White')) {
  print(
    collisions_cleaned %>%
      filter(Latitude != 0, Longitude != 0, VictimEthnicity == local(race)) %>%
      group_by(Latitude = round(Latitude, 3), Longitude = round(Longitude, 3), VictimEthnicity) %>%
      summarise(Collisions = n()) %>%
      arrange(Collisions) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point() +
      xlim(-118.7, -118.1) +
      ylim(33.6, 34.4) +
      ggtitle(paste('Collision Density Map for ', local(race), ' Victims', sep=''))
  )
}
```

## Collision rates over time
```{r}
print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_cleaned %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    mutate(CollisionRate = Collisions/TotalPopulation) %>%
    ggplot(aes(x=Year, y=CollisionRate)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA))
)

print(
  census_data %>%
    group_by(Year) %>% summarise(TotalPopulation = first(TotalPopulation)) %>%
    left_join(
      collisions_cleaned %>% group_by(OccurredYear) %>% summarise(Collisions = n()),
      by = c('Year' = 'OccurredYear')
    ) %>%
    left_join(commuting_data_overall, by='Year') %>%
    mutate(CollisionRate = Collisions/TotalPopulation,
           CarCommutersByPopulation = (Driving + Carpooling)/TotalPopulation
           ) %>%
    ggplot(aes(x=CarCommutersByPopulation, y=CollisionRate, color=as.factor(Year))) +
    geom_point() +
    scale_y_continuous(labels=scales::percent, limits=c(0, NA)) +
    scale_x_continuous(labels=scales::percent) +
    labs(x='Number of Car Commuters per Population', y='Collisions per Population', color='Year')
)
```

# Commutes over Time
## Commuting trends over time
```{r}
# Proportion of each mode of transportation over time
print(
  commuting_data_overall %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)

print(
  commuting_data_by_ethnicity %>%
    pivot_longer(cols=c('Driving', 'Carpooling', 'PublicTransit', 
                        'Walking', 'TaxiBikeAndOther', 'WorkingFromHome'), 
                 names_to='ModeOfTransportation', values_to='Values') %>%
    mutate(Proportion = round(Values/TotalWorkers, 2)) %>%
    ggplot(aes(x=Year, y=Proportion, color=ModeOfTransportation)) +
    geom_line() +
    facet_wrap(~RaceEthnicity) +
    theme_set(theme_gray(base_size = 11)) +
    labs(color='Mode of Transportation') +
    scale_y_continuous(labels=scales::percent)
)
```

## Commuting shifts after COVID
```{r}
print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=CommutingByCarRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Commuting by Car Rate', fill='') +
    ggtitle('Proportion of Workers Commuting by Car, Pre- and Post-COVID')
)

print(
  commuting_by_covid_period %>%
    ggplot(aes(x=RaceEthnicity, y=WFHRate, fill=Period)) +
    geom_bar(stat='identity', position='dodge') +
    scale_y_continuous(labels=scales::percent) +
    labs(x='Race/Ethnicity', y='Working from Home Rate', fill='') +
    ggtitle('Proportion of Workers Working from Home, Pre- and Post-COVID')
)
```

## Comparing number of collision victims by ethnicity, compared to WFH rate
Too correlated w/ decrease in collisions after COVID
```{r}
print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=DrivingProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Driving to Work vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)

print(
  collisions_by_ethnicity %>% 
    filter(VictimEthnicity != 'Other', 
           VictimEthnicity != 'Two or More Races',
           OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear > 2020, 'Post-COVID', 'Pre-COVID')) %>%
    ggplot(aes(x=WFHProportion, y=CollisionRate, color=VictimEthnicity, shape=Period)) +
    geom_point(size=3) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Proportion of Workers Working from Home vs Collision Rate of Population', 
            subtitle='By Ethnicity')
)
```

## Distributions of collision time over the years
```{r}
print(
  collisions_cleaned %>%
    filter(!is.na(OccurredTime), OccurredYear >= 2016, 
           OccurredYear < 2025, OccurredYear != 2020) %>%
    mutate(Period = ifelse(OccurredYear < 2020, 'Pre-COVID', 'Post-COVID'),) %>%
    group_by(Period) %>% mutate(Total = n()) %>% ungroup() %>%
    group_by(OccurredTime = round(OccurredTime, -2), Period) %>%
    summarise(Distribution = n()/first(Total)) %>%
    ggplot(aes(x=OccurredTime, y=Distribution, color=Period)) +
    geom_line() +
    scale_y_continuous(labels=scales::percent) +
    ggtitle('Distribution of Collision Occurrence Time, Pre- and Post-COVID')
)
```

# Breaking Down Demographics by ZIP Code
## Observing the census and commuter data
```{r}
# sex
print(
  census_by_zip %>%
    filter(LabelGroup1 == 'SEX AND AGE', LabelGroup2 %in% c('MALE', 'FEMALE')) %>%
    group_by(Sex = LabelGroup2, Year) %>% summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=Sex)) + geom_line() + ylim(0, NA) +
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population by Sex in Relevant LA ZIP Codes over Time'))

print(
  census_by_zip %>%
    filter(LabelGroup1 == 'SEX AND AGE', LabelGroup2 %in% c('MALE', 'FEMALE')) %>%
    group_by(Year) %>% mutate(TotalPopulation = sum(Population, na.rm=TRUE)) %>% ungroup() %>%
    group_by(Sex = LabelGroup2, Year) %>% summarise(Proportion = sum(Population, na.rm=TRUE)/first(TotalPopulation)) %>%
    ggplot(aes(x=Year, y=Proportion, color=Sex)) + geom_line() + 
    scale_y_continuous(labels=scales::percent, limits = c(0, NA)) + 
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population Proportion by Sex in Relevant LA ZIP Codes over Time'))

# age
print(
  census_by_zip %>%
    filter(LabelGroup1 == 'SEX AND AGE', !(LabelGroup2 %in% c('MALE', 'FEMALE'))) %>%
    group_by(Age = LabelGroup2, Year) %>% summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=Age)) + geom_line() + ylim(0, NA) +
    facet_wrap(~Age) +
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population by Age in Relevant LA ZIP Codes over Time'))

print(
  census_by_zip %>%
    filter(LabelGroup1 == 'SEX AND AGE', !(LabelGroup2 %in% c('MALE', 'FEMALE'))) %>%
    group_by(Year) %>% mutate(TotalPopulation = sum(Population, na.rm=TRUE)) %>% ungroup() %>%
    group_by(Age = LabelGroup2, Year) %>% summarise(Proportion = sum(Population, na.rm=TRUE)/first(TotalPopulation)) %>%
    ggplot(aes(x=Year, y=Proportion, color=Age)) + geom_line() + 
    facet_wrap(~Age) +
    scale_y_continuous(labels=scales::percent, limits = c(0, NA)) + 
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population Proportion by Age in Relevant LA ZIP Codes over Time'))

# race/ethnicity
print(
  census_by_zip_pivoted %>%
    select(Year, White, Black, Indigenous, Asian, PacificIslander, Other, TwoOrMoreRaces) %>%
    pivot_longer(cols=c('White', 'Black', 'Indigenous', 'Asian', 'PacificIslander', 'Other', 'TwoOrMoreRaces'),
                 names_to = 'RaceEthnicity', values_to = 'Population') %>%
    group_by(RaceEthnicity, Year) %>% summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=RaceEthnicity)) + geom_line() + ylim(0, NA) +
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    labs(x='Year', y='Proportion', color='Race') +
    ggtitle('Population by Race in Relevant LA ZIP Codes over Time'))

print(
  census_by_zip_pivoted %>%
    select(Year, White, Black, Indigenous, Asian, PacificIslander, Other, TwoOrMoreRaces) %>%
    pivot_longer(cols=c('White', 'Black', 'Indigenous', 'Asian', 'PacificIslander', 'Other', 'TwoOrMoreRaces'),
                 names_to = 'RaceEthnicity', values_to = 'Population')  %>% 
    group_by(Year) %>% mutate(TotalPopulation = sum(Population, na.rm=TRUE)) %>% ungroup() %>%
    group_by(RaceEthnicity, Year) %>% summarise(Proportion = sum(Population, na.rm=TRUE)/first(TotalPopulation)) %>%
    ggplot(aes(x=Year, y=Proportion, color=RaceEthnicity)) + geom_line() + 
    scale_y_continuous(labels=scales::percent, limits = c(0, NA)) + 
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    labs(x='Year', y='Proportion', color='Race') +
    ggtitle('Population Proportion by Race in Relevant LA ZIP Codes over Time'))

print(
  census_by_zip_pivoted %>%
    select(Year, WhiteAlone, BlackAlone, IndigenousAlone, AsianAlone, 
           PacificIslanderAlone, OtherAlone, TwoOrMoreRacesAlone, Hispanic) %>%
    pivot_longer(cols=c('WhiteAlone', 'BlackAlone', 'IndigenousAlone', 'AsianAlone', 
                        'PacificIslanderAlone', 'OtherAlone', 'TwoOrMoreRacesAlone', 'Hispanic'),
                 names_to = 'RaceEthnicity', values_to = 'Population') %>%
    group_by(RaceEthnicity, Year) %>% summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=RaceEthnicity)) + geom_line() + ylim(0, NA) +
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    labs(x='Year', y='Proportion', color='Race/Ethnicity') +
    ggtitle('Population by Race/Ethnicity in Relevant LA ZIP Codes over Time'))

print(
  census_by_zip_pivoted %>%
    select(Year, WhiteAlone, BlackAlone, IndigenousAlone, AsianAlone, 
           PacificIslanderAlone, OtherAlone, TwoOrMoreRacesAlone, Hispanic) %>%
    pivot_longer(cols=c('WhiteAlone', 'BlackAlone', 'IndigenousAlone', 'AsianAlone', 
                        'PacificIslanderAlone', 'OtherAlone', 'TwoOrMoreRacesAlone', 'Hispanic'),
                 names_to = 'RaceEthnicity', values_to = 'Population') %>%
    group_by(Year) %>% mutate(TotalPopulation = sum(Population, na.rm=TRUE)) %>% ungroup() %>%
    group_by(RaceEthnicity, Year) %>% summarise(Proportion = sum(Population, na.rm=TRUE)/first(TotalPopulation)) %>%
    ggplot(aes(x=Year, y=Proportion, color=RaceEthnicity)) + geom_line() + 
    scale_y_continuous(labels=scales::percent, limits = c(0, NA)) + 
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    labs(x='Year', y='Proportion', color='Race/Ethnicity') +
    ggtitle('Population Proportion by Race/Ethnicity in Relevant LA ZIP Codes over Time'))
```

```{r}
print(
  commuting_by_zip_pivoted %>%
    pivot_longer(cols=c('Carpooling', 'Driving', 'PublicTransit', 'Biking', 'TaxiAndOther', 'Walking', 'WorkingFromHome'),
                 names_to = 'Commute', values_to = 'Population')  %>%
    group_by(Commute, Year) %>% summarise(Population = sum(Population, na.rm=TRUE)) %>%
    ggplot(aes(x=Year, y=Population, color=Commute)) + geom_line() + ylim(0, NA) +
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population by Commute in Relevant LA ZIP Codes over Time'))

print(
  commuting_by_zip_pivoted %>%
    pivot_longer(cols=c('Carpooling', 'Driving', 'PublicTransit', 'Biking', 'TaxiAndOther', 'Walking', 'WorkingFromHome'),
                 names_to = 'Commute', values_to = 'Population')  %>%
    group_by(Year) %>% mutate(TotalPopulation = sum(Population, na.rm=TRUE)) %>% ungroup() %>%
    group_by(Commute, Year) %>% summarise(Proportion = sum(Population, na.rm=TRUE)/first(TotalPopulation)) %>%
    ggplot(aes(x=Year, y=Proportion, color=Commute)) + geom_line() + 
    scale_y_continuous(labels=scales::percent, limits = c(0, NA)) + 
    scale_x_continuous(breaks=scales::breaks_pretty()) +
    ggtitle('Population Proportion by Commute in Relevant LA ZIP Codes over Time'))
```

## Comparing the census and commuter data to collisions data
```{r}
population_rates_graph(zip_codes_all_metrics, population = 'Hispanic', collision_type = 'HispanicVictim',
                       graph_title = 'Hispanic Population Proportion vs Collision Victim Proportion', 
                       legend_position='right')
```
```{r}
population_rates_graph(zip_codes_all_metrics, population = 'WhiteAlone', collision_type = 'WhiteVictim',
                       graph_title = 'White, Non-Hispanic Population Proportion vs Collision Victim Proportion', 
                       legend_position='right')
```

```{r}
collision_volume_graphs(
  zip_codes_all_metrics %>% mutate(All = 1), population = 'WorkingFromHome', collision_type = 'All',
  graph_title = 'Population Working From Home vs Total Collision Volume', legend_position = 'right')
```



