---
title: "Traffic Collisions MO Code Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.stat_bin.inform = FALSE)
```

# Comparing Intersections
```{r}
print(
  collisions_with_labeled_intersections %>% 
    filter(OccurredDate >= '2013-01-01',
           OccurredDate < '2025-03-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='months'),
             CollisionAtIntersection = ifelse(CollisionAtIntersection == 1, 
                                              'At Intersection', 'Not at Intersection')) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, 
               color=as.factor(CollisionAtIntersection))) +
    geom_line() +
    guides(
      colour = guide_legend(
        title = "Collision Location"
        #, theme = theme(legend.title = element_text(hjust = 0))
      )
    )
)
```

```{r}
print(
  collisions_with_labeled_intersections %>% 
    filter(OccurredDate >= '2013-01-01',
           OccurredDate < '2025-03-01') %>%
    group_by(OccurredWeek = floor_date(OccurredDate, unit='months'),
             CollisionAtIntersection = 
               case_when(CollisionAtIntersection == 1 & ListedAsIntersection == 1 ~ 'At Intersection', 
                         CollisionAtIntersection == 1 ~ 'Near Intersection',
                         TRUE ~ 'Not at Intersection')) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=OccurredWeek, y=Collisions, 
               color=as.factor(CollisionAtIntersection))) +
    geom_line() +
    guides(
      colour = guide_legend(
        title = "Collision Location"
        #, theme = theme(legend.title = element_text(hjust = 0))
      )
    )
)
```

## Seeing how often intersections have crashes in the vicinity that aren't at intersections - basically none
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(year(OccurredDate) == 2024,
           Latitude != 0, Longitude != 0) %>%
    group_by(RoundedLat = round(Latitude, 3), 
             RoundedLon = round(Longitude, 3)) %>%
    summarise(NumCollisionsAtIntersection = sum(as.numeric(CrossStreet != '')),
              NumCollisionsNotAtIntersection = sum(as.numeric(CrossStreet == '')),
              ProportionAtIntersection = round(NumCollisionsAtIntersection/
                (NumCollisionsAtIntersection + NumCollisionsNotAtIntersection), 2)) %>%
    ggplot(aes(x=ProportionAtIntersection)) +
    geom_histogram(stat='count')
)
```

## Comparing the severity of crashes at intersections vs non-intersections
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(OccurredYear == 2024) %>%
    mutate(WasCollisionAtIntersection = ifelse(CollisionAtIntersection == 1, 
                                               'At Intersection', 'Not at Intersection')) %>%
    group_by(WasCollisionAtIntersection) %>%
    mutate(TotalCollisions = n()) %>%
    group_by(WasCollisionAtIntersection, 
             MostSevereInjuryReported = factor(
               MostSevereInjuryReported, levels = c('No Injury', 'Complaint of Injury', 'Visible Injury',
                                                    'Severe Injury', 'Fatal Injury', 'Unspecified'))) %>%
    summarise(InjuryRate = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=MostSevereInjuryReported, y=InjuryRate, fill=WasCollisionAtIntersection)) +
    geom_bar(stat='identity') +
    facet_wrap(~WasCollisionAtIntersection) +
    theme(axis.text.x = element_text(angle = 70, vjust = 1, hjust=1))
)
```

## Looking for dangerous intersections
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(OccurredYear == 2024,
           CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>% 
    mutate(TotalCollisions = n()) %>%
    ungroup() %>%
    filter(TotalCollisions > 5) %>%
    group_by(Latitude, Longitude, 
             MostSevereInjuryReported = factor(MostSevereInjuryReported, 
                                               levels = c('No Injury', 'Complaint of Injury', 'Visible Injury',
                                                          'Severe Injury', 'Fatal Injury', 'Unspecified'))) %>%
    summarise(InjuryRate = n()/first(TotalCollisions)) %>%
    ggplot(aes(x=InjuryRate, fill=MostSevereInjuryReported)) +
    geom_histogram(binwidth = 0.1) +
    facet_wrap(~MostSevereInjuryReported)
)
```

```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(
      # OccurredYear == 2024,
      CollisionAtIntersection == 1,
      Latitude != 0, Longitude != 0) %>%
    group_by(
      RoundedLat = round(Latitude, 4), 
      RoundedLon = round(Longitude, 4)) %>%
    summarise(Collisions = n(), 
              NonInjuries = sum(as.numeric(NonInjury)),
              Injuries = sum(as.numeric(InjuryInvolved))) %>%
    ggplot(aes(x=Collisions, y=Injuries)) +
    geom_point()
)
```

```{r}
collisions_with_labeled_intersections %>%
  filter(
    # OccurredYear == 2024,
    CollisionAtIntersection == 1,
    Latitude != 0, Longitude != 0) %>%
  group_by(Latitude, Longitude) %>%
  summarise(Collisions = n(), 
            Injuries = sum(as.numeric(InjuryInvolved)),
            InjuryRate = mean(as.numeric(InjuryInvolved)),
            VisibleInjuryRate = mean(as.numeric(VisibleInjury)),
            SevereInjuryRate = mean(as.numeric(SevereInjury))
            ) %>%
  filter(Collisions > 50) %>%
  arrange(desc(InjuryRate))
```

# Collision Object
```{r}
print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    group_by(CollisionObject) %>%
    summarise(InjuryRate = mean(InjuryInvolved, na.rm=TRUE)) %>%
    ggplot(aes(x=CollisionObject, y=InjuryRate, fill=CollisionObject)) +
    geom_bar(stat='identity')
)

print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    mutate(SevereOrFatalInjury = as.numeric(SevereInjury | FatalInjury)) %>%
    group_by(CollisionObject) %>%
    summarise(SevereOrFatalInjuryRate = mean(SevereOrFatalInjury, na.rm=TRUE)) %>%
    ggplot(aes(x=CollisionObject, y=SevereOrFatalInjuryRate, fill=CollisionObject)) +
    geom_bar(stat='identity') +
    scale_y_continuous(labels = scales::percent) +
    labs(x='Collision Object', y='Severe/Fatal Injury Rate') +
    theme(legend.position='none') +
    ggtitle('Severe/Fatal Injury Rate by Collision Object in LA, 2024')
)

print(
  collisions_by_object %>%
    filter(OccurredYear == 2024) %>%
    group_by(CollisionObject, CollisionAtIntersection) %>%
    summarise(Collisions = n()) %>%
    ggplot(aes(x=CollisionObject, y=Collisions, fill=CollisionObject)) +
    geom_bar(stat='identity') +
    facet_wrap(~CollisionAtIntersection)
)
```

## Finding the intersections with the most/highest rate of pedestrian or bike collisions
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisionRate = round(mean(CollisionWithPedestrian), 2)) %>%
    filter(Collisions > 50) %>%
    arrange(desc(PedestrianCollisionRate)) %>%
    ggplot(aes(x=PedestrianCollisionRate)) +
    geom_histogram()
)

print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisions = sum(CollisionWithPedestrian)) %>%
    filter(Collisions > 50) %>%
    arrange(desc(PedestrianCollisions)) %>%
    ggplot(aes(x=PedestrianCollisions)) +
    geom_histogram()
)

```

## Map of pedestrian collision hot spots
```{r}
print(
  collisions_with_labeled_intersections %>%
    filter(CollisionAtIntersection == 1,
           Latitude != 0, Longitude != 0) %>%
    group_by(Latitude, Longitude) %>%
    summarise(
      Collisions = n(),
      PedestrianCollisions = sum(CollisionWithPedestrian),
      PedestrianCollisionRate = mean(CollisionWithPedestrian)
    ) %>%
    filter(Collisions >= 50) %>%
    mutate(
      PedestrianCollisionHotspot = case_when(
        PedestrianCollisions >= 25 & PedestrianCollisionRate >= 0.15 ~ 'High Rate and Volume',
        PedestrianCollisions >= 25 ~ 'High Volume',
        PedestrianCollisionRate >= 0.15 ~ 'High Rate',
        TRUE ~ 'Not Pedestrian Collision Hotspot'
      )
    ) %>%
    group_by(PedestrianCollisionHotspot) %>% mutate(Total = n()) %>% ungroup() %>%
    arrange(desc(Total)) %>%
    ggplot(aes(x=Longitude, y=Latitude, color=PedestrianCollisionHotspot)) +
    geom_point(size=1.7) +
    labs(x='', y='', color='Is Hotspot for\nPedestrian Collisions') +
    ggtitle('Map of Pedestrian Collisions Hotspots, LA')
)
```

# Using the Defined Function for Analysis
## Injury Type
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'MostSevereInjuryReported', 
             label='Injury Severity', year=2024)
```

### Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'FatalInjury', label='Fatal Injuries')

mapping_hotspots(collisions_with_labeled_intersections, 
                 'SevereInjury', label='Severe Injuries')
```

## Collision Object
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'MostVulnerableCollisionObject', 
             label='Collision Object', year=2024)
```

### Finding the intersections with the most/highest rate of pedestrian/bike collisions
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'CollisionWithPedestrian', label='Collision w/ Pedestrians')

mapping_hotspots(collisions_with_labeled_intersections, 
                 'CollisionWithBike', label='Collision w/ Bikes')
```

### Injuries by collision object
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'MostVulnerableCollisionObject', 
                      label1='Collision Object', label2='Collision Object')

graphing_injury_rates(collisions_with_labeled_intersections, 'CollisionWithPedestrian', 
                      label1='Collision w/ Pedestrians', label2='Collision w/ Pedestrians')
```

## Hit and Runs
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'HitAndRunSeverity', 
             label='Hit & Run Severity', year=2024)

intro_graphs(collisions_with_labeled_intersections, 'HitAndRun', 
             label='Hit & Run', year=2024)
```


### Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'HitAndRun', label='Hit & Runs')
```

### Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'HitAndRunSeverity', 
                      label1='Hit & Run Severity', label2='Hit & Run')

graphing_injury_rates(collisions_with_labeled_intersections, 'HitAndRun', 
                      label1='Hit & Run', label2='Hit & Run')
```

## DUIs
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'DUISeverity', 
             label='DUI Severity', year=2024)
```


### Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'DUI', label='DUIs')
```

### Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'DUISeverity', 
                      label1='DUI Severity', label2='DUI')

graphing_injury_rates(collisions_with_labeled_intersections, 'DUI', 
                      label1='DUI', label2='DUI')
```

## Private Property Involved
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'PrivatePropertyInvolvement', 
             label='Private Property Involved', year=2024)
```


### Finding the intersections with the most/highest rate of hit and runs
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'PrivatePropertyInvolved', label='Private Property')
```

### Injuries from hit & runs
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'PrivatePropertyInvolved', 
                      label1='Private Property Involvement', label2='Private Property Collision')
```

## Public Property Involved
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'CityePropertyInvolvement', 
             label='City Property Involvement', year=2024)
```


### Finding the intersections with the most/highest rate of public property involvement
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'CityPropertyInvolved', label='City Property')
```

### Injuries from public property collisions
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'CityePropertyInvolvement', 
                      label1='City Property Involvement', label2='City Property Collision')
```

## Unlicensed Motorists
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'UnlicensedMotorist', 
             label='Unlicenced Motorist', year=2024)
```

### Finding the intersections with the most/highest rate of unlicensed motorists
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'UnlicensedMotorist', label='Unlicenced Motorists')
```

### Injuries from unlicensed motorists
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'UnlicensedMotorist', 
                      label1='Unlicensed Motorist')
```

## Homeless Victim
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'HomelessVictim', 
             label='Homeless Victim', year=2024)
```


### Finding the intersections with the most/highest rate of homeless victimes
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 'HomelessVictim', label='Homeless Victims')
```

### Injuries to homeless victims
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'HomelessVictim', 
                      label1='Homeless Victim')
```

## Victim Descent
```{r, fig.height=2, fig.width=4}
intro_graphs(collisions_with_labeled_intersections, 'VictimEthnicity', 
             label='Victim Ethnicity', year=2024)

# Descent Code: A - Other Asian B - Black C - Chinese D - Cambodian F - Filipino G - Guamanian H - Hispanic/Latin/Mexican I - American Indian/Alaskan Native J - Japanese K - Korean L - Laotian O - Other P - Pacific Islander S - Samoan U - Hawaiian V - Vietnamese W - White X - Unknown Z - Asian Indian
```


### Finding the intersections with the most/highest rate of Hispanic victims
```{r, fig.height=2, fig.width=4}
mapping_hotspots(collisions_with_labeled_intersections, 
                 'HispanicOrLatinoVictim', label='Hispanic/Latino Victims')
```

### Injuries by Victim Ethnicity
```{r, fig.height=2, fig.width=4}
graphing_injury_rates(collisions_with_labeled_intersections, 'VictimEthnicity', 
                      label1='Victim Ethnicity')
```

