---
title: "Traffic Volume Exploratory Analysis"
output: html_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(readxl)
library(forecast)
library(tseries)
library(sgo)
```

# LADOT Data - insufficient
```{r}
ladot_traffic_df <- read.csv('Data/LADOT_Traffic_Counts_Summary.csv')

ladot_traffic_cleaned <- ladot_traffic_df %>% rename(
  PrimaryStreet = `Primary.Street`,
  Direction = Dir,
  CrossStreet = `Cross.Street`,
  District = Dist,
  CountDate = `Count.Date`,
  DayOfWeek = Day,
  WestBoundTraffic = `W.B`,
  EastBoundTraffic = `E.B`,
  NorthBoundTraffic = `N.B`,
  SouthBoundTraffic = `S.B`,
  TotalTraffic = Total
) %>% mutate(
  PrimaryStreet = gsub(' st ', 'st ', PrimaryStreet),
  PrimaryStreet = gsub(' nd ', 'nd ', PrimaryStreet),
  PrimaryStreet = gsub(' rd ', 'rd ', PrimaryStreet),
  PrimaryStreet = gsub(' th ', 'th ', PrimaryStreet),
  CountDate = ifelse(CountDate == 'January 10 0411', 'January 10 2011', CountDate),
  CountDate = as.Date(CountDate, format='%B %d %Y')
) %>% mutate( # these are data transformations I added later, after checking the data a bit more
  # Direction = 
  Type = str_trim(Type),
  Type = case_when(
    Type == "AUTO'" ~ 'AUTO',
    Type %in% c('MIO', 'AMAN') ~ 'MISC',
    TRUE ~ Type
  ),
  District = str_trim(District),
  
)
ladot_traffic_cleaned$CountDate %>% unique()
ladot_traffic_cleaned %>% arrange((CountDate))
ladot_traffic_cleaned %>% arrange((PrimaryStreet), (CrossStreet))
ladot_traffic_cleaned %>% filter(
  PrimaryStreet == '1st ST',
  CrossStreet == 'MC CADDEN PL'
) %>% group_by(District) %>% summarise(n())
```

```{r}
for (district in ladot_traffic_cleaned$District %>% unique()) {
  print(ladot_traffic_cleaned %>%
    filter(District == local(district)) %>%
    group_by(District, PrimaryStreet, CrossStreet) %>%
    summarise(n()))
}
```

```{r}
ladot_traffic_cleaned
```

## Time series graphs
```{r}
traffic_by_day <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  mutate(StreetCombo = paste0(PrimaryStreet, ' X ', CrossStreet, sep='')) %>%
  group_by(CountDate) %>% 
  summarise(
    TotalTraffic= sum(TotalTraffic, na.rm=TRUE),
    NumUniqueStreets = n_distinct(StreetCombo)
  )

traffic_by_week <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(TotalTraffic= sum(TotalTraffic, na.rm=TRUE))

print(traffic_by_day %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))
print(traffic_by_day %>%ggplot(aes(x=CountDate, y=NumUniqueStreets)) +
    geom_line() + ggtitle('Number of Streets Tracked by Date'))

print(traffic_by_week %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))

traffic_cleaned %>% filter(CountDate == as.Date('2011-01-05'))
```

### Looking at it by week and by area
```{r}
traffic_by_week <- traffic_cleaned %>% 
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

traffic_by_area_and_week <- traffic_cleaned %>% 
  group_by(AreaName, CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(traffic_by_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(traffic_by_area_and_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```
### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- traffic_cleaned %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```
### Occurred time distribution
```{r}
occurred_time_dist_over_time <- traffic_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(CountDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())

print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

## Scatterplot of Latitude, Longitude, 
```{r}
traffic_by_coordinates_and_year <- traffic_cleaned %>% 
  group_by(OccurredYear = year(CountDate), 
           Latitude = floor(Latitude*100)/100, 
           Longitude = floor(Longitude*100)/100) %>%
  summarise(Collisions = n()) %>%
  filter(Latitude != 0 | Longitude != 0) %>%
  arrange(Collisions)

for (year in c(2011:2025)) {
  print(
    traffic_by_coordinates_and_year %>% filter(OccurredYear == local(year)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point(size=4) +
      ggtitle(paste0('Collisions by Coordinates in ', year, sep=''))
  )
}
```

## Forecasting Overall Collision Counts
### Pre-COVID
```{r}
traffic_over_time_pre_covid <- traffic_cleaned %>% 
  filter(CountDate < as.Date('2020-03-15')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_pre_covid$CountDate)), 
                month(min(traffic_over_time_pre_covid$CountDate)))

traffic_pre_covid_ts <- ts(traffic_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_pre_covid_ts, round(length(traffic_pre_covid_ts) * 0.8))
h <- length(traffic_pre_covid_ts) - length(train)
test <- tail(traffic_pre_covid_ts, h)
```


```{r}
traffic_pre_covid_auto_model <- auto.arima(train)
traffic_pre_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train_dff1) # 3-4
```
#### MA
```{r}
acf(train_dff1) #2
```
#### ARIMA
```{r}
traffic_pre_covid_model <- Arima(train, order=c(3, 1, 2))
traffic_pre_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_pre_covid_auto_model$aic - traffic_pre_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_pre_covid_model, h=length(test))

autoplot(traffic_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

### Post-COVID
```{r}
traffic_over_time_post_covid <- traffic_cleaned %>% 
  filter(CountDate >= as.Date('2021-01-03')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_post_covid$CountDate)), 
                month(min(traffic_over_time_post_covid$CountDate)))

traffic_post_covid_ts <- ts(traffic_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_post_covid_ts, round(length(traffic_post_covid_ts) * 0.8))
h <- length(traffic_post_covid_ts) - length(train)
test <- tail(traffic_post_covid_ts, h)
```


```{r}
traffic_post_covid_auto_model <- auto.arima(train)
traffic_post_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train) # 1-2
```
#### MA
```{r}
acf(train) #3
```
#### ARIMA
```{r}
traffic_post_covid_model <- Arima(train, order=c(1, 0, 3))
traffic_post_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_post_covid_auto_model$aic - traffic_post_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_post_covid_model, h=length(test))

autoplot(traffic_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

## Time series graphs
```{r}
traffic_by_day <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  mutate(StreetCombo = paste0(PrimaryStreet, ' X ', CrossStreet, sep='')) %>%
  group_by(CountDate) %>% 
  summarise(
    TotalTraffic= sum(TotalTraffic, na.rm=TRUE),
    NumUniqueStreets = n_distinct(StreetCombo)
  )

traffic_by_week <- traffic_cleaned %>% 
  filter(!is.na(CountDate)) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(TotalTraffic= sum(TotalTraffic, na.rm=TRUE))

print(traffic_by_day %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))
print(traffic_by_day %>%ggplot(aes(x=CountDate, y=NumUniqueStreets)) +
    geom_line() + ggtitle('Number of Streets Tracked by Date'))

print(traffic_by_week %>%ggplot(aes(x=CountDate, y=TotalTraffic)) +
    geom_line() + ggtitle('Total Traffic by Date'))

traffic_cleaned %>% filter(CountDate == as.Date('2011-01-05'))
```

### Looking at it by week and by area
```{r}
traffic_by_week <- traffic_cleaned %>% 
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

traffic_by_area_and_week <- traffic_cleaned %>% 
  group_by(AreaName, CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(Collisions = n())

print(traffic_by_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
    geom_line() + ggtitle('Collisions by Week of Occurred Date'))
```

```{r, fig.width=10, fig.height=8}
print(traffic_by_area_and_week %>% ggplot(aes(x=CountDate, y=Collisions)) +
        facet_wrap(~AreaName) + geom_line() +
        ggtitle('Collisions by Area and Week of Occurred Date'))
```
### Mean and median occurred time over time
```{r}
occurred_time_stats_over_time <- traffic_cleaned %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>% 
  summarise(
    MeanOccurredTime = mean(OccurredTime, na.rm=TRUE),
    MedianOccurredTime = median(OccurredTime, na.rm=TRUE)
  )
print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MeanOccurredTime)) +
    geom_line() + ggtitle('Mean Collision Time by Week of Occurred Date'))

print(occurred_time_stats_over_time %>% ggplot(aes(x=CountDate, y=MedianOccurredTime)) +
    geom_line() + ggtitle('Median Collision Time by Week of Occurred Date'))
```
### Occurred time distribution
```{r}
occurred_time_dist_over_time <- traffic_cleaned %>% 
  mutate(
    Last2Digits = OccurredTime - floor(OccurredTime/100)*100,
    RoundedOccurredTime = case_when(
      Last2Digits >= 45 ~ floor(OccurredTime/100)*100 + 100,
      Last2Digits >= 15 ~ floor(OccurredTime/100)*100 + 30,
      TRUE ~ floor(OccurredTime/100)*100
    ),
    RoundedOccurredTime = ifelse(RoundedOccurredTime >= 2400, 
                                 RoundedOccurredTime-2400, RoundedOccurredTime)
    ) %>% 
  group_by(Year = year(CountDate), RoundedOccurredTime) %>%
  summarise(Collisions = n())

print(
  occurred_time_dist_over_time %>%
    ggplot(aes(x=RoundedOccurredTime, y=Collisions)) +
    geom_bar(stat='identity', width=50) +
    facet_wrap(~Year) +
    ggtitle('Distribution of Collision Occurrence Time by Year')
)
```

## Scatterplot of Latitude, Longitude, 
```{r}
traffic_by_coordinates_and_year <- traffic_cleaned %>% 
  group_by(OccurredYear = year(CountDate), 
           Latitude = floor(Latitude*100)/100, 
           Longitude = floor(Longitude*100)/100) %>%
  summarise(Collisions = n()) %>%
  filter(Latitude != 0 | Longitude != 0) %>%
  arrange(Collisions)

for (year in c(2011:2025)) {
  print(
    traffic_by_coordinates_and_year %>% filter(OccurredYear == local(year)) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=Collisions)) +
      geom_point(size=4) +
      ggtitle(paste0('Collisions by Coordinates in ', year, sep=''))
  )
}
```

## Forecasting Overall Collision Counts
### Pre-COVID
```{r}
traffic_over_time_pre_covid <- traffic_cleaned %>% 
  filter(CountDate < as.Date('2020-03-15')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_pre_covid$CountDate)), 
                month(min(traffic_over_time_pre_covid$CountDate)))

traffic_pre_covid_ts <- ts(traffic_over_time_pre_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_pre_covid_ts, round(length(traffic_pre_covid_ts) * 0.8))
h <- length(traffic_pre_covid_ts) - length(train)
test <- tail(traffic_pre_covid_ts, h)
```


```{r}
traffic_pre_covid_auto_model <- auto.arima(train)
traffic_pre_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train_dff1) # 3-4
```
#### MA
```{r}
acf(train_dff1) #2
```
#### ARIMA
```{r}
traffic_pre_covid_model <- Arima(train, order=c(3, 1, 2))
traffic_pre_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_pre_covid_auto_model$aic - traffic_pre_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_pre_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_pre_covid_model, h=length(test))

autoplot(traffic_pre_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```

### Post-COVID
```{r}
traffic_over_time_post_covid <- traffic_cleaned %>% 
  filter(CountDate >= as.Date('2021-01-03')) %>%
  group_by(CountDate = floor_date(CountDate, unit='weeks')) %>%
  summarise(Collisions = n())

start_time <- c(year(min(traffic_over_time_post_covid$CountDate)), 
                month(min(traffic_over_time_post_covid$CountDate)))

traffic_post_covid_ts <- ts(traffic_over_time_post_covid$Collisions, 
                    start = start_time, frequency = 52)

train <- head(traffic_post_covid_ts, round(length(traffic_post_covid_ts) * 0.8))
h <- length(traffic_post_covid_ts) - length(train)
test <- tail(traffic_post_covid_ts, h)
```


```{r}
traffic_post_covid_auto_model <- auto.arima(train)
traffic_post_covid_auto_model
```

#### Differencing
```{r}
train %>% adf.test()
train %>% ggtsdisplay(main='')

train_dff1 <- train %>% diff()
train_dff1 %>% adf.test()
train_dff1 %>% ggtsdisplay(main='')
```

#### AR
```{r}
pacf(train) # 1-2
```
#### MA
```{r}
acf(train) #3
```
#### ARIMA
```{r}
traffic_post_covid_model <- Arima(train, order=c(1, 0, 3))
traffic_post_covid_model
```

#### Comparing AIC
```{r}
exp((traffic_post_covid_auto_model$aic - traffic_post_covid_model$aic)/2)
```
#### Forecast
```{r}
forecast_1 <- forecast(traffic_post_covid_auto_model, h=length(test))
forecast_2 <- forecast(traffic_post_covid_model, h=length(test))

autoplot(traffic_post_covid_ts)
autoplot(forecast_1)
autoplot(forecast_2)
```


# Cleaned Traffic Data
## Load Data
```{r}
traffic_cleaned <- read.csv('Cleaned Data/LATrafficOverTimeByPostmile.csv')

# Check for duplicates
traffic_cleaned %>% 
  distinct() %>%
  drop_na() %>%
  group_by(District, Route, RouteSuffix, County, PMPrefix, 
           PostMile, AlignCode, Year, Description) %>%
  summarise(Count=n()) %>% arrange(desc(Count))

traffic_cleaned %>% 
  distinct() %>%
  drop_na() %>%
  group_by(District, Route, RouteSuffix, County, PMPrefix, 
           PostMile, AlignCode, Year, Description) %>%
  summarise(Count=n()) %>% arrange(desc(Count))
```

# Comparing to Pre-Joined Data
```{r}
traffic_cleaned_original <- read.csv('Cleaned Data/LATrafficOverTime.csv') %>%
  mutate(
    TotalPeakHour = BackPeakHour + AheadPeakHour,
    TotalPeakMonth = BackPeakMonth + AheadPeakMonth,
    TotalAnnualAvgDaily = BackAnnualAvgDaily + AheadAnnualAvgDaily,
    
    RouteSuffix = ifelse(is.na(RouteSuffix), '', RouteSuffix),
    PMPrefix = ifelse(is.na(PMPrefix), '', PMPrefix),
    AlignCode = ifelse(is.na(AlignCode), '', AlignCode)
  ) %>%
  select(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode, Year, Description,
         TotalPeakHour, TotalPeakMonth, TotalAnnualAvgDaily) %>%
  as.data.frame()

traffic_data_comparison <- traffic_cleaned_original %>%
  mutate(PM = PostMile, PostMile = round(PostMile, 1)) %>%
  left_join(traffic_cleaned, by=c('District', 'Route', 'RouteSuffix', 'County',
                                  'PMPrefix', 'PostMile', 'AlignCode', 'Year', 'Description'))

# No duplicates w/ the rounded PostMiles, so I actually didn't need to worry about aggregating
traffic_data_comparison %>% 
  group_by(District, Route, RouteSuffix, County, PMPrefix, PostMile, AlignCode, Year) %>%
  summarise(Count = n()) %>% arrange(desc(Count))
```

# Breakdown of Fields
## Categorical Fields
```{r, fig.height=5, fig.width=7}
for (col in c('TotalPeakHour', 'TotalPeakMonth', 'TotalAnnualAvgDaily')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      ggplot(aes(x=TargetFeature)) +
      geom_histogram() +
      ggtitle(paste0(col, ' Distribution'))
  )
}

for (col in c('District', 'Route', 'County')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      group_by(FeatureName = local(col), TargetFeature, Year) %>%
      summarise(TotalAnnualAvgDaily = sum(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      group_by(FeatureName, TargetFeature) %>%
      summarise(AvgDailTraffic = mean(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      ggplot(aes(x=TargetFeature, y=AvgDailTraffic)) +
      geom_bar(stat='identity') +
      theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust=1)) +
      ggtitle(paste0(col, ' Distribution'))
  )
}
```

## Over Time
```{r}
for (col in c('District', 'County')) {
  print(
    traffic_cleaned %>% rename(TargetFeature = as.symbol(col)) %>%
      group_by(FeatureName = local(col), TargetFeature, Year) %>%
      summarise(TotalAnnualAvgDaily = sum(TotalAnnualAvgDaily, na.rm=TRUE)) %>%
      ggplot(aes(x=Year, y=TotalAnnualAvgDaily, color=as.factor(TargetFeature))) +
      geom_line() + ggtitle(paste0('Traffic by ', col, ' Over Time'))
  )
}
```
## Heatmap
```{r}
collisions_max_lat <- (collisions_cleaned %>% group_by(1) %>%
  summarise(MaxLat = max(Latitude)))$MaxLat[1]
collisions_max_lon <- (collisions_cleaned %>% group_by(1) %>%
  summarise(MaxLon = max(Longitude)))$MaxLon[1]
collisions_min_lat <- (collisions_cleaned %>% group_by(1) %>%
  summarise(MinLat = min(Latitude)))$MinLat[1]
collisions_min_lon <- (collisions_cleaned %>% group_by(1) %>%
  summarise(MinLon = min(Longitude)))$MinLon[1]

traffic_by_coordinates_and_year <- traffic_cleaned %>% 
  group_by(Year, Latitude = floor(Latitude*100)/100, 
           Longitude = floor(Longitude*100)/100) %>%
  summarise(
    AvgPeakHour = mean(TotalPeakHour, na.rm=TRUE),
    AvgPeakMonth = mean(TotalPeakMonth, na.rm=TRUE),
    AnnualAvgDaily = mean(TotalAnnualAvgDaily, na.rm=TRUE)
  ) %>%
  filter(Latitude != 0 | Longitude != 0,
         # Latitude <= local(collisions_max_lat),
         # Latitude >= local(collisions_min_lat),
         # Longitude <= local(collisions_max_lon),
         # Longitude >= local(collisions_min_lon),
         Latitude <= local(34.4),
         Latitude >= local(33.6),
         Longitude <= local(-118.125),
         Longitude >= local(-118.75),
         )
```

```{r, fig.width=6, fig.height=4}
for (year in c(2013:2023)) {
  print(
    traffic_by_coordinates_and_year %>% filter(Year == local(year)) %>%
      arrange(AvgPeakHour) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=AvgPeakHour)) +
      geom_point(size=4) +
      ggtitle(paste0('Avg Peak Hour Traffic by Coordinates in ', year, sep=''))
  )
}
```

```{r, fig.width=6, fig.height=4}
for (year in c(2013:2023)) {
  print(
    traffic_by_coordinates_and_year %>% filter(Year == local(year)) %>%
      arrange(AvgPeakMonth) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=AvgPeakMonth)) +
      geom_point(size=4) +
      ggtitle(paste0('Avg Peak Month Traffic by Coordinates in ', year, sep=''))
  )
}
```

```{r, fig.width=6, fig.height=4}
for (year in c(2013:2023)) {
  print(
    traffic_by_coordinates_and_year %>% filter(Year == local(year)) %>%
      arrange(AnnualAvgDaily) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=AnnualAvgDaily)) +
      geom_point(size=4) +
      ggtitle(paste0('Annual Avg Daily Traffic by Coordinates in ', year, sep=''))
  )
}
```
# Intersection of Traffic Data and Collisions Data
Seems faulty, probably because the traffic data only views freeways, ignores lane width, intersection density
```{r}
rounding_factor <- 1000

traff_and_collisions_by_coord <- traffic_cleaned %>% 
  group_by(Latitude = floor(Latitude*local(rounding_factor))/local(rounding_factor), 
           Longitude = floor(Longitude*local(rounding_factor))/local(rounding_factor)) %>%
  summarise(
    AvgPeakHour = mean(TotalPeakHour, na.rm=TRUE),
    AvgPeakMonth = mean(TotalPeakMonth, na.rm=TRUE),
    AnnualAvgDaily = mean(TotalAnnualAvgDaily, na.rm=TRUE),
    NumTrafficRecords = n()
  ) %>%
  filter(Latitude != 0 | Longitude != 0,
         # Latitude <= local(collisions_max_lat),
         # Latitude >= local(collisions_min_lat),
         # Longitude <= local(collisions_max_lon),
         # Longitude >= local(collisions_min_lon),
         Latitude <= local(34.4),
         Latitude >= local(33.6),
         Longitude <= local(-118.125),
         Longitude >= local(-118.75),
         ) %>%
  inner_join(collisions_cleaned %>% 
               group_by(Address, CrossStreet, Latitude, Longitude) %>% 
               summarise(Collisions = n()) %>% 
               group_by(Latitude = floor(Latitude*local(rounding_factor))/local(rounding_factor), 
                        Longitude = floor(Longitude*local(rounding_factor))/local(rounding_factor)) %>% 
               summarise(TotalCollisions = sum(Collisions),
                         AvgCollisionsPerLocation = mean(Collisions, na.rm=TRUE),
                         NumCollisionRecords = n()) %>%
               filter(Latitude != 0 | Longitude != 0), 
             by=c('Latitude', 'Longitude'))

traff_and_collisions_by_coord
```

```{r, fig.width=6, fig.height=4}
for (year in c(2013:2023)) {
  print(
    traff_and_collisions_by_coord %>% filter(Year == local(year)) %>%
      arrange(TotalCollisions) %>%
      ggplot(aes(x=Longitude, y=Latitude, color=TotalCollisions)) +
      geom_point(size=4) +
      ggtitle(paste0('Collisions by Coordinates with Traffic Data in ', year, sep=''))
  )
}

print(
  traff_and_collisions_by_coord %>% 
    arrange(TotalCollisions) %>%
    ggplot(aes(x=Longitude, y=Latitude, 
               color=NumTrafficRecords, size=NumCollisionRecords)) +
    geom_point() +
    ggtitle(paste0('Data Records by Coordinates', sep=''))
)
```
```{r}
for (metric in c('AvgPeakHour', 'AvgPeakMonth', 'AnnualAvgDaily')) {
  print(traff_and_collisions_by_coord %>%
          rename(TargetFeature = local(metric)) %>%
          ggplot(aes(x=TargetFeature, y=AvgCollisionsPerLocation, 
                     color=NumTrafficRecords, size=NumCollisionRecords)) +
          geom_point() +
          ggtitle(paste0('Collisions vs ', local(metric), ' Traffic', sep='')))
}
```

```{r}
traff_and_collisions_by_coord %>% arrange(desc(Collisions))
```

```{r}
traff_and_collisions_by_coord %>% arrange(desc(AvgPeakHour))
```
